### 2016-11-02 paper/paper-v2.tex
draft v2 minor changes

\documentclass[a4paper]{article}

\usepackage{geometry}
\usepackage{natbib}
\bibpunct[:]{(}{)}{,}{a}{}{;}

%--------------------
%\usepackage{gb4e}
%\noautomath

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{mathrsfs}
\usepackage{nicefrac}
%\usepackage{stmaryrd}
%\usepackage{multicol}
\usepackage{graphicx}
\usepackage{color}
\usepackage{booktabs}
%\newcommand{\mvalueof}[1]{\llbracket#1\rrbracket}
\newcommand{\citeposs}[2][]{\citeauthor{#2}'s (\citeyear[#1]{#2})}
\newcommand{\tuple}[1]{\ensuremath{\left\langle #1 \right\rangle}}

\newcommand{\hl}[1]{\textcolor[rgb]{.8,.33,.0}{#1}}% prints in orange
%\newcommand{\argmax}[1]{\underset{#1}{\operatorname{arg}\,\operatorname{max}}\;}
%\newcommand{\argmin}[1]{\underset{#1}{\operatorname{arg}\,\operatorname{min}}\;}
%\newcommand{\sbna}{\exists\lnot\forall}

%--------------------
%
%\usepackage{setspace}
%\onehalfspacing
%
%-------------------

\title{Communicative pressures at the semantics-pragmatics interface}

\author{%\bf NAME1 and NAME2\\
    ( -- draft \today --- )
}

\date{}

\begin{document}
\maketitle

\begin{abstract}
According to standard linguistic theory, the conventional meaning of expressions is systematically modified by general pragmatic rules governing their use. This implies that models of the cultural evolution of meaning should likewise take into consideration that observable language use is a complex interaction of semantic representations and pragmatic rules. To this end, we present a game theoretic model of the cultural evolution of language where communicative pressures work on abstract semantic representations and pragmatic usage rules. Our model traces two evolutionary forces and their interaction: (i) fitness-based pressure towards communicative efficiency and (ii) systematic transmission perturbations when linguistic knowledge is transferred from one agent to another. The latter can contain general cognitive or learning biases, but can also contain other sources of systematic noise, e.g., resulting from errors in perception. We illustrate the model based on a case study showing that cognitive biases that favor simple semantic representations can prevent the lexicalization of pragmatic inferences. We also show that, technically speaking, it is possible that environmental factors, such as perceptual errors during acquisition, can produce evolutionary outcomes that look as if such cognitive biases are present even if they are not. 
\end{abstract}

%Mutual reasoning endows speakers with the ability to convey information that goes beyond the literal meaning of expressions. A particularity of certain classes of lexical meanings is that they enable for such pragmatic enrichments in a notably productive fashion. This raises the challenge to justify their regular selection over other alternatives that, e.g., do not require such enrichments. We suggest that such an account can be provided by an analysis of the linguistic pressures that apply at the semantics-pragmatics interface. To this end, we lay out a general model that combines functional pressure on language use and a pressure for learnability in acquisition in populations of probabilistic language users with varied degrees of pragmatic sophistication. We then apply this model in a case study on the (lack of) lexicalization of scalar implicatures. The results suggest that such inferences fail to lexicalize when languages are pressured towards simpler semantic representations in learning, provided that pragmatic reasoning can compensate for the disadvantage in expressivity that users of such languages otherwise incur. This result sheds light on the lack of lexicalization of scalar inferences, as well as the semantics-pragmatics distinction more generally.

\section{Introduction}\label{sec:introduction}
What is conveyed usually goes beyond what is said. A request for a blanket can be politely veiled by uttering ``I'm cold'', a temporal succession of events can be conveyed by the order in which conjuncts appear as in ``I traveled to Paris and got married'', or an invitation can be declined by saying ``I have to work''. An influential explanation of the relation between the literal meaning of expressions (semantics) and what they are intended and interpreted to convey (pragmatics) due to \citet{grice:1975} views the latter as a product of a process of mutual reasoning about rational language use. For instance, under the assumption that the speaker is cooperative and relevant, ``I have to work'' may be interpreted as providing a reason why the speaker will not be able to accept the invitation -- why would she have said so otherwise? Crucially, many classes of lexical meanings allow for pragmatic enrichments in a notably systematic fashion.

Productive pragmatic inferences such as scalar and manner implicatures have been at the focus of many studies on the semantics-pragmatics interface. However, an issue that has received little attention is the justification of semantic structure in light of the systematic enrichments offered by pragmatics. Similarly, multiple models of the cultural evolution of language have been put forward to analyze properties such as compositionality and combinatoriality, but the selection of particular semantics in light of pragmatics remains uncharted territory. The present investigation seeks to fill these gaps by analyzing the effects linguistic pressures have on the selection and pervasiveness of particular lexical meanings under consideration of their systematic pragmatic modification. Accordingly, our main goal is to uncover how transmission biases and functional pressure interact in the cultural evolution of meaning when pragmatics is taken into account. Crucially, we argue that transmission biases need not be cognitive in nature but can also ensue from perception and the environment.

\subsection{Linguistic pressures at the semantics-pragmatics interface}
The emergence and change of linguistic structure is influenced by many intertwined factors. These range from biological and socio-ecological to cultural ones \citep{steels:2011}. Social and ecological pressures determine communicative needs while biology determines the architecture that enables and constrains the means by which they can be fulfilled. In the following, our focus lies on cultural aspects, wherein processes of linguistic change are viewed as shaped by language use and its transmission. That is, as a result of a process of cultural evolution.

The idea that language is influenced by communicative pressures has played a pivotal role in synchronic and diachronic analyses at latest since \citeposs{zipf:1949} rationalization of the approximation of word frequency rankings by a power law distribution as competing hearer and speaker preferences (e.g. in \citealt{martinet:1962, horn:1984,jaeger+vRooij:2007,jaeger:2007, piantadosi:2014,kirby+etal:2015}). In recent years this line of research has led to a surge of approaches that seek to analyze the effects of such pressures from multiple angles ranging from simulations to experiments with human and robotic agents (see \citealt{steels:2015} and \citealt{tamariz+kirby:2016} for recent overviews). Our starting point is given by the overarching argument that has crystalized from this literature: Natural languages need to be well-adapted to communicative needs within a linguistic community, but also need to be learnable to survive their faithful transmission across generations. More succinctly; natural languages are pressured for expressivity and learnability.

The opposition of expressivity and learnability becomes particularly clear when considering their consequences in the extreme (cf. \citealt{kemp+regier:2012,kirby+etal:2015}). A language with a single form-meaning association is easy to learn but lacking in expressivity. Conversely, a language that associates a distinct form with all possible meanings one may want to convey is maximally expressive but challenging to acquire. The most prominent problem that arises from this tension is that of acquiring a language to express a potentially infinite set of meanings through finite means \citep{kirby:2002}. However, this is not the only challenge learners confront. More central to our explanandum is the issue of selecting particular hypotheses out of a potentially infinite space of alternatives compatible with the data learners are exposed to. At the semantics-pragmatics interface this concerns the selection between functionally (near-)equivalent lexical meanings, noting in particular that what is systematically conveyed through pragmatics could alternatively be codified lexically. Notwithstanding, lexical meanings that allow for pragmatic modifications often seem to resist the lexicalization of their pragmatic component. The question is why.

We assume an integral part of the answer to lie in the effects of transmission perturbations on language when passed from one generation to the next. Such perturbations may take the form of learning biases, but may also stem from extraneous factors such as a noisy perception of the environment. In particular, we show how the presence of systematic noise can introduce regularizations that resemble those obtained under the assumption of a learning bias.

In the following, we model these components using the replicator-mutator dynamics, combining functional pressure on successful communication with effects of transmission perturbations on (iterated) Bayesian learning \citep{griffiths+kalish:2007}. The semantics-pragmatics distinction and its bearing on production and comprehension are captured by a probabilistic model of rational language use with different degrees of pragmatic sophisitication and languages \citep{frank+goodman:2012,franke+jaeger:2014, bergen+etal:2016}. The remainder of this section introduces these components individually together with the assumptions underlying them. These are: the representation of languages and their use (\S\ref{sec:languages+use}), pressures towards expressivity (\S\ref{sec:expressivity}) and learnability (\S\ref{sec:learnability}) regulated by the replicator and mutator dynamics, respectively. After laying out the model, we analyze its predictions in a case-study on the lack of lexicalization of scalar implicatures in \S\ref{sec:si-case-study} and compare the outcome predicted by different transmission biases in \S\ref{subsec:bias} and \S\ref{subsec:noise}.

\subsection{Lexica and linguistic behavior}\label{sec:languages+use}
Lexica codify the truth-conditions of a language's expressions. Following \citet{franke+jaeger:2014}, a convenient way to represent lexica is by $(|S|,|M|)$-Boolean matrices, where $S$ is a set of states of affairs (meanings) to convey and $M$ a set of messages (forms available in the language).

We distinguish between two kinds of linguistic behavior. {\em Literal interlocutors} produce and interpret messages literally, being guided only by their lexica. In contrast, {\em pragmatic interlocutors} engage in mutual reasoning to inform their choices. Following models of rational language use such as Rational Speech Act models \citep{frank+goodman:2012} and their game-theoretic predecessors, the Iterated Best/Quantal Response models \citep{franke:2009,franke+jaeger:2014}, these behaviors are captured by a reasoning hierarchy. The hierarchy's bottom, level $0$, corresponds to literal language use. Pragmatic language users of level $n + 1$ behave rationally according to expected level $n$ behavior of their interlocutors. Below, (\ref{h:level0}) and (\ref{h:leveln}) specify the behavior of literal and pragmatic hearers of a language $L$. Their speaker counterparts are given in (\ref{s:level0}) and (\ref{s:leveln}).

\begin{flalign}
&H_{0}(s|m;L) \propto pr(s) L_{sm} \label{h:level0}\\
&S_{0}(m|s;L) \propto \exp(L_{sm}^\alpha) \label{s:level0}\\
&H_{n+1}(s|m;L) \propto pr(s) S_{n}(m|s;L) \label{h:leveln}\\
&S_{n+1}(m|s;L) \propto  \exp(\lambda \; H_{n}(s|m;L)^\alpha) \label{s:leveln}
\end{flalign}

According to (\ref{h:level0}), a literal hearer's interpretation of a message $m$ as a state $s$ depends on her lexicon and her prior over states, $pr \in \Delta(S)$. In the following this prior is assumed to be uniform across hearers. The behavior of literal speakers, given in (\ref{s:level0}), is regulated by a parameter $\alpha$ which controls the sensitivity to which speakers prefer one signal over another based on its expected communicative success.

Pragmatic behavior is similar to its literal counterparts. Their difference lies in that level $n+1$ speakers/hearers reason about level $n$ hearer/speaker behavior instead of solely relaying on their lexicon. That is, they reason about the way a rational level $n$ interlocutor would use or interpret a message, and behave according to these expectations.  Speaker behavior is further influenced by a soft-max parameter $\lambda$, $\lambda \geq 1$ \citep{luce:1959,sutton+barto:1998}. As $\lambda$ increases, choices made in production are more rational in that higher values lead to behavior that is increasingly in line with expected utility maximization.

We call the combination of a lexicon with its use, i.e., a level in the reasoning hierarchy, a {\em type}. These are the basic units on which our model's dynamics operate.

\subsection{Expressivity}\label{sec:expressivity}
Expressivity has received particular attention from investigations using evolutionary game theory (e.g. \citealt{nowak+krakauer:1999,nowak+etal:2000, nowak+etal:2002}). Under this view, a type's ability to convey and interpret information successfully confers it a higher fitness, a measure that is relative to the success of other types in the population. In the simplest models, fitness directly translates into the proportion of types present in the population after a generational turnover. This association of communicative success within a population with changes in the proportion of types present in it creates a feedback loop that pressures the population towards greater expressivity.

The replicator equation gives us the means to make these dynamics precise. The proportion of types in a given population is codified in a vector $x$, where $x_i$ is type $i$'s proportion. As noted above, the fitness of type $i$ is equal to its relative communicative success within this population, $f_i = \sum_j x_j \text{EU}(t_i,t_j)$. The expected communicative success of $i$ and $j$ is obtained by considering the average success of $i$ conveying information to $j$ and vice versa: $\text{EU}(t_i,t_j) = [U_S(t_i,t_j) + U_R(t_i,t_j)]/2$. $U_S(x,y)$ and $U_R(x,y)$ are, respectively, $\sum_s P(s)\sum_m S_n(m|s;L) \sum_{s'} R_o(s'|m;L) \delta(s,s')$ and $U_S(y,x)$, for $n$ and $o$ being the reasoning level of $x$ and $y$, and $\delta(s,s') = 1$ iff $s = s'$ and $0$ otherwise. This quantity is symmetric, reflecting the probability of two types' mutual understanding. The average fitness of the population is given by $\Phi$, $\Phi = \sum_i x_i f_i$. This term serves as a normalizing constant for the discrete replicator equation: $\dot{x}_i = \frac{x_i f_i}{\Phi}$

Under its biological interpretation, this equation captures the idea of fitness-relative selection whereby fitter types produce more offspring, leading to their propagation in subsequent generations. As noted above, many aspects of language are subject to processees of transmission and change that can be likened to these biological dynamics. Amongst others, replication can be construed as modelling language acquisition, as e.g. in \citealt{nowak+etal:2002}, but also as a process of horizontal adaptation in a single generation (see \citealt[\S3.3]{benz+etal:2005b} for discussion).

In their series of papers on language evolution, Nowak and colleagues did not only consider expressivity but also recognized the central role of the fidelity by which languages are transmitted. Crucially, linguistic production can be prone to errors, states or messages may be percieved incorrectly, and multiple languages may be compatible with the data learners are exposed to. These sources of uncertainty introduce variation in their transmission from one generation to the next. In keeping the analogy to evolutionary processes this variation can be likened to mutation to the effect that a type's offspring may adopt a different type than that of its parent. Importantly, the resulting generational turnovers should depend on the relative learnability of a type instead of being a constant that is equal to all types as in \citet{nowak+etal:2002}. For this purpose, we turn to a different strand of research in cultural evolution: {\em iterated learning}.

%errors were considered in N+K:1999, mutation in N+etal:2002

\subsection{Learnability}\label{sec:learnability}
Iterated learning is a process in which the behavior of one individual serves as learning input for another. This learner, upon acquisition of this behavior, then goes on to produce behavior that serves as input for a new learner. This process can be thought of as a progression through chains of parents and children; the parent produces linguistic data from which the child infers a language. The latter, now a parent, goes on to produce linguistic data for a new generation of learners. Following \citet{griffiths+kalish:2007} we model iterated learning as a repeated process of Bayesian inference in which learners combine the likelihood of a type producing the received learning input with prior inductive biases.

Due to the pressure towards learnability it exherts, iterated learning generally leads to simpler and more regular languages (see \citealt{kirby+etal:2014} and \citealt{tamariz+kirby:2016} for recent surveys). Importantly, experimental and mathematical results suggest that the outcome of this process reflects learners' a priori biases. In a Bayesian setting these biases can be codified in a prior $P \in \Delta(T)$ and a way to think about it is as the amount of data a learner would require in order to adopt a language (cf. \citealt[450]{griffiths+kalish:2007}). Or, in our case, a combination of a lexicon and a signaling behavior. The extent of the prior's influence has been shown to heavily depend on the learning strategy assumed to underly the inference process. On the one hand, simulation results suggested that weak biases could be magnified by exposing learners to only small data samples (e.g. in \citealt{brighton:2002}). On the other, \citeposs{griffiths+kalish:2007} mathematical characterization showed that iterated learning converged to the prior, i.e., the resulting distribution over languages corresponds to the learners' prior distribution and is not influenced by the amount of input given to them. This difference in predictions can be traced back to differences in the selection of hypotheses from the posterior. Griffith \& Kalish's convergence to the prior holds for learners that sample from the posterior. More deterministic strategies such as the adoption of the type with the highest posterior probability, so-called {\it maximum a posterior estimation} (MAP), increase the influence of both the prior and the data \citep{griffiths+kalish:2007,kirby+etal:2007}. In the following, we parametrize the posterior with a parameter $l \geq 1$. In doing so, we obtain a range of learning strategies between posterior sampling and MAP. When $l = 1$ learners sample from the posterior. As $l$ increases towards infinity their propensity to maximize the posterior grows as well.

The data learners are exposed to is described by a set $D$ containing sequences of state-message pairs, e.g., $\tuple{\tuple{s_i,m_v},...\tuple{s_j,m_w}}$. These are sequences of language use witnessed by learners, the length of which we denote by $k$. %Put differently, each datum $d \in D$ contains $k$ members of the set $\{\tuple{s_i,m_v} | s_i \in S, m_v \in M\}$.

%To summarize, the parametrized posterior, $P(d|t)^l$, is obtained from combining the prior $P(t)$ and the likelihood $P(d|t)$. Since types are combinations of a lexicon and its use, the latter can be computed straightforwardly from a type's production behavior. 
We integrated iterated learning in the replicator dynamics by codifying the former in a transition matrix $Q$, where $Q_{ij}$ indicates the probability that a child of a parent of type $i$ adopts type $j$. This quantity is proportional to the probability of $i$ producing the learning data and that of inferring $j$ given the data:

\[
 Q_{ij} \propto \sum_{d \in D} P(d|t_i) F(t_j,d), \; \text{where } F(t_j,d) \propto P(t_j|d)^l \text{ and } P(t_j|d) \propto P(t_j) P(d|t_j).
\]

\subsection{Summary}
Drawing from past research, we argued that expressivity and learnability are central to the cultural evolution of language. We propose these components to be modelled respectively as communicative efficiency-relative replication and (iterated) Bayesian learning. Taken together their interaction is described by the replicator-mutator dynamics \citep{hofbauer+sigmund:2003}:

\[ 
\hat{x}_i = \sum_j Q_{ji} \frac{x_jf_j}{\Phi}
\]

The units that the dynamics operate on are a combination of a lexicon and a degree of pragmatic sophistication determining its use. We call this combination a type. A type's expressivity depends on its communicative efficiency within a population while its learnability depends on the fidelity by which it is inferred by new generations of learners. The learners' task is consequently to perform a joint inference over types of linguistic behavior and lexical meaning. With this model at hand we turn to the analysis of the lack of lexicalization of productive pragmatic inferences in a case study on scalar implicatures.

\section{Scalar implicatures}\label{sec:si-case-study}
%
Scalar implicatures are a particularly well-studied type of conventional pragmatic inferences. They are licensed for groups of expressions ordered in terms of informativity, here understood as an entailment induced order. For instance, {\em some} is entailed by {\em all}. If it were true that `All students came to class', it would also be true that `Some students came to class'. However, while weaker expressions such as {\em some} are truth-conditionally compatible with stronger alternatives such as {\em all}, this is not what their use is normally taken to convey. Instead, the use of a less informative expression when a more informative one could have been used can license a defeasible inference that stronger alternatives do not hold (cf. \citealt{horn:1972,gazdar:1979}). That is, a hearer who assumes the speaker to be able and willing to provide all relevant information can infer that stronger alternatives do not hold because the speaker used a weaker alternative instead. In this way, `Some students came to class' is strengthened to convey that some but not all students came to class. Conversely, speakers can rely on their interlocutors to draw this inference. A bound that rules out stronger alternatives is thusly not codified in the lexical meaning of weak alternatives but instead pragmatically supplied.

This kind of strengthening is captured by the signaling behavior of pragmatic types introduced in \S\ref{sec:languages+use}: A pragmatic hearer who reasons about a speaker's use of message involving {\em some} will associate it more strongly with a state in which {\em some but not all} holds than with one in which the stronger alternative {\em all} holds, as the {\em all} already unambiguously conveys the latter state. Conversely, a pragmatic speaker will reason about her interlocutor's expected interpretation and use the messages at her disposition accordingly.

Our initial question can now be narrowed to the case of scalar implicatures by asking for a justification for the lack of lexical upper-bounds in weak scalar alternatives. That is, why they are regularly selected for over other alternatives such as that of codifying the bound semantically. More poignantly, would it not serve language users better if weak(er) expressions such as {\em warm}, {\em or}, {\em some} and {\em big} were truth-conditionally incompatible with stronger alternatives such as {\em hot}, {\em and}, {\em all} and {\em huge}?  This question is particularly striking considering the number of expressions that license such inferences across languages.

We see two main explanations for the lack of upper-bounds in the lexical meaning of weak scalar expressions. The first is that their truth-conditional compatibility with stronger expressions endows them with a broader range of applicability by allowing them to occur in contexts in which their upper-bounded reading is absent. This can happen when embedded in downward-entailing contexts, when the speaker is likely uncertain about whether the upper bounded reading is true, or when the distinction between an upper-bounded reading and the simple, only lower-bounded reading, is not relevant. For instance, if for all the speaker knows `Some students came' but she doesn't know whether all came, then the use of {\em some} lacking an upper-bound succinctly conveys her uncertainty. This may suggest a functionalist argument for why upper-bounded meanings do not conventionalize: should contextual cues provide enough information to the hearer to identify whether a bound is intended to be conveyed pragmatically, then this is preferred over expressing it overtly through longer expressions. For example, by saying {\em some but not all} explicitly. Importantly, although morphosyntactic disambiguation may be dispreferred due to its relative length and complexity \citep{piantadosi+etal:2012b}, it allows speakers to enforce an upper-bound and override contextual cues that might otherwise mislead the hearer. In a nutshell, this explanation posits that scalar implicatures fail to lexicalize because, all else being equal, speakers prefer to communicate as economically as possible and pragmatic reasoning enables them to do so. Compare this with a hypothetical language that lexicalizes two expressions for each weak scalar expression -- one with and one lacking an upper-bound. We see four conditions along this functionalist explanation that may pressure languages for English-like semantics over this alternative. First, contextual cues are very reliable. Second, morphosyntactic disambiguation is seldom necessary. Third, morphosyntactic disambiguation is only marginally dispreferred. Fourth, larger lexica are costly. Overall, neither condition is convincing as a pivotal explanatory device for such a wide-spread phenomenon. The first two conditions put a heavy burden on the ability to retrieve contextual cues to a degree that seems unlikely to undercut the benefit of unambiguous communication. It is likely that human language users are very good at retrieving cues from context, but to stipulate that they are so good as to undercut the benefit of safe communication provided by this hypothetical alternative strikes us as too strong of an assumption.  As for the third and fourth condition, these seem mostly like technical solutions without a proper empirical basis.

Instead, the systematicity and typological spread of scalar implicatures together with the observation that monomorphemic expressions that lexically rule out stronger alternatives are unattested across languages (\citealt[252-267]{horn:1984}, \citealt{horn:1972,traugott:2004,vdAuwera:2010}) suggests that other forces may be at play. In what follows we investigate the hypothesis that the lack of lexicalization of scalar inferences may be accounted for by the relative representational simplicity of lexical meanings lacking an upper-bound over those that explicitly codify it. This difference is reflected in a learning bias towards more compressed lexical representations. That is, rational learners should prefer simpler over more complex explanations of the data they witness \citep{feldman:2000, chater+vitanyi:2003, piantadosi+etal:2012a, kirby+etal:2015,piantadosi+etal:underreview}. More narrowly, codifying more semantically is a priori dispreferred and in cases where pragmatics offers conventional means of enrichment, the communicative disadvantage that speakers would otherwise incurr by the use of simpler semantics is leveled. In this way, these domains play a synergic role in overcoming linguistic pressures which may prevent the lexicalization of pragmatic inferences.

While we do not want to argue that functional aspects as the ones discussed above do not play a role, we do see a clear benefit in exploring whether matters of transmission biases would not give us additional explanatory leverage. Note however that we do not represent the contrast between lexical representations explicitly. Instead, the learning bias towards a lack of upper-bounds in weak scalar alternatives is directly encoded in the learners' prior over types. In principle this difference could be made precise with an adequate representational language, e.g., through measures over representational complexity such as minimal description length.  There is a growing effort to develop such empirically  testable  representational  languages. For  instance, the so-called {\em language of thought} has been put to test in various rational probabilistic models that show encouraging results (see e.g. \citealt{katz+etal:2008, piantadosi+etal:underreview, piantadosi+etal:2012a} and references therein). We think that our assumption is well-warranted as a working hypothesis and decide against such an enrichment at present in order to focus on the effects of linguistic pressures predicted by our model instead. In line with our aim to analyze the effects of transmission perturbations more broadly, we also show how noisy perception of the environment can lead to the same outcome as that predicted by a learning bias.

\subsection{Analysis}
We analyze the model's predictions in populations of types with two signaling behaviors; literal or pragmatic. The former correspond to level $0$ reasoners who only take their lexica into consideration and the latter to level $1$ reasoners. Higher level reasoning is not required to derive scalar implicatures from the lexica we consider here, nor do they leave room for substantial pragmatic refinement.

The space of lexica we consider is given in Table \ref{tab:lexica}. These $(2,2)$-Boolean matrices are the simplest ones that allow us to make the contrast between the presence or abscence of an upper-bound and the use of scalar implicatures precise. One may think of the state corresponding to the first row of a matrix as a ``some but not all''-state and the second as an ``all''-state. The literal meaning of weak scalar expressions such as English {\em some} then corresponds to a message true of both rows and rows in these fragments. While there are $16$ possible $(2,2)$-lexica, a number of them are identical both in terms of expressivity and the learning bias. The competition between such types is determined by the initial configuration of a population -- the proportion with which each type starts out. However, this fact can be obscured when averaging across simulations. Simulations conducted with the full space of possible lexica confirm that the results reported here do not hinge on their exclusion. We therefore focus on this smaller representative subset. Lexica $L_1$ to $L_3$ are not optimal for communication because they assign the same state to all their messages. This failure to associate a single form to a meaning both semantically and pragmatically inevitably leads to their disadvantage in language use. $L_4$ and $L_5$ are our target lexica. They codify upper-bounded semantics for the message corresponding to the first matrix's column and a lack thereof. Lastly, $L_6$ is similar to $L_5$ in that two messages are true of the same state but differs from it in assigning upper-bounded semantics to the first column's message.

\begin{table}[t]
\centering 
\begin{tabular}{l c l}
$L_1$ = $\begin{pmatrix} 0 & 0 \\ 1 & 1 \end{pmatrix}$ & 
$L_2$ = $\begin{pmatrix} 1 & 1 \\ 0 & 0 \end{pmatrix}$ & 
$L_3$ = $\begin{pmatrix} 1 & 1 \\ 1 & 1 \end{pmatrix}$\\[0.5cm]

$L_4$ = $\begin{pmatrix} 1 & 0 \\ 0 & 1 \end{pmatrix}$ &
$L_5$ = $\begin{pmatrix} 1 & 0 \\ 1 & 1 \end{pmatrix}$ &
$L_6$ = $\begin{pmatrix} 1 & 1 \\ 0 & 1 \end{pmatrix}$
\end{tabular}
\caption{Space of possible lexica.}
\label{tab:lexica}
\end{table}

Combining a signaling behavior with each of these $6$ lexica yields a total of $12$ distinct types. Note in particular  that a type that has conventionalized upper bounds to realize a (quasi-)partition of the relevant semantic space, such as $L_4$, will produce speaker behavior that is {\em almost} indistinguishable from that of a language that lacks upper bounds, but with pragmatic speakers, such as $L_5$. Almost, because there may be slight differences between the probability with which speakers would (erroneously) use a semantically false description and the probability with which speakers would (erroneously) use a pragmatically suboptimal description. Due to this possibly marginal difference between pragmatic $L_4$ and $L_5$, the selection of one type over the other is expected to mainly depend on the transmission bias. Things are less clear for literal $L_5$ contrasted with literal/pragmatic $L_4$ as the former has a learning advantage but is expected to fare worse in communicative terms.

The dynamics are initialized with an arbitrary distribution over types, constituting the population's first generation. The results for each parameter setting were obtained from $1000$ independent runs, each consisting of $20$ generations. This corresponds to a developmental plateau after which no noteworthy change was registered. As specified in \S\ref{sec:learnability}, the learning matrix $Q$ can be obtained by considering all possible state-message sequences of length $k$. Given that this is intractable for large $k$, matrices were approximated by sampling $10$ sequences from each type's production probabilities and a type's children being exposed only to this subset.

%For convenience, the model's parameters are summarized in Table \ref{tab:summary}.
%
%\begin{table}
%\centering
%\begin{tabular}{l|l|l}
%    \multicolumn{1}{c}{parameter} & \multicolumn{1}{c}{explanation} & \multicolumn{1}{c}{locus}\\ \hline
%    $\lambda \geq 1$ & rationality parameter & $S_{n+1}(m|s;L) \propto \text{exp}(\lambda \; H_{n}(s|m;L)^\alpha)$\\
%    $\alpha \in [0,1)$ & semantics-pragmatics tension & $S_{n+1}(m|s;L) \propto \text{exp}(\lambda \; H_{n}(s|m;L)^\alpha)$\\ 
%    $|D|$ & number of data produced per parent type & $P(d|t_j)P(t_i|d)$\\
%    $k = |d|$ & number of observations per datum& $P(d|t_j)P(t_i|d)$\\
%    $l \geq 1$ & posterior parameter from sampling to MAP & $P(t_i|d) \propto [P(t_i)P(d|t_i)]^l$\\
%    $c \in [0,1]$ & learning bias for lack of upper-bounds &  $P(t_i)$
%\end{tabular}
%\caption{Summary of model parameters.} 
%\label{tab:summary}
%\end{table}
\subsection{Learning bias}\label{subsec:bias}
Following our assumption of a preference for simple lexical representations, the prior biases learners against lexica in which a message holds true only of the first row. That is, those messages that lexicalize an upper-bound that rules out the ``all''-state. All other semantics are assumed to be a priori equally probable. Accordingly, this prior is captured by $P(t_i) \propto n - c \cdot r$, where $n$ is the total number of states and $r$ is the number of messages only true of row $1$ in $t_i$'s lexicon, $c \in [0,1]$. Increments in the prior parameter $c$ therefore bring about a stronger bias against languages that lexicalize upper-bounds, i.e., $L_2, L_4$ and $L_6$.

According to our hypothesis, functional pressure on successful communication combined with learning pressures in the form of a bias against upper-bounds may lead to the selection of $L_5$-like semantics. It is instructive to first inspect the effect of these pressures in isolation. For this purpose we focus our attention on three pragmatic types.\footnote{Pragmatic reasoning allows language users to refine their (possibly erroneous) choices. Therefore, it is advantageous even for those types that codify an upper-bound lexically.} Pragmatic $L_3$, a type that is lacking in expressivity but is a priori preferred for its lack of upper-bounds. Pragmatic $L_4$, a type that is functionally advantageous but biased against. And pragmatic $L_5$, combining virtues of the latter two.

\paragraph{Expressivity only.} The replicator dynamics are sensitive to $\lambda$ and $\alpha$ as both have a bearing on a type's fitness. The influence of the rationality parameter for $\alpha = 1$ is depicted in Figure \ref{fig:either-R-or-M}.A. As expected, the less expressive $L_3$ speakers fare the worse and are influenced the least by variations in $\lambda$. In contrast, low values of $\lambda$ result in a higher proportion of $L_4$ speakers relative to $L_5$. This is expected given role of rationality in producing more deterministic behavior in users of $L_5$-like languages. As the rationality parameter increases, the functional difference between $L_4$ and $L_5$ is leveled. Overall, the outcome from only a pressure towards expressivity come close to an even share of pragmatic $L_4, L_5$ and $L_6$ types (the latter follows the same trajectory as $L_5$ in Figure \ref{fig:either-R-or-M}.A).  This illustrates the problem we started out with: expressivity alone can not differentiate between (near-)functional equivalents to a degree that justifies the systematic prevalence of $L_5$-like semantics.

%In particular, low $\alpha$ disadvantages types that rely on pragmatic reasoning to the gain of those that codify more semantically. The rationality parameter $\lambda$ has a similar effect for different reasons: Low rationality leads to a less pronounced preference for the choice(s) expected to succeed best in communication. That is, $\lambda$ regulates the strength by which users of $L_5$ associate non-upper-bounded $m_1$ exclusively with the ``some''-state $s_1$ over the ``all''-state $s_2$. Speakers of $L_4$ need not rely on $\lambda$ for this as the association of $m_1$ with $s_1$ is already part of their language's semantics.

\begin{figure}
\centering
\includegraphics[scale=.5]{./only-R-or-M}
\caption{Mean proportions of target types after $20$ generations in $1000$ populations with only a pressure for expressivity in A ($\alpha = 1$) and only for learnability in B ($\alpha =1, \lambda = 30, k = 5$, $l=1$).}
\label{fig:either-R-or-M}
\end{figure}

\paragraph{Learnability only.} The effect of iterated learning with posterior sampling but without a pressure for expressivity is shown in Figure \ref{fig:either-R-or-M}.B. In line with our expectations, the share of $L_4$ speakers decreases as the bias against upper-bounds increases. In turn, this benefits $L_3$ and, in particular, $L_4$. However, note that even a strong bias against lexical upper-bounds leads only to a moderate advantage for $L_5$ over $L_4$. Crucially, a pressure only towards learnability can promote functionally defective languages such as $L_3$.

Inspecting these pressures separately not only gives some intuitions about the parameters' influence, but also highlights some of their broader implications. First and foremost, neither dynamic comes close to converging to a monomorphic population under most parameter configurations. For instance, while $L_4$ speakers can come to take over a substantial proportion of the population, this only happens in a restricted range of low degrees of rationality. Apart from polymorphy, both pressures make undesirable predictions. A pressure only towards expressivity leads to the to the ejection of $L_1$, $L_2$ and $L_3$ from the population but it can not explain the regular selection of $L_5$-like semantics over either of its functionally similar alternatives. In contrast, a pressure only towards learnability has a modest but clear effect in differentiating $L_5$ from these alternatives but fails to rule out functionally suboptimal types such as tautological $L_3$.

\begin{figure}
\centering
\includegraphics[scale=.5]{./fig2-rmd}
\caption{Mean proportions of target types after $20$ generations in $1000$ populations across bias values $c \in [0,1]$ with $l =1$ in A and $l = 3$ in B ($\alpha =1, \lambda = 20, k = 5$).}
\label{fig:cost}
\end{figure}

\paragraph{Expressivity and learnability.} Figure \ref{fig:cost} illustrates the effect of the learning bias for posterior sampling (\ref{fig:cost}.A) and slightly more MAP-like learning (\ref{fig:cost}.B) when pressured for both expressivity and learnability.\footnote{\hl{Robert suggests to illustrate the effect with a higher l-value. E.g., l = 1 and l = 10}} More detailed results for all types across a sample of $c$-values for $l = 1$ and $l = 3$ are presented in Table \ref{tab:numeric-results}. These results show that a weak bias is sufficient to lead to a selection of $L_5$ over $L_4$ in the present setup. As in the outcomes that only considered learnability, this effect increases with the bias' strength provided $L_5$ users are pragmatic. Importantly, the addition of a pressure towards expressivity magnifies this effect and dampens the proliferation of functionally suboptimal types advantaged by the learning bias. As stressed above, this indicates that neither a learning bias nor functional pressure alone but their combination may lead to the lack of upper-bounds in the lexical meaning of scalar expressions.

\begin{table}
\centering 
\begin{tabular}{l | c c c c c| c c c c c c c c}
\multicolumn{1}{c}{~} & \multicolumn{5}{c}{$l = 1$} & ~ & \multicolumn{5}{c}{$l = 3$}\\ \hline \hline
  c        &  0  & .1  & .5  & .8  & .9 & ~ & 0 & .1 & .5 & .8 & .9\\ \hline \hline
lit. $L_1$ & .03 & .03 & .04 & .04 & .04& ~ & 0 &0 & 0 & 0 & 0\\ 
lit. $L_2$ & .03 & .03 & .02 & .01 & .04& ~ & 0& 0 & 0 & 0 & 0\\
lit. $L_3$ & .03 & .03 & .04 & .04 & .04& ~ & 0& 0 & 0 & 0 & 0\\
lit. $L_4$ & .07 & .07 & .06 & .06 & .05& ~ & 0& 0 & 0 & 0 & 0\\
lit. $L_5$ & .04 & .05 & .05 & .06 & .06& ~ & 0& 0 & 0 & 0 & 0\\
lit. $L_6$ & .04 & .04 & .04 & .04 & .03& ~ & 0& 0 & 0 & 0 & 0\\ \hline
prg. $L_1$ & .03 & .03 & .04 & .04 & .04& ~ & 0& 0 & 0 & 0 & 0 \\
prg. $L_2$ & .03 & .03 & .02 & .01 & .04& ~ & 0& 0 & 0 & 0 & 0 \\
prg. $L_3$ & .03 & .03 & .04 & .04 & .04& ~ & 0& 0 & 0 & 0 & 0 \\ 
prg. $L_4$ & .22 & .22 & .2 & .18 & .17& ~ & .33& .31 & .23 & .15 & .12 \\
prg. $L_5$ & .22 & .23 & .27 & .3  & .32& ~ & .33& .37 & .54 & .7 & .75 \\
prg. $L_6$ & .22 & .22 & .2 & .18 & .17& ~ & .33& .31 & .23 & .15 & .12
\end{tabular}
\caption{Mean proportions of types in $1000$ populations after $20$ generations across bias values $c \in [0,1]$ with $l =1$ and $l = 3$ ($\alpha = 1, \lambda = 30, k = 5$)}%, $\epsilon < 0.005$.}
\label{tab:numeric-results}
\end{table}

The resulting proportion of pragmatic $L_5$ speakers primarily hinges on four parameters. First, the degree to which linguistic behavior is deterministic, controlled by $\lambda$ and $\alpha$, as it plays a role both for expressivity as well as in producing data that allows learners to discriminate this type from others. Second,  the learning bias $c$ which controls the learners preference for simpler lexical representations. Lastly, the posterior parameter $l$, which magnifies the effects of the learning bias in tandem with replication.

As discussed in relation to Figure \ref{fig:cost}.A, posterior sampling can lead to the incumbency of pragmatic $L_5$. However, not even a strong favorable learning bias combined with a pressure for expressivity completely drives out competing types. This is not so for more posterior maximizing behavior. As shown in Figure \ref{fig:prior-posterior}, the range of bias values within which $L_5$ takes over the population increases with MAP-like learning. In other words, the strength of the learning bias required for a given final proportion of $L_5$ speakers strongly depends on learners' inference mechanism. As for the effect of the other parameters not mentioned so far, changes in sequence length influence the population in a predictable way: smaller values lead to more heterogeneous populations that reflect the learner's prior more faithfully. Larger ones lead to more pronounced differences amongst equally preferred types. This is expected insofar as the likelihood that a small sequence was produced by any type is relatively uniform (modulo prior) whereas the likelihood of types with lexica $L_1$ - $L_3$ to produce larger sequences with the same state-message combination is less likely than for pragmatic types using $L_4$ - $L_6$, or literal $L_4$. Thus, while noteworthy, sequence length has no direct bearing on the contrast of interest.

\begin{figure}
\centering
\includegraphics[scale=.5]{../presentations/01heatmap}
\caption{Mean proportion of pragmatic $L_5$ in $1000$ populations after $20$ generations ($\alpha = 1, \lambda = 30, k = 5$)}
\label{fig:prior-posterior}
\end{figure}

\subsection{Noisy transmission}\label{subsec:noise}
The assumption of a learning bias was pivotal in introducing transmission perturbations that differentiate similar types and homogenize the population. However, such perturbations need not necessarily arise from cognitive factors. Instead, as shown in the following, similar outcomes can arise from environmental actors, such as perceptual errors during production and comprehension. We should stress that our goal is not to suggest noisy perception to underlie the selection of a lack of semantic upper-bounds in scalar implicatures, but to elucidate the role of transmission perturbations in this process and the cultural evolution of meaning more broadly.

The core components of the model remain as before. All that is required is an adjustment to the possibility of confusing one state of affairs with another. Letting $\epsilon$ stand for the probability of confusing the first row state with the second, and vice-versa for $\delta$, we denote the probability that the teacher (learner) observes state $s_t$ ($s_l$) when the actual state is $s_a$ as $P_N(s_t \mid s_a)$ ($P_N(s_l \mid s_a)$). The probability that $s_a$ is the actual state when the learner observes $s_l$ is just Bayes rule, combining prior of $s_a$ with the noise from above:
\begin{align*}
  P_N(s_a \mid s_l) \propto P(s_a) \ P_N(s_l \mid s_a)\,.
\end{align*}
Secondly, the probability that the teacher observes $s_t$ when the learner observes $s_l$ is:
\begin{align*}
  P_N(s_t \mid s_l) = \sum_{s_a} P(s_a \mid s_l) \ P_N(s_t \mid s_a)\,.
\end{align*}
Finally, this gives us the probability that a teacher of type $t$ produces a datum that is
perceived by the listener as $d = \tuple{s_l, m}$:
\begin{align*}
  P_N(\tuple{s_l, m} \mid t) = \sum_{s_t} P_N(s_t \mid s_l) \ P(m \mid s_t, t)\,.
\end{align*}
Generalize this to a sequence of perceived data $d_l$ and write $P_N(d_l \mid t)$. We then
define the noise-perturbed mutation matrix as:
\begin{align*}
  Q_{ij}  \propto \sum_{d_l \in D} P(d_l \mid t_i) F(t_j,d_l) \,, \ \  \text{where $F(t_j,d)$
    as before.}
\end{align*}
In words, it may be the case that learner and/or teacher do not perceive the actual state as what it is. They are not aware of this, and produce/learn as if what they observed was the actual state. In particular, the learner does not reason about noise when he tries to infer the speaker's type. He takes what he observes a state to be as the actual state that the teacher has seen as well and infers which type would have most likely generated the message to this state. This can lead to biases of inferring the ``wrong'' teacher type, if the noise makes some types err in a way that resembles the noiseless behavior of other types. That is, such environmental factors can, in principle, induce transmission biases that look as is there was a cognitive bias in favor of a particular type, simply because that type better explains the noise.

Apart from changing the mutation matrix in this way, we also need to adapt the calculation of expected utilities, taking into consideration that states are perceived noisily:
\begin{align*}
  U_S(t_i,t_j) = \sum_{s_a}  P(s_a) \sum_{s_t} P_N(s_t \mid s_a) \sum_m S_n(m|s_t;L) \sum_{s'} R_o(s'|m;L) \delta(s,s')\,.
\end{align*}

\paragraph{Results.} As shown in Table \ref{tab:num-noise}, a prevalence of pragmatic $L_5$ can also arise from noisy transmission without a prior  learning bias for particular types. In contrast to the outcome of the non-noisy model, favourable outcomes for this type concentrate in noise patterns with $\epsilon > \delta$, $\alpha \geq 5$, $1 < k < 20$ and $l > 3$. Crucially, these results show that the mere presence of systematic noise in the transmission of strategies can introduce regularization that looks as if the agents have a learning bias. But, most importantly, this disturbance of transmission fidelity is due to perceptual noise and properties of the environment, not learning biases. In other words, learning biases are clearly not the only transmission biases that can shape evolution alongside functional pressure. Environmental and perceptual noise can play a role too.

\begin{table}
\centering 
\begin{tabular}{l | c c c c c| c c c c c c c c}
\multicolumn{1}{c}{~} & \multicolumn{5}{c}{$l = $} & ~ & \multicolumn{5}{c}{$l = 15$}\\ \hline \hline
 ($\epsilon,\delta$)        &  (.1,.1)  & (.1,.3)  & (.3,.1)  & (.8,.1)  & (.1,.8) & ~ & (.1,.1) & (.1,.3) & (.3,.1) & (.8,.1) & (.1,.8)\\ \hline \hline
lit. $L_1$ & .01 & .02 & .03 & .03 & .03& ~ & 0& .02&  .02 & .01 & .01\\ 
lit. $L_2$ & .01 & .02 & .03 & .03 & .03& ~ & 0& .02&  .02 & .01 & .01\\
lit. $L_3$ & .01 & .02 & .03 & .01 & .03& ~ & 0& .02 & .02 & .01 & .01\\
lit. $L_4$ & .03 & .03 & .03 & .05 & .01& ~ & .04& .03 & .03 & .01 & .01\\
lit. $L_5$ & .02 & .04 & .03 & .02 & .02& ~ & .01& .03 & .02 & .02 & .01\\
lit. $L_6$ & .02 & .03 & .04 & .03 & .05& ~ & .01& .03 & .04 & .01 & .02\\ \hline
prg. $L_1$ & .01 & .02 & .03 & .03 & .03& ~ & 0&  .02& .02 & .01 &  .01\\
prg. $L_2$ & .01 & .02 & .03 & .03 & .03& ~ & 0&  .02& .02& .01&  .01\\
prg. $L_3$ & .01 & .02 & .03 & .03 & .03& ~ & 0&  .02& .02 & .01 & .01 \\ 
prg. $L_4$ & .45 & .11 & .11 & .02 & .02& ~ & .44& .11 & .1 & .22 & .02 \\
prg. $L_5$ & .22 & .17 & .47 & .04 & .68& ~ & .24& .18 & .56 & .02 & .84\\
prg. $L_6$ & .22 & .48 & .17 & .69 & .04& ~ & .24& .52 & .16 & .84 & .02
\end{tabular}
\caption{\hl{Currently: Mean proportions of types from $50$ runs per $100$ independently generated Q-matrices per parameter setting after $20$ generations across noisy values for $\epsilon$ and $\delta$ with $l = 5$ and $l = 10$ ($\alpha = 10, \lambda = 20, k = 5$). In principle, we could skip every second row because pragmatic $L5$ and $L6$ mirror each other.}}
\label{tab:num-noise}
\end{table}

\subsection{Discussion}
Under the assumption of a learning bias for simpler representations, our results suggest that a lack of semantic upper-bounds coupled with pragmatic reasoning can overcome communicative pressures and stabilize in a population. This prediction hinges on three assumptions. First, that language is pressured toward both expressivity and learnability. Second, that language use is not too unpredictable -- low $\alpha$ or $\lambda$ render languages that rely on pragmatics too prone to communicative failure and more difficult to learn. Lastly, that learners prefer simpler over more complex lexical representations. An important addendum to this third condition being that a combination of rationality in choice and maximization in learning requires a weaker bias towards simplicity. Under these conditions the selection of lexical meanings lacking upper-bounds in populations of pragmatic speakers is robust against parameter perturbations.  This outcome is particularly encouraging in light of other advantages a lack of semantic upper-bounds may confer.

While a major share of non upper-bounded lexical meanings is predicted by the literature, it is less clear to what extent other types should be present in the final population, if at all. It seems reasonable to expect functionally suboptimal types $L_1$, $L_2$ and $L_3$ to be ruled out because they fail to enable their users to communicate effectively. However, this is not true of $L_4$.\footnote{$L_6$ presents a special case. In our current setup, it mirrors $L_5$ in enabling for the pragmatic narrowing of message $m_2$ rather than $m_1$. However, this association of $s_2$ with $m_2$ under favorable parameter conditions is achieved by ruling out the ``some but not all''-state $s_1$ and not, as with scalar implicatures, the ``all''-state $s_2$. $L_6$ speakers therefore strengthen a ``some''-message to convey something paraphrasable as `some but not [some but not all]'. Our representation of lexica as Boolean matrices is blind to this anomaly.} Notwithstanding, the prediction that natural language communities are homogeneous or that a single speaker may entertain $L_4$-like semantics for one scalar expression and $L_5$-like semantics for another is is not implausible (\hl{references on individual level differences in the derivation of scalar implicatures}). Alternatively, a stronger tendency for posterior maximization has to be assumed (cf. Figure \ref{fig:prior-posterior}). This empirical issue relates to other two aspects left undiscussed in our present analysis: disadvantages of pragmatic reasoning and the effect of state frequencies on the fossilization of pragmatic inferences. We tacitly assumed pragmatic reasoning to come at no cost. However, there is experimental evidence that suggests  that the pragmatic derivation of upper-bounds costs effort and takes additional processing time (cf. \citealt{deNeys+schaeken:2007, huang+snedeker:2009}). This raises the question at which point such usage-based cost undercuts the learnability advantage of simpler semantic representations. Should cost play a role, then its effect is bound to depend on the frequency with which a given scalar expression is used. We conject that frequently drawn scalar implicatures might fossilize to avoid cost, while infrequent ones could still be computed on-line. This opens a possible venue to address the question about the expected presence of $L_4$-like semantics.

More broadly, we have shown that environmental factors such as noisy perception by language learners and users can lead to outcomes indistinguishible from those generated by cognitive biases. We hope this finding to invite further investigation on causes and interactions of such transmission perturbations. \hl{elaborate}

\section{General discussion}
We laid out a model that combines game-theoretical models of functional pressure towards efficient communication \citep{nowak+krakauer:1999}, effects of transmission biases on (iterated) language learning \citep{griffiths+kalish:2007}, probabilistic speaker and listener types of varied degrees of pragmatic sophistication \citep{frank+goodman:2012, franke+jaeger:2014} as well as different lexica \citep{bergen+etal:2012,bergen+etal:2016}. This model generates predictions about lexicalization patterns and, more generally, effects of communicative pressures on language. We argued that the puzzle raised by semantics in light of pragmatics is hard to explain on purely functional grounds and that part of the answer may instead lie in learnability. In particular, in perturbations in the transmission of language. In the realm of learning biases, simpler semantic representations are more likely to be learned while pragmatic reasoning can counteract functional disadvantages otherwise incurred. This result is of particular relevance for the longstanding assumption of a divide and interaction between semantics and pragmatics by offering an account of why (certain) pragmatic inferences fail to lexicalize. More broadly, we showed that systematic noise in perception can produce outcomes that are indistinguishible from those generated by cognitive biases.

The main innovations of the model are its modular separation of expressivity and learnability, allowing for their isolated and combined analysis, the learning process involving a joint inference over types of pragmatic behavior and lexical meaning, as well as in its accommodation of a range of transmission perturbations that go beyond learning biases. The goal to decouple but model both expressivity and learnability has also recently been addressed by \citet{kirby+etal:2015}. In contrast to our proposal, Kirby et al. model expressivity as exerting its force only in the production of learning data. This model's expressivity parameter thereby fulfills a similar role  to high values of $\lambda$ in making speaker behavior more deterministic. In this way, it ``favors'' unambiguous languages. Crucially, the degree of mutual understanding of interlocutors central to replication and to our notion of expressivity is not taken into consideration. That is, while our proposal combines bidirectional horizontal transmission with its vertical and unidirectional counterpart, Kirby et al.'s model only considers the latter's influence. Our reasoning behind the inclusion of the former lies in the empirical and theoretical observation that learnability alone can lead the selection of functionally defective languages, as showcased by the tautological language $L_3$ in our analysis. This outcome has been reported in a number of laboratory experiments where the participants' task was to learn and subsequently reproduce the language produced by a previous participant, leading to a proliferation of languages that associated a large number of meanings with a single form (see e.g. \citealt{silvey+etal:2014} and experiment 1 in \citealt{kirby+etal:2008}). In contrast, experiments involving an interactive component have been found to foster languages that enable interlocutors to distinguish meanings accurately  (e.g. \citealt{fay+etal:2013}; for a review of laboratory results under the iterated learning paradigm and further discussion see \citealt{kirby+etal:2015, tamariz+kirby:2016}). It is not evident how to compare these findings given that they consider distinct meaning spaces, modes of transmission, iterations and feedback given to participants. However, we take these results to suggest that there is an important difference between a language generating learnable linguistic data and its actual performance as a means of information transfer. The former solely depends on the mechanism by which speakers associate form and meaning. The latter additionally hinges on the addressee's linguistic experience and her ability to interpret linguistic input based on this experience. In sum, we contend that successful information transfer in a linguistic community is central to the adoption of a communication system and that this measure is not adequately reflected by production alone.

\hl{One of our central  assumptions is the learning bias for simpler lexical representations. Albeit implicit in our formalization, the conceptual foundation for a compression preference at the lexical level is grounded in work in the language of thought tradition put forward by \citet{piantadosi+etal:2012a,piantadosi+etal:underreview}. Following Piantadosi and colleagues we assumed that lexical meanings are expressed in a representational language and that the learners' task is to infer lexica that best explain the data witnessed. Furthermore, we assumed learners to not only infer lexical meaning but types of linguistic use. Whether language use is culturally transmitted is an empirical matter that can not be settled here. However, this idealized model suggests that semantic structure evinced cross-linguistically that lends itself for pragmatic exploitation reflects lexical information that is both easy to learn and, perhaps more interestingly, also easy to learn how to exploit via mutual reasoning.}
\hl{Shift this a bit, and involve the noisy transmission}

\section{Conclusion}
\hl{add noise}
Language change is affected by intertwined pressures. We investigated the selection of lexical meanings lacking upper-bounds in a case study on role of semantics and pragmatics in satisfying such pressures. We showed that, when pressured for learnability and expressivity, the former drives for simpler semantic representations inasmuch as pragmatics can compensate for their lack of expressivity in use. That is, the relative learnability advantage of simpler semantics may offer an answer to why natural languages fail to lexicalize certain pragmatic inferences. More broadly, we argued that semantic patterns can be explained by taking into consideration the way in which they are used in interaction as well as their representational complexity and learnability.

\bibliographystyle{plainnat}
\bibliography{./bounds-rmd}

\newpage

\section{Some thoughts + what may be missing}
\begin{itemize}
  \item I still need to review, read and find some references I'm withholding in the current draft (marked as ``\hl{references}'' in color). 
  \item The data used in plots and slides should be generated as uniformly as possible. I will do this once we settle for everything else. Any input on this is welcome, but I think it boils down to adjusting the non-noisy simulations a bit to the noisy ones (the number of samples is what mostly diverges). It's also up to debate whether we should run the non-noisy simulations with multiple Qs, as their noisy counterparts. 
  \item The analysis may benefit from further data and less of a qualitative overview of the parameters' effect (e.g. the regression analysis Michael conducted);
  \item In principle, we could make a direct comparison between the case-study of \citet{griffiths+kalish:2007} on compositionality using iterated learning and iterated learning + functional pressure. This could be relegated to the appendix. However, I'm inclined to think this is somewhat tangential and wouldn't add much. The same goes for an explicit comparison of {\em some} vs. {\em some but not all} using a language of thought-style representation. Let me know if you see worth in either option.
\end{itemize}

\end{document}