### 2016-12-29 paper/new_simulation_setup.tex
new simulation specs initial

\documentclass[a4paper]{article}

\usepackage{geometry}
\usepackage{natbib}
\bibpunct[:]{(}{)}{,}{a}{}{;}

%--------------------
%\usepackage{gb4e}
%\noautomath

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{mathrsfs}
\usepackage{nicefrac}
%\usepackage{stmaryrd}
%\usepackage{multicol}
\usepackage{graphicx}
\usepackage{color}
\usepackage{booktabs}
%\newcommand{\mvalueof}[1]{\llbracket#1\rrbracket}
\newcommand{\citeposs}[2][]{\citeauthor{#2}'s (\citeyear[#1]{#2})}
\newcommand{\tuple}[1]{\ensuremath{\left\langle #1 \right\rangle}}

\newcommand{\hl}[1]{\textcolor[rgb]{.8,.33,.0}{#1}}% prints in orange
%\newcommand{\argmax}[1]{\underset{#1}{\operatorname{arg}\,\operatorname{max}}\;}
%\newcommand{\argmin}[1]{\underset{#1}{\operatorname{arg}\,\operatorname{min}}\;}
%\newcommand{\sbna}{\exists\lnot\forall}

\definecolor{Red}{RGB}{178,34,34}
\newcommand{\mf}[1]{\textcolor{Red}{[MF: #1]}} 
\newcommand{\tb}[1]{\textcolor[rgb]{.8,.33,.0}{[TB: #1]}}% prints in orange

\usepackage{blkarray}
\usepackage{xspace}

%%% MF's commands
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\card}[1]{\left \lvert \, #1 \, \right\rvert}
\newcommand{\abs}[1]{\lvert #1 \rvert}
\newcommand{\States}{\ensuremath{S}\xspace}		% Set of States
\newcommand{\state}{\ensuremath{s}\xspace}		% single states
\newcommand{\mystate}[1]{\ensuremath{\state_{\text{#1}}}\xspace} %meaningful states
\newcommand{\Messgs}{\ensuremath{M}\xspace}		% Set of Messages
\newcommand{\messg}{\ensuremath{m}\xspace}		% single messages
\newcommand{\mymessg}[1]{\ensuremath{\messg_{\text{#1}}}\xspace} %meaningful messages
\newcommand{\ssome}{\mystate{\ensuremath{\exists\neg\forall}}}
\newcommand{\sall}{\mystate{\ensuremath{\forall}}}
\newcommand{\snone}{\mystate{\ensuremath{\emptyset}}}
\newcommand{\msome}{\mymessg{some}}
\newcommand{\mall}{\mymessg{all}}
\newcommand{\asome}{\myact{\ensuremath{\exists\neg\forall}}}
\newcommand{\aall}{\myact{\ensuremath{\forall}}}
\definecolor{mygray}{cmyk}{0.35,0.35,0.35,0.35}
\newcommand{\mygray}[1]{{\textcolor{mygray}{#1}}}
%%%

%--------------------
%
%\usepackage{setspace}
%\onehalfspacing
%
%-------------------

\title{Tracing the cultural evolution of meaning at the semantics-pragmatics interface}

\author{%\bf NAME1 and NAME2\\
    ( -- draft \today --- )
}

\date{}

\begin{document}

We consider three states $\States = \set{\snone, \ssome, \sall}$. This state space can be
thought of as a partition of possible worlds into cells where none, some or all of the $A$s are
$B$s, for some arbitrary fixed predicates $A$ and $B$. Eight concepts can be distinguished
based on their truth or falsity in three world states, six of which are not contradictory
(always false) or tautologous (always true). These are listed with mnemonic names in
Table~\ref{tab:concepts}.

\begin{table}
  \centering
\begin{center}
  \begin{tabular}{lccccl}
    \toprule
    intuitive name
    & \snone
    & \ssome
    & \sall
    & min.~derivation length
    & formula
    \\ \midrule
    ``all''
    & 0
    & 0
    & 1
    & $3$
    & $A \subseteq B$
    \\
    ``some but not all''
    & 0
    & 1
    & 0
    & $7$
    & $A \cap B \neq \emptyset \wedge A \neq \emptyset$
    \\    
    ``some''
    & 0
    & 1
    & 1
    & $4$
    & $A \cap B \neq \emptyset$
    \\
    ``none''
    & 1
    & 0
    & 0
    & $4$
    & $A \cap B = \emptyset$
    \\
    ``none or all''
    & 1
    & 0
    & 1
    & $8$
    & $\neg(A \cap B \neq \emptyset \wedge A \neq \emptyset)$
    \\
    ``not all''
    & 1
    & 1
    & 0
    & $4$
    & $\neg (A \subseteq B)$
    \\
    \bottomrule
  \end{tabular}
\end{center}
\caption{Available concepts and their minimal derivation length}
\label{tab:concepts}
\end{table}

Identifying three world states requires three messages in $\Messgs$. As before, we call a
lexicon $L$ a mapping $\Messgs \rightarrow C$ from messages to concepts. There are $6^3 = 343$
possible lexica, but many of these assign the same concept to more than one message. For
simplicity, attention could be restricted to only those lexica which avoid expressive
redundancy. Hardcoding such an \emph{mutual exclusivity bias}
\citep[e.g.][]{Clark2009:Lexical-Meaning}, there are 20 non-redundant
lexica.\footnote{Functional pressure towards efficient communication will evidently favor
  non-redundant lexica, so this restriction is fairly innocuous but practical.}

As before, the type of a player is a combination of a lexicon and a manner of pragmatic
language use: literal or pragmatic. With this, there are 40 types in this model.

\begin{table}
  \centering
  \begin{align*}
    C & \rightarrow C \wedge C 
    & 
    X & \rightarrow \set{A,B} \\
    C & \rightarrow\neg C 
    & 
    X & \rightarrow X \cap X \\
    C & \rightarrow X \subseteq X
    & 
    X & \rightarrow X \cup X \\
    C & \rightarrow X \neq \emptyset
    & 
    X & \rightarrow X \setminus X \\
    C & \rightarrow X = \emptyset     
    & 
    X & \rightarrow \overline{M}
  \end{align*}
  \caption{Toy grammar in a set-theoretic "language of thought"}
  \label{tab:grammar}
\end{table}

The prior probability of a type is just the prior probability of its lexicon. The prior of a
lexicon is a function of the complexity of the concepts in its image set. Lexica that use
simpler concepts are \emph{a priori} more likely. This can be motivated by assuming that
learners beam search for suitable concepts to map onto overt signals by (probabilistically)
considering simpler concepts first. There are many ways for defining complexity of a
concept. Following \citet{piantadosi+etal:underreview} and related work, we define complexity
of a concept as a function of its derivation length in a generative ``language of
thought''. For concreteness of example, consider the toy grammar of concepts in
Table~\ref{tab:grammar}. This grammar uses basic set-theoretic operations to form expressions
which can be evaluated as true or false in our three world states. Table~\ref{tab:concepts}
lists the minimal length of a derivation of an expression that represents each concept. A
simple way of forming priors over a lexicon is:
\begin{align*}
  P(L)  & = \prod_{c \in Im(L)} P(c)  \ \ \ \text{, with} & 
  P(c) & \propto \max_{c'}Compl(c') - Compl(c')\,,
\end{align*}
where $Compl(c)$ is the minimum length of a derivation of concept $c$.

\bibliographystyle{plainnat}
\bibliography{./bounds-rmd}
\end{document}