### 2017-03-23 paper/paper-v3.tex
revisions up to general discussion

\documentclass[a4paper, 11pt]{article}

\usepackage{geometry}
\usepackage{natbib}
\bibpunct[:]{(}{)}{,}{a}{}{;}

%--------------------
%\usepackage{gb4e}
%\noautomath

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{mathrsfs}
\usepackage{nicefrac}
%\usepackage{stmaryrd}
%\usepackage{multicol}
\usepackage{graphicx}
\usepackage{color}
\usepackage{booktabs}
\usepackage{pgfplots}
\usepackage{subcaption}
\pgfplotsset{compat=1.3}
\usetikzlibrary{pgfplots.groupplots,decorations.markings}

\newtheoremstyle{Satz}
   {}                      %Space above
   {1em}                   %Space below
   {\normalfont}           %Body font
   {}                      %Indent amount (empty = no indent,
                           %\parindent = para indent)
   {\normalfont}           %Thm head font
   {.}                     %Punctuation after thm head
   {.8em}                  %Space after thm head: " " = normal interword
                           %space; \newline = linebreak
   {\bfseries\thmname{#1}\thmnote{ (#3)}}
                           %Thm head spec (can be left empty, meaning
                           %`normal')

\theoremstyle{Satz}
\newtheorem{example}{Example}

%\newcommand{\mvalueof}[1]{\llbracket#1\rrbracket}
\newcommand{\citeposs}[2][]{\citeauthor{#2}'s (\citeyear[#1]{#2})}
\newcommand{\tuple}[1]{\ensuremath{\left\langle #1 \right\rangle}}

\newcommand{\hl}[1]{\textcolor[rgb]{.8,.33,.0}{#1}}% prints in orange
%\newcommand{\argmax}[1]{\underset{#1}{\operatorname{arg}\,\operatorname{max}}\;}
%\newcommand{\argmin}[1]{\underset{#1}{\operatorname{arg}\,\operatorname{min}}\;}
%\newcommand{\sbna}{\exists\lnot\forall}

\definecolor{Red}{RGB}{178,34,34}
\newcommand{\mf}[1]{\textcolor{Red}{[MF: #1]}} 
\newcommand{\tb}[1]{\textcolor[rgb]{.8,.33,.0}{[TB: #1]}}% prints in orange

\usepackage{blkarray}
\usepackage{xspace}

\usepackage{tgtermes}
\renewcommand{\baselinestretch}{1.2}

%%% MF's commands
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\card}[1]{\left \lvert \, #1 \, \right\rvert}
\newcommand{\abs}[1]{\lvert #1 \rvert}
\newcommand{\States}{\ensuremath{S}\xspace}		% Set of States
\newcommand{\state}{\ensuremath{s}\xspace}		% single states
\newcommand{\mystate}[1]{\ensuremath{\state_{\text{#1}}}\xspace} %meaningful states
\newcommand{\mylang}[1]{\ensuremath{L_{\text{#1}}}\xspace} %meaningful states
\newcommand{\Messgs}{\ensuremath{M}\xspace}		% Set of Messages
\newcommand{\messg}{\ensuremath{m}\xspace}		% single messages
\newcommand{\mymessg}[1]{\ensuremath{\messg_{\text{#1}}}\xspace} %meaningful messages
\newcommand{\ssome}{\mystate{\ensuremath{\exists\neg\forall}}}
\newcommand{\sall}{\mystate{\ensuremath{\forall}}}
\newcommand{\snone}{\mystate{\ensuremath{\emptyset}}}
\newcommand{\msome}{\mymessg{some}}
\newcommand{\mall}{\mymessg{all}}
\newcommand{\mnone}{\mymessg{none}}
\newcommand{\Lall}{\mylang{all}}
\newcommand{\Lbound}{\mylang{bound}}
\newcommand{\Llack}{\mylang{lack}}
\newcommand{\asome}{\myact{\ensuremath{\exists\neg\forall}}}
\newcommand{\aall}{\myact{\ensuremath{\forall}}}

\definecolor{mygray}{cmyk}{0.35,0.35,0.35,0.35}
\newcommand{\mygray}[1]{{\textcolor{mygray}{#1}}}
%%%

%--------------------
%
%\usepackage{setspace}
%\onehalfspacing
%
%-------------------

\title{Co-evolution of lexical meaning \& pragmatic use}

\author{%\bf NAME1 and NAME2\\
    ( -- draft \today --- )
}

\date{}

\begin{document}

  %% for arrow head placement
  \tikzset{->-/.style={decoration={
  markings,
  mark=at position #1 with {\arrow{>}}},postaction={decorate}}}
  %%%

\maketitle

\begin{abstract}
  According to standard linguistic theory, the meaning of an utterance is the product of
  conventional semantic meaning and general pragmatic rules on language use. To investigate how
  cultural evolution of language plays out under this picture of the semantics-pragmatics
  interface, we present a game theoretic model of the competition between types of language
  users, each endowed with a selection of lexical concepts and a particular pragmatic
  disposition to act on them. Our model traces two evolutionary forces and their interaction:
  (i) fitness-based pressure towards communicative efficiency and (ii) potential learning
  biases during the transfer of linguistic knowledge. We illustrate the model based on a case
  study on scalar implicatures. In this case study learning biases that favor simple semantic
  representations can boost the evolution of more sophisticated pragmatic reasoning types and
  so prevent the lexicalization of scalar implicatures.
\end{abstract}

\section{Introduction}\label{sec:introduction}
What is conveyed usually goes beyond what is said. A request for a blanket can be politely
veiled by uttering ``I'm cold;'' temporal succession of events can be communicated by the order
in which conjuncts appear as in ``I traveled to Paris and got married;'' an invitation can be
declined by saying ``I have to work.'' An influential explanation of the relation between the
literal meaning of expressions and what they may convey in context is due to
\citet{grice:1975}, who characterizes pragmatic use and interpretation as a process of mutual
reasoning about rational language use. For instance, under the assumption that the speaker is
cooperative and relevant, ``I have to work'' may be interpreted as providing a reason why the
speaker will not be able to accept an invitation, going beyond its literal meaning. Some of
these enrichments are rather \emph{ad hoc}. Others show striking regularities, such as the use
of ability questions for polite requests (``Could you please \dots?''), or certain enrichments
of lexical meanings such as \emph{and} to convey \emph{and then}.

A particularly productive and well studied class of systematic pragmatic enrichments are scalar
implicatures
\citep{horn:1984,Hirschberg1985:A-Theory-of-Sca,LevinsonPragmatics1983,Geurts2010:Quantity-Implic}. Usually,
the utterance of a sentence like ``I own some of Johnny Cash's albums'' will be taken to mean
that the speaker does not own all of them. This is because, if the speaker owned them all, she
could have used the word \emph{all} instead of \emph{some} in his utterance, thereby making a
more informative statement. Scalar implicatures, especially the inference from \emph{some} to
\emph{some but not all}, have been studied extensively, both theoretically
\citep[e.g.,][]{Sauerland2004:Scalar-Implicat,ChierchiaFox2008:The-Grammatical,Rooyvan-RooijJagerde-Jager2012:Explaining-Quan}
as well as experimentally
\citep[e.g.,][]{BottNoveck2004:Some-Utterances,huang+snedeker:2009,GrodnerKlein2010:Some-and-Possib,GoodmanStuhlmuller2013:Knowledge-and-I,DegenTanenhaus2012:Processing-Scal}. While
there is much dispute in this domain about many details, a position endorsed by a clear
majority is that a scalar item like \emph{some} is underspecified to mean \emph{some and maybe
  all} and that the enrichment to \emph{some but not all} is part of some regular process with
roots in pragmatics.

If this majority view is correct, the question arises how such a division of labor between
semantics and pragmatics could have evolved and why it would be so pervasive across natural
languages. Models of language evolution abound. There are simulation-based models studying the
evolution of language in populations of communicating agents
\citep{Hurford1989:Biological-Evol,Steels1995:A-Self-Organizi,LenaertsJansen2005:The-Evolutionar,SteelsBelpaeme2005:Coordinating-Pe,BaronchelliPuglisi2008:Cultural-route-,steels:2011,SpikeStadler2016:Minimal-Require}
and there are mathematical models of language evolution, mostly coming from game theory
\citep{lewis:1969,Warneryd1993:Cheap-Talk-Coor,BlumeKim1993:Evolutionary-St,nowak+krakauer:1999,Huttegger2007:Evolution-and-t,Skyrms2010:Signals}. Much
of this work has focused on explaining basic properties such as compositionality and
combinatoriality
\citep[e.g.,][]{Batali1998:Computational-S,nowak+krakauer:1999,nowak+etal:2000,KirbyHurford2002:The-Emergence-o,kirby:2002,SmithKirby2003:Iterated-Learni,Gong2007:Language-Evolut,kirby+etal:2015,verhoef+etal:2014,Franke2015:Proto-Syntax},
but little attention has been paid to the interaction between conventional meaning and
pragmatic use. What is more, many mathematical models explain evolved meaning as a regularity
in the overt behavior of agents. In contrast, we will here look at language users with a richer
cognitive make-up.

We spell out a model of the co-evolution of conventional meaning and pragmatic reasoning
types. The objects of replication and selection are pairs of lexical meanings and general types
of pragmatic behavior, which we represent using probabilistic models of pragmatic language use
\citep{frank+goodman:2012,FrankeJager2015:Probabilistic-p,GoodmanFrank2016:Pragmatic-Langu}. Replication
and selection are described by the \emph{replicator mutator dynamic}, a general and established
model of evolutionary change in large and homogeneous populations
\citep{Hofbauer1985:The-Selection-M,nowak+etal:2000,NowakKomarova2001:Evolution-of-Un,hofbauer+sigmund:2003,Nowak2006:Evolutionary-Dy}. The
approach allows us to study the interaction between (i) evolutionary pressure towards
communicative efficiency and (ii) possible infidelity in the transmission of linguistic
knowledge, caused by factors such as inductive learning biases and sparse learning
data. Considering transmission of linguistic knowledge is important because neither semantic
meanings nor pragmatic usage patterns are directly observable. Instead, language learners have
to infer these unobservables from the observable behavior in which they result. We formalize
this process as a form of Bayesian inference. Our approach thereby contains a well-understood
model of iterated Bayesian learning \citep{griffiths+kalish:2007} as a special case, but
combines it with functional selection, here formalized as the most versatile dynamic from
evolutionary game theory; the replicator dynamic
\citep{TaylorJonker1978:Evolutionary-St}. Section~\ref{sec:model} introduces this model.

Section~\ref{sec:si-case-study} applies this model to a case study on scalar implicatures. We
discuss a setting in which the majority view of underspecified lexical meanings and pragmatic
enrichments emerges if selection and transmission infidelity are combined. In particular, we
show that inductive learning biases of Bayesian learners that favor simpler lexical meanings
can prevent the lexicalization of scalar inferences and lead to the emergence of Gricean-like
pragmatic reasoning types. Results of this case study are critically assessed in the light of
the assumptions that feed our model in Section~\ref{sec:discussion}.

% We see the main contribution of this work as conceptual and technical, not as a definite answer
% to the question why scalar implicatures emerge. It rather demonstrates how current
% probabilistic cognitive modeling of language use and evolutionary modeling can be fruitfully
% combined to study the co-evolution of semantics and pragmatics side-by-side. Reversely, the
% approach taken here may be seen as a first step towards giving an evolutionary rationale for
% empirically successful probabilistic models of language use that embrace the majority view of
% the division of labor between semantics and pragmatics. Section~\ref{sec:discussion} elaborates
% on these points.

\section{A model of co-evolving lexical concepts and pragmatic behavior}
\label{sec:model}

\subsection{Expressivity and learnability}

The emergence and change of linguistic structure is influenced by many intertwined
factors. These range from biological and socio-ecological to cultural ones \citep{benz+etal:2005b,steels:2011,tamariz+kirby:2016}. Social and ecological pressures determine communicative needs, while
biology determines the architecture that enables and constrains the means by which they can be
fulfilled. In the following, our focus lies on cultural aspects, wherein processes of
linguistic change are viewed as shaped by language use and its transmission, i.e., as a result
of a process of cultural evolution
\citep{Pagel2009:Human-Language-,ThompsonKirby2016:Culture-Shapes-}.

The idea that language is an adaptation to serve a communicative function is fundamental to
many synchronic and diachronic analyses at least since \citeposs{zipf:1949} explanation of word
frequency rankings as a result of competing hearer and speaker preferences \citep[e.g.,
in][]{martinet:1962, horn:1984,jaeger+vRooij:2007,jaeger:2007,
  piantadosi:2014,kirby+etal:2015}. If processes of selection, such as conditional imitation or
reinforcement, favor more communicatively efficient types of behavior, languages are driven
towards semantic expressivity \citep[e.g.,][]{nowak+krakauer:1999,Skyrms2010:Signals}. But
pressure towards communicative efficiency is not the only force that shapes
language. Learnability is another. Natural languages need to be learnable to survive their
faithful transmission across generations. Even small learning biases implicit in acquisition
can build up and have quite striking effects on an evolving language in a process of iterated
learning
\citep{KirbyHurford2002:The-Emergence-o,SmithKirby2003:Iterated-Learni,kirby+etal:2014}.

While natural languages are pressured for both expressivity and learnability these forces may
pull in opposite directions. Their opposition becomes particularly clear when considering the
extreme (cf. \citealt{kemp+regier:2012,kirby+etal:2015}). A language consisting of a single
form-meaning association is easy to learn but lacking in expressivity. Conversely, a language
that lexicalizes a distinct form for a large number of different meanings is highly expressive
but challenging to acquire.

\subsection{The replicator mutator dynamic}

An elegant formal approach to capture the interaction between expressivity and learnability is
the \emph{replicator mutator dynamic}
\citep{Hofbauer1985:The-Selection-M,nowak+etal:2000,NowakKomarova2001:Evolution-of-Un,hofbauer+sigmund:2003,Nowak2006:Evolutionary-Dy}. In
its simplest, discrete-time formulation, the RMD defines the frequency $x'_i$ of each type $i$
in a population at the next time step as a function of: (i) the frequency $x_i$ of each type
$i$ before the update step, (ii) the fitness $f_i$ of each type $i$ before the update, and
(iii) the probability $Q_{ji}$ that an agent who wants to imitate, adopt, or learn the type of
an agent with type $j$ ends up acquiring type $i$:
\begin{align}
  \label{eq:RMD_discrete}
  x'_i = \sum_j Q_{ji} \frac{x_jf_j}{\sum_k x_k f_k}\,.
\end{align}
The RMD consists of two components: fitness-based selection and transmission biases. This
becomes most transparent when we consider an equivalent formulation in terms of a step-wise
application of the discrete-time replicator dynamic \citep{TaylorJonker1978:Evolutionary-St} on the initial population vector $\vec{x}$
and its subsequent multiplication with a mutation matrix $Q$:
\begin{align}
  \label{eq:RMD_discrete_recast}
  x'_i & = (\text{M}(\text{RD}(\vec{x})))_i\,,
\end{align}
where
\begin{align*}
      \left ( \text{RD}(\vec{x}) \right )_i 
         = \frac{x_i f_i}{\sum_k x_k f_k}
 \ \ \ \ \text{and} \ \ \ \ 
  (\text{M}(\vec{x}))_i = (\vec{x} \cdot Q)_i = \left ( \sum_j
          x_j Q_{ji} \right)_i\,.
\end{align*}
If the transmission matrix $Q$ is trivial in the sense that $Q_{ji}=1$ whenever $j=i$, the
dynamic reduces to the replicator dynamic. The replicator dynamic is a model of fitness-based
selection in which the relative frequency of type $i$ will increase with a gradient
proportional to its average fitness in the population. This dynamic is popular and
versatile because it can be derived from many abstract processes of biological and cultural
transmission and selection \citep[for overview and several derivations
see][]{Sandholm2010:Population-Game}, including conditional imitation
\citep[e.g.,][]{Helbing1996:A-Stochastic-Be,Schlag1998:Why-Imitate-and} or reinforcement
learning \citep[e.g.,][]{BorgersSarin997:Learning-Throug,Beggs2005:On-the-Converge}. If fitness
$f_i$ is the same for all types $i$, the replicator step is the identity map
$ \left ( \text{RD}(\vec{x}) \right )_i = x_i$ and the dynamic reduces to a process of
iteration of the transmission bias encoded in $Q$. In this way, the process in
(\ref{eq:RMD_discrete}), equivalently (\ref{eq:RMD_discrete_recast}), contains a model of
iterated learning \citep{griffiths+kalish:2007}.

\begin{example}
  Consider a simple coordination game. Agents are of two types: positive and negative. If
  agents of different type play with each other, they obtain a payoff of 0. If negative meets
  negative, each receives a payoff of 1. If positive meets positive, they get a payoff of 2. A
  population state is completely characterized by the proportion $x$ of negatives. The fitness
  of negatives in population state $x$ is $f_n(x) = x$, that of positives is $f_p(x) = 2-2x$.
  The average fitness is $\Phi(x) = x f_n(x) + (1-x) f_p(x) = 3x^2 - 4x + 2$. Without mutation,
  the replicator dynamic will then update $x$ to $RD(x) = \nicefrac{x^2}{\Phi(x)}$. The update
  function $RD(x)$ of the replicator step is plotted in Figure~\ref{fig:Updates_RMS} as the
  blue line. Rest points, for which $RD(x)=x$, are at $x=0$, $x=1$ and $x= \nicefrac{2}{3}$.
  The former are attractors as nearby points converge to them. Points near $x=\nicefrac{2}{3}$
  move towards 0 or 1. This is schematically pictured in the topmost phase portrait in
  Figure~\ref{fig:Phase_RD}.

  Adding mutation changes the dynamic and its rest points. Let's assume that $Q_{ji} = .9$
  when $j=i$. The update effect of mutation alone is $M(x) = .9 x + .1 (1-x) = .8x + .1$ and is
  plotted as the linear green line in Figure~\ref{fig:Updates_RMS}. It has only one stable rest
  point at $x = 0.5$ (see Figure~\ref{fig:Phase_RD}). If we first take the replicator step and
  then the mutation step in sequence, we obtain the replicator mutator dynamic
  $RMD(x) = M(RD(x)) = \nicefrac{.9x^2 - .2x +.2}{3x^2 - 4x + 2}$, which is plotted in red in
  Figure~\ref{fig:Updates_RMS}. The rest points are at $x=.121$, $x=.903$ and $x=.609$. The
  former two are attractors (see Figure~\ref{fig:Phase_RD}).
\end{example}

\begin{figure}[t]
  \centering

  %% notice that the figure is hacked, because somehow the curve does not come out right under
  %% PGFplots; I don't know why!?!?

    \begin{subfigure}[b]{0.45\textwidth}
      \centering
      \begin{tikzpicture}
      \begin{axis}[xlabel=$x$, ylabel = $x'$,
                   grid=major, width=6.85cm, height = 5.5cm,
                   legend style = {legend pos = north west},
                   ymin=0,ymax=1, xmin=0,xmax=1]

      \addplot[smooth, color=blue] {x^2 / (3*x^2 - 4*x + 2) + 0.019 * x};
      \addplot[smooth, color=green] {.8*x + .1};

      \addplot[smooth, color=red] {(.9*x^2 - .2*x +.2)/(3*x^2 - 4*x + 2)};

      \addplot[smooth, color=gray] {x};

      \legend{RD, M, RMD}

      \end{axis}
    \end{tikzpicture}

        \caption{Update functions: the population state $x$ is mapped onto $x'$ in one update step}
        \label{fig:Updates_RMS}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.5\textwidth}    
      \centering
    \begin{tikzpicture}[x=120]
      \node[draw=black, circle, fill = black, minimum size = 0.25cm]
      (0) at (0,0) {};

      \node[draw=black, circle, fill = black, minimum size = 0.25cm]
      (1) at (1,0) {};

      \draw[-, thick] (0) -- (1);

      \node[draw=black, circle, fill = white, minimum size = 0.25cm, thick]
      (mid) at (2/3,0) {};

      \node[]  (0.label) at (0,-0.5) {0};

      \node[]  (1.label) at (1,-0.5) {1};

      \node[]  (label) at (0.5,-0.5) {$x$};

      \node[]  (label) at (0.5,0.55) {RD};

      \draw[->-=0.8,very thick] (mid) -- (0);

      \draw[->-=0.8,very thick] (mid) -- (1);

    \end{tikzpicture}

    \vspace*{0.35cm}

    \begin{tikzpicture}[x=120]

      \node (0) at (-0.05,0) {};

      \node (1) at (1.05,0) {};

      \draw[-, thick] (0) -- (1);

      \node[draw=black, circle, fill = black, minimum size = 0.25cm, thick]
      (mid) at (0.5,0) {};

      \node[]  (0.label) at (0,-0.3) {0};

      \node[]  (1.label) at (1,-0.3) {1};

      \node[]  (label) at (0.5,-0.5) {$x$};

      \node[]  (label) at (0.5,0.55) {M};

      \draw[->-=0.5,very thick] (0) -- (mid);

      \draw[->-=0.5,very thick] (1) -- (mid);

    \end{tikzpicture}

    \vspace*{0.35cm}

    \begin{tikzpicture}[x=120]

      \node (0) at (-0.05,0) {};

      \node (1) at (1.05,0) {};

      \node[draw=black, circle, fill = black, minimum size = 0.25cm]
      (S0) at (0.121,0) {};

      \node[draw=black, circle, fill = black, minimum size = 0.25cm]
      (S1) at (0.903,0) {};

      \draw[-, thick] (0) -- (1);

      \node[draw=black, circle, fill = white, minimum size = 0.25cm, thick]
      (mid) at (0.609,0) {};

      \node[]  (0.label) at (0,-0.3) {0};

      \node[]  (1.label) at (1,-0.3) {1};

      \node[]  (label) at (0.5,-0.5) {$x$};

      \node[]  (label) at (0.5,0.55) {RMD};

      \draw[->-=0.5,very thick] (mid) -- (S0);

      \draw[->-=0.5,very thick] (mid) -- (S1);

      \draw[->-=0.5,very thick] (0) -- (S0);

      \draw[->-=0.5,very thick] (1) -- (S1);

    \end{tikzpicture}

    \caption{Phase portraits for RD, M and RMD: unstable rest points are hollow, attractors are
      solid.}
        \label{fig:Phase_RD}
    \end{subfigure}

  \caption{Example}
  \label{fig:Example_RMD}
\end{figure}

\subsection{Fitness \& learnability of lexical meanings \& pragmatic strategies}
\label{sec:fitn--learn}

Moving beyond an abstract example, our goal is to apply the RMD to investigate the co-evolution
of lexical concepts and pragmatic behavior. To do so, we need to fix three things: (i) what the
relevant types are, (ii) how fitness derives from communicative success and (iii) how the
mutation matrix is computed. These issues are addressed, one by one, in the following.

\subsubsection{Types: Lexica and pragmatic strategies}
\label{sec:languages+use}

Types are what evolution operates on. They define an agent's fitness, usually through a payoff
accrued in single interactions with other agents. Often types can be identified as the possible
acts in a game, e.g., either cooperating or defecting in a prisoner's dilemma. In other cases,
they may be thought of as general properties of an agent that influences her fitness, such as
being positive or negative in our previous example (whatever that means). For our present
purposes, types are identified more concretely by specific assumptions about their cognitive
make-up. Since we are interested in the evolutionary competition between different lexical
concepts and ways of using them in communication, a type is here defined as a pair consisting
of a lexicon and a pragmatic strategy.

Agents play signaling games, in which the speaker wants to communicate a world state $s$ with a
message $m$ to a hearer who receives $m$ but does not know $s$
\citep[e.g.][]{lewis:1969,Skyrms2010:Signals}. A lexicon associates each message with a
(possibly fuzzy) set of states. A pragmatic type maps a lexicon onto a probabilistic speaker
rule (a probabilistic choice of message for each state) and a probabilistic listener rule (a
probabilistic choice of state for each message). There are many ways of making these general
notions more concrete. Here is what we will assume in the remainder of this paper.

Lexica codify the truth-conditions of expressions. A convenient way to represent lexica is by
$(\card{\States}, \card{\Messgs})$-Boolean matrices, where $\States$ is a set of states
(meanings) and $M$ a set of messages (forms available in the language). For example, suppose
that there are two relevant world states $\States = \set{\ssome, \sall}$. In state $\ssome$
Chris owns some but not all of Johnny Cash's albums while in $\sall$ Chris owns them
all. Suppose that there are two messages $\Messgs = \set{\msome, \mall}$ where $\msome$ is
short for a sentence like \emph{Chris owns some of Johnny Cash's albums} and $\mall$ for the
same sentence with \emph{some} replaced by \emph{all}.  Lexica for this case would assign a
Boolean truth value, either $0$ for false or $1$ for true, to each state-message pair. The
following two lexica are minimal examples for the distinction between a lexicalized upper-bound
for \emph{some} in $\Lbound$ and the widely assumed logical semantics with only a lower-bound
in $\Llack$.

\begin{align*}
  \Lbound & = \begin{blockarray}{lcc}
    & \mygray{\msome} & \mygray{\mall} \\
    \begin{block}{l[cc]}
      \mygray{\ssome} & 1 & 0 \\
      \mygray{\sall}  & 0 & 1 \\
    \end{block}
  \end{blockarray} &
  \Llack & = \begin{blockarray}{lcc}
    & \mygray{\msome} & \mygray{\mall} \\
    \begin{block}{l[cc]}
      \mygray{\ssome} & 1 & 0 \\
      \mygray{\sall}  & 1 & 1 \\
    \end{block}
  \end{blockarray}
\end{align*}

Pragmatic types define dispositions of produce and interpret messages given a lexicon. We
distinguish between two kinds of pragmatic types. {\em Literal interlocutors} produce and
interpret messages literally, being guided only by their lexica. {\em Pragmatic interlocutors}
instead engage in mutual reasoning to inform their choices. Recent probabilistic models of
rational language use
\citep{frank+goodman:2012,FrankeJager2015:Probabilistic-p,GoodmanFrank2016:Pragmatic-Langu}
capture different types of pragmatic behavior in a reasoning hierarchy. The hierarchy's bottom,
level $0$, corresponds to literal language use, as in Equations~(\ref{h:level0}) and
(\ref{s:level0}). Pragmatic language users of level $n + 1$ act (approximately) rational with
respect to level-$n$ behavior of their interlocutors, as in Equations~(\ref{h:leveln}) and
(\ref{s:leveln}).
\begin{alignat}{3}
&H_{0}(s \mid m;L) &&\propto pr(s) L_{sm} \label{h:level0}\\
&S_{0}(m \mid s;L) &&\propto \exp(\lambda \; L_{sm}) \label{s:level0}\\
&H_{n+1}(s|m;L) &&\propto pr(s) S_{n}(m|s;L) \label{h:leveln}\\
&S_{n+1}(m|s;L) &&\propto  \exp(\lambda \; H_{n}(s|m;L)) \label{s:leveln}
\end{alignat}

According to (\ref{h:level0}), a literal hearer's interpretation of a message depends on her
lexicon and her prior over states, $pr \in \Delta(S)$, which is here assumed flat, for
simplicity. Literal interpreters thereby choose an arbitrary true interpretation for each
message. Pragmatic hearers, defined in (\ref{h:leveln}), use Bayes rule to weigh
interpretations based on a conjecture about speaker behavior. Speaker behavior is regulated by
a soft-max parameter $\lambda$, $\lambda \geq 0$ \citep{luce:1959,sutton+barto:1998}. As
$\lambda$ increases, choices approximate strict maximization of expected utilities. Expected
utility of a message $\messg$ in state $\state$ for a level $n+1$ speaker is here defined as
$H_{n}(s|m;L)$, the probability that the hearer will assign to or choose the correct
meaning. For literal speakers, utility only tracks truthfulness. Literal speakers choose any
true message with equal probability but may send false messages as well with a probability
dependent on $\lambda$.

Here are some concrete examples. \mf{please double-check the numbers!} A literal interpreter with lexicon $\Lbound$ assigns $\ssome$
a probability of $H_0(\ssome \mid \msome; \Lbound) = 1$ after hearing $\msome$, while a literal
interpreter with $\Llack$ has $H_0(\ssome \mid \msome; \Llack) = 0.5$:
\begin{align*}
  H_0(\cdot \mid \cdot, \Lbound) & = \begin{blockarray}{lcc}
    & \mygray{\ssome} & \mygray{\sall} \\
    \begin{block}{l[cc]}
      \mygray{\msome} & 1 & 0 \\
      \mygray{\mall}  & 0 & 1 \\
    \end{block}
  \end{blockarray} &
  H_0(\cdot \mid \cdot, \Llack) & = \begin{blockarray}{lcc}
    & \mygray{\ssome} & \mygray{\sall} \\
    \begin{block}{l[cc]}
      \mygray{\msome} & .5 & .5 \\
      \mygray{\mall}  & 0 & 1 \\
    \end{block}
  \end{blockarray} 
\end{align*}
If $\lambda = 1$ literal speaker behavior looks like this:
\begin{align*}
  S_0(\cdot \mid \cdot, \Lbound) & \approx \begin{blockarray}{lcc}
    & \mygray{\msome} & \mygray{\mall} \\
    \begin{block}{l[cc]}
      \mygray{\ssome} & .73 & .27 \\
      \mygray{\sall}  & .27 & .73 \\
    \end{block}
  \end{blockarray} &
  S_0(\cdot \mid \cdot, \Llack) & \approx \begin{blockarray}{lcc}
    & \mygray{\msome} & \mygray{\mall} \\
    \begin{block}{l[cc]}
      \mygray{\ssome} & .73 & .27 \\
      \mygray{\sall}  & .5 & .5 \\
    \end{block}
  \end{blockarray} 
\end{align*}
If $\lambda = 20$, literal speakers will choose with less slack:
\begin{align*}
  S_0(\cdot \mid \cdot, \Lbound) & \approx \begin{blockarray}{lcc}
    & \mygray{\msome} & \mygray{\mall} \\
    \begin{block}{l[cc]}
      \mygray{\ssome} & 1 & 0 \\
      \mygray{\sall}  & 0 & 1 \\
    \end{block}
  \end{blockarray} &
  S_0(\cdot \mid \cdot, \Llack) & \approx \begin{blockarray}{lcc}
    & \mygray{\msome} & \mygray{\mall} \\
    \begin{block}{l[cc]}
      \mygray{\ssome} &  1 & 0 \\
      \mygray{\sall}  & .5 & .5 \\
    \end{block}
  \end{blockarray} 
\end{align*}
Pragmatic hearers of level 1 have the following probabilistic interpretation behavior for
$\lambda = 1$:
\begin{align*}
  H_1(\cdot \mid \cdot, \Lbound) & \approx \begin{blockarray}{lcc}
    & \mygray{\ssome} & \mygray{\sall} \\
    \begin{block}{l[cc]}
      \mygray{\msome} & .73 & .27 \\
      \mygray{\mall}  & .27 & .73 \\
    \end{block}
  \end{blockarray} &
  H_1(\cdot \mid \cdot, \Llack) & \approx \begin{blockarray}{lcc}
    & \mygray{\ssome} & \mygray{\sall} \\
    \begin{block}{l[cc]}
      \mygray{\msome} & .59 & .41 \\
      \mygray{\mall}  & 0.35 & 0.65 \\
    \end{block}
  \end{blockarray} 
\end{align*}
For $\lambda = 20$, there wil less slack again:
Pragmatic listeners of level 1 have the following probabilistic interpretation behavior for
$\lambda = 1$:
\begin{align*}
  H_1(\cdot \mid \cdot, \Lbound) & \approx \begin{blockarray}{lcc}
    & \mygray{\ssome} & \mygray{\sall} \\
    \begin{block}{l[cc]}
      \mygray{\msome} & 1 & 0 \\
      \mygray{\mall}  & 0 & 1 \\
    \end{block}
  \end{blockarray} &
  H_1(\cdot \mid \cdot, \Llack) & \approx \begin{blockarray}{lcc}
    & \mygray{\ssome} & \mygray{\sall} \\
    \begin{block}{l[cc]}
      \mygray{\msome} & 0.67 & 0.33 \\
      \mygray{\mall}  & 0 & 1 \\
    \end{block}
  \end{blockarray} 
\end{align*}
With $\lambda = 1$, pragmatic speakers of level 1 will have:
\begin{align*}
  S_1(\cdot \mid \cdot, \Lbound) & \approx \begin{blockarray}{lcc}
    & \mygray{\msome} & \mygray{\mall} \\
    \begin{block}{l[cc]}
      \mygray{\ssome} & .73 & .37 \\
      \mygray{\sall}  & .37 & .73 \\
    \end{block}
  \end{blockarray} &
  S_1(\cdot \mid \cdot, \Llack) & \approx \begin{blockarray}{lcc}
    & \mygray{\msome} & \mygray{\mall} \\
    \begin{block}{l[cc]}
      \mygray{\ssome} & .62 & .38 \\
      \mygray{\sall}  & .38 & .62 \\
    \end{block}
  \end{blockarray} 
\end{align*}
For high $\lambda = 20$, pragmatic speakers of level 1 give us:
\begin{align*}
  S_1(\cdot \mid \cdot, \Lbound) & \approx \begin{blockarray}{lcc}
    & \mygray{\msome} & \mygray{\mall} \\
    \begin{block}{l[cc]}
      \mygray{\ssome} & 1 & 0 \\
      \mygray{\sall}  & 0 & 1 \\
    \end{block}
  \end{blockarray} &
  S_1(\cdot \mid \cdot, \Llack) & \approx \begin{blockarray}{lcc}
    & \mygray{\msome} & \mygray{\mall} \\
    \begin{block}{l[cc]}
      \mygray{\ssome} & 1 & 0 \\
      \mygray{\sall}  & 0 & 1 \\
    \end{block}
  \end{blockarray} 
\end{align*}

It is important to note that pragmatic agents of level 1 using $\Llack$ associate $\msome$
preferentially with $\ssome$, in contrast to their literal counterparts of level 0. This
association is not perfect, and usually less strong than what agents with a lexicalized upper
bound in $\Lbound$ can achieve, even without pragmatic reasoning. Higher oder iteration of
reasoning types leads to stronger associations of $\msome$ and $\ssome$ also for the
hearer. Still, the case study presented in Section~\ref{sec:si-case-study} will consider
speaker and hearer behavior at levels 0 and 1, as the latter are the simplest pragmatic
reasoning types which show a tendency to communicatively attuned pragmatic enrichment. Using
only level 1 reasoning types is therefore a conservative choice that works against the
fitness-based selection of pragmatic language use for a notion of fitness based on
communicative efficiency, which is introduced next.

\subsubsection{Fitness \& fitness-based selection based on expressivity}\label{sec:expressivity}

Under the replicator dynamic the proportion of type $i$ in a population will increase or
decrease as a function of its relative fitness $f_i$. In the context of language evolution,
fitness is usually associated with expressivity, i.e., the ability to successfully communicate
with other language users from the same population
\citep[e.g.,][]{nowak+krakauer:1999,nowak+etal:2000, nowak+etal:2002}. Under a biological
interpretation the assumption is that organisms have a higher chance of survival and
reproduction if they are able to share and receive useful information via communication with
peers. Under a cultural interpretation the picture is that agents themselves strive towards
communicative success and therefore occasionally adapt or revise their behavior to achieve
higher communicative success (see \citealt[\S3.3]{benz+etal:2005b} for discussion).

The replicator equation gives us the means to make the ensuing dynamic precise, without
necessarily committing to a biological or cultural interpretation. As above, the proportion of
types in a given population is codified in a vector $\vec{x}$, where $x_i$ is the proportion of
type $i$. The fitness of type $i$ is its average expected communicative success, or
\emph{expected utility} (EU), given the frequencies of types in the current population:
\begin{align*}
  f_i = \sum_j x_j \text{EU}(t_i,t_j)\,.
\end{align*}
The expected utility $\text{EU}(t_i,t_j)$ for type $i$ when communicating with type $j$ is the
average success of $i$ when talking or listening to $j$. If agents are speakers half of the
time, this yields:
\begin{align*}
  \text{EU}(t_i,t_j) = \nicefrac{1}{2} \, \text{EU}_S(t_i,t_j) + \nicefrac{1}{2} \, \text{EU}_H(t_i,t_j)\,,
\end{align*}
where $\text{EU}_S(t_i,t_j)$ and $\text{EU}_H(t_i,t_j)$ are the expected utilities for $i$ as a
speaker and as a hearer when communicating with $j$, defined as follows, where $n_i$ and $n_j$
are type $i$'s and type $j$'s pragmatic reasoning types and $L_i$ and $L_j$ are their lexica:
\begin{flalign*}
  & \text{EU}_S(t_i,t_j)  = \sum_s P(s)\sum_m S_{n_i}(m \mid s;L_i) \sum_{s'} H_{n_j}(s' \mid m;L_j)
  \delta(s,s') \,, \\
 & \text{EU}_H(t_i,t_j)  = \text{EU}_S(t_j,t_i)\,.
\end{flalign*}
As usual, $\delta(s,s') = 1$ iff $s = s'$ and $0$ otherwise.

\subsubsection{Learnability}
\label{sec:learnability}

Languages are shaped not only by functionalist forces towards greater expressivity. Another
important factor is the fidelity with which linguistic knowledge is transmitted. Among others,
linguistic production can be prone to errors, states or messages may be perceived incorrectly,
and multiple languages may be compatible with the data learners are exposed to. These sources
of uncertainty introduce variation in the transmission of linguistic knowledge from one
generation to the next. In particular, learning biases in the iterated transmission process can
influence language evolution substantially.

In biological evolution, where types are expressed genetically, transmission infidelity comes
into the picture through infrequent and mostly random genetic mutations. However, an agent's
lexicon and pragmatic reasoning behavior is likely not inherited genetically. They need to be
learned from observation. Concretely, when agents of type $j$ want to adopt or imitate the
linguistic behavior of type $i$, they observe the overt linguistic behavior of type $i$ and
need to infer the covert type that most likely produced the visible behavior. Iterated learning
is a process in which languages are learned repeatedly from the observation of linguistic
behavior of agents who have themselves acquired the language from observation and inference. In
the simplest case there is a single teacher and a single learner. After sufficient training the
learner becomes a teacher and produces behavior that serves as input for a new learner. Due to
the pressure towards learnability it exerts, iterated learning alone generally leads to simpler
and more regular languages (see \citealt{kirby+etal:2014} and \citealt{tamariz+kirby:2016} for
recent surveys).

Following \citet{griffiths+kalish:2007} we model language acquisition as a process of Bayesian
inference in which learners combine the likelihood of a type producing the witnessed learning
input with prior inductive biases. Experimental and mathematical results on iterated learning
suggest that the outcome of this process reflects learners' inductive biases
\citep[e.g.,][]{kirby+etal:2014}. In a Bayesian setting these biases can be codified in a prior
$P \in \Delta(T)$, which reflects the amount of data a learner requires to faithfully acquire
the language of the teacher \citep[cf.][450]{griffiths+kalish:2007}. The extent of the prior's
influence has been shown to heavily depend on the learning strategy assumed to underlie the
inference process. On the one hand, early simulation results suggested that weak biases could
be magnified by exposing learners to only small data samples
\citep[e.g. in][]{brighton:2002}. On the other, \citeposs{griffiths+kalish:2007} mathematical
characterization showed that iterated learning converges to the prior in the limit, i.e., that
the resulting distribution over languages corresponds to the learners' prior distribution and
is not influenced by the amount of input given to them. This difference in predictions can be
traced back to differences in the selection of hypotheses from the posterior. Griffith \&
Kalish's convergence to the prior holds for learners that sample from the posterior. More
deterministic strategies such as the adoption of the type with the highest posterior
probability, so-called {\it maximum a posterior estimation} (MAP), increase the influence of
both the prior and the data \citep{griffiths+kalish:2007,kirby+etal:2007}. In the following, we
use a parameter $l\ge1$ to modulate between posterior sampling and the MAP strategy. When
$l = 1$ learners sample from the posterior. The learners' propensity to maximize the posterior
grows as $l$ increases.

Let $D$ be the set of possible data that learners may be exposed to. This set $D$ contains all
sequences of state-message pairs of length $k$, e.g.,
$\tuple{\tuple{s_1,m_1},\dots , \tuple{s_k,m_k}}$. As $k$ increases, learners have more data to base their inference on and so tend to
recover the true types that generated a given sequence with higher probability. The mutation
matrix $Q$ of the replicator mutator dynamic in (\ref{eq:RMD_discrete}) can then be defined as
follows: $Q_{ji}$ is the probability that a learner acquires type $i$ when learning from an
agent of type $j$. The learner observes a length-$k$ sequence $d$ of state-message pairs, but
the probability $P(d \mid t_j)$ with which sequence $d = \tuple{\tuple{s_1,m_1},\dots , \tuple{s_k,m_k}}$ is observed depends on type $j$'s
behavior:
\begin{align*}
  P(d = \tuple{\tuple{s_1,m_1},\dots , \tuple{s_k,m_k}} \mid t_j) = \prod_{i = 1}^k S_{n_j}(m_i
  \mid s_i; L_{j})\,,
\end{align*}
where, as before, $n_j$ is $j$'s pragmatic reasoning type and $L_j$ is $j$'s lexicon. For a
given observation $d$, the probability of acquiring type $i$ is $F(t_i \mid d)$, so that:
\begin{flalign*}
  Q_{ji} \propto \sum_{d \in D} P(d \mid t_j) F(t_i \mid d)\,.
\end{flalign*}
The acquisition probability $F(t_i \mid
d)$ given datum $d$ is obtained by probability matching $l = 1$ or a tendency towards choosing
the most likely type $l > 1$ from the posterior distribution $P(\cdot \mid d)$ over types given
the data, which is calculated by Bayes' rule:
\begin{flalign*}
  & F(t_i \mid d) \propto P(t_i \mid d)^l \; \text{ and }\\
  & P(t_i \mid d) \propto P(t_i) P(d \mid t_i)\,.
\end{flalign*}

\subsection{Model summary}

Expressivity and learnability are central to the cultural evolution of language. These
components can be modelled, respectively, as replication based on a measure of fitness in terms
of communicative efficiency and iterated Bayesian learning. Their interaction is described by
the discrete time replicator mutator dynamic in (\ref{eq:RMD_discrete}), repeated here:
\begin{align*}
  x'_i = \sum_j Q_{ji} \frac{x_jf_j}{\sum_k x_k f_k}\,.
\end{align*}
This equation defines the frequency $x'_i$ of type $i$ at the next time step, based on its
frequency $x_i$ before the step, its fitness $f_i$, and the probability that a learner infers
$i$ when observing the behavior of a type-$j$ agent. Fitness-based selection is here thought of
not as biological (fitness as expected relative number of offspring) but cultural (fitness as
likelihood of being imitated or repeated) evolution, since the types that the dynamic operates
on are pairs consisting of a lexicon and a pragmatic use pattern. A type's expressivity depends
on its communicative efficiency within a population while its learnability depends on the
fidelity by which it is inferred by new generations of learners. The learners' task is
consequently to perform a joint inference over types of linguistic behavior and lexical
meaning.

The model has three parameters: $\lambda$ regulates the degree to which pragmatic speakers
choose messages that maximize the chance of communicative success; $k$ is the number of
observations for each language learner; $l$ regulates where the learners' inference behavior
lies on a spectrum from probability matching to choice of the most likely parent type.

\section{Case study: scalar implicatures}
\label{sec:si-case-study}

The model of the previous section formalizes the evolutionary competition between different
pairs of lexical concepts and ways of using them. This section looks at a concrete case study
on scalar implicatures. Concretely, we will engage in the following formal thought experiment:
if a population of language users could freely combine different lexica with different
pragmatic strategies, what there the conditions under which the majority view of scalar
implicatures could have evolved?

Recall that the majority view is that scalar implicatures are non-lexicalized pragmatic
enrichments. Scalar implicature triggers like \emph{some}, \emph{warm} or \emph{may} are
semantically weak expressions for which logically stronger expressions are salient, e.g.,
\emph{all}, \emph{hot} or \emph{must}. For instance, {\em some} is entailed by {\em all}. If
the sentence ``Chris owns all of Johnny Cash's albums'' is true, then ``Chris owns some of
Johnny Cash's albums'' is also true. However, while weaker expressions such as {\em some} are
truth-conditionally compatible with stronger alternatives such as {\em all}, this is not what
their use is normally taken to convey. Instead, the use of a less informative expression when a
more informative one could have been used can license a defeasible inference that stronger
alternatives do not hold (cf. \citealt{horn:1972,gazdar:1979}). That is, a hearer who assumes
the speaker to be able and willing to provide all relevant information can infer that stronger
alternatives do not hold because the speaker used a weaker alternative instead. In this way,
``Chris owns some of Johnny Cash's albums'' is strengthened to convey that she owns {\em some
  but not all} albums. The upper bound, so the majority view, is a pragmatic inference but not
part of the conventional meaning. \mf{I wonder if this paragraph is really still necessary
  after the introduction?}

In the following we consider a specific application of the model from Section~\ref{sec:model}
which allows to address the question if or when scalar inferences might (not) lexicalize. We
consider what is perhaps one of the simplest non-trivial setups that speak to this matter and
reflect on its limitations in Section~\ref{sec:discussion}. The setup is introduced in
Section~\ref{sec:setup:-stat-conc}. Section~\ref{sec:simulation-results} describes the simulations
and their results.

% Our initial question about the division of labor between semantics and pragmatics can be narrowed to the case of scalar implicatures by asking for a justification for the lack of lexical upper-bounds in weak scalar alternatives. That is, we ask why lexical meanings that lack upper-bounds and convey it pragmatically are regularly selected for over alternatives such as that of codifying the bound semantically. More poignantly, would it not serve language users better if weak(er) expressions such as {\em warm}, {\em or}, {\em some} and {\em big} were truth-conditionally incompatible with stronger alternatives such as {\em hot}, {\em and}, {\em all} and {\em huge} to avoid misunderstanding?  This question is particularly striking considering the number of expressions that license such inferences across natural languages.

\subsection{Setup}
\label{sec:setup:-stat-conc}

To fill the model from Section~\ref{sec:model} with life, we need to specify the sets of
states, messages and lexica. Additionally, we want to explore the effects of a learning bias in
favor of simple lexical concepts. One way of motivating and formalizing such a bias is
introduced below.

\subsubsection{States, messages, concepts and Lexica}

Consider a state space with three states $\States = \set{\snone, \ssome, \sall}$ and think of
it as a partition of possible worlds into cells where none, some or all of the $A$s are $B$s,
for some arbitrary fixed predicates $A$ and $B$. Eight concepts can be distinguished based on
their truth or falsity in three world states, six of which are not contradictory or
tautological.

A lexicon $L$ is a mapping $\Messgs \rightarrow C$ from messages to concepts. With three
messages there are $6^3 = 216$ possible lexica. Some assign the same concept to more than one
message and others lexicalize the same concepts but associate them with different messages. Out
of these possible lexica, three kinds are of particular relevance. First, lexica that assign
the same concept to more than one message. Such lexica lack in expressivity but may be favored
by particular learning biases nonetheless. Second, lexica that conventionalize upper-bounds to
realize a (quasi-)partition of the state space. Finally, lexically that do not lexicalize an
upper bound but allow it to be conveyed pragmatically due to the presence of a stronger lexical
item. There are six lexica of the second kind and six of the third. The following three lexica
exemplify each kind:
\begin{align*}
  \begin{blockarray}{lccc}
    & & \underline{\Lall} & \\
    & \mygray{\mnone} & \mygray{\msome} & \mygray{\mall} \\
    \begin{block}{l[ccc]}
     \mygray{\snone}  & 0 & 0 & 0\\
     \mygray{\ssome}  & 0 & 0 & 0\\
    \mygray{\sall}   & 1 & 1 & 1\\
    \end{block}
% \\ & & P(\Lall)\rlap{ $\propto 9$} & \\
  \end{blockarray} & &
 \begin{blockarray}{lccc}
    & & \underline{\Lbound} & \\
    & \mygray{\mnone} & \mygray{\msome} & \mygray{\mall} \\
    \begin{block}{l[ccc]}
       & 1 & 0 & 0\\
       & 0 & 1 & 0\\
       & 0 & 0 & 1\\
    \end{block}
%  \\& & P(\Lbound)\rlap{ $\propto 96$} & \\
  \end{blockarray} & &
  \begin{blockarray}{lccc}
    & & \underline{\Llack} & \\
    & \mygray{\mnone} & \mygray{\msome} & \mygray{\mall} \\
    \begin{block}{l[ccc]}
       & 1 & 0 & 0\\
       & 0 & 1 & 0\\
       & 0 & 1 & 1\\
    \end{block}
%   \\& & P(\Llack)\rlap{ $\propto 48$} & \\
  \end{blockarray}
\end{align*}
Lexicon $\Lall$ is clearly bad for communication: all message and interpretation choices will
be equally likely for all types; no information will be conveyed with it. Users of $\Lbound$
need not resort to pragmatic reasoning to convey an upper-bound with weak scalar
expressions. Instead, this information is already provided by the semantics of their lexicon,
giving them an advantage in terms of expressivity. Types with such lexica may be expected to
generally achieve higher communicative success than pragmatic users with a lexicon of type
$\Llack$ (see Section~\ref{sec:languages+use}).

Recall that types are a combination of a lexicon and a manner of language use. We analyze the
model's predictions in populations of types with one of the two behaviors introduced earlier;
literal or pragmatic. The former correspond to level $0$ reasoners and the latter to ones of
level $1$. Accordingly, we consider a total of $432$ types. Six are variants of pragmatic
language users with $\Llack$-like lexica. These will be our target types, because they
represent lexica and language use that conform to the majority view of scalar
implicatures. Twelve types are either literal or pragmatic speakers with lexical of the
$\Lbound$ kind. These will be the target types main competitors in evolutionary competition.

\subsubsection{An inductive learning bias for semantic simplicity}
\label{sec:an-induct-learn}

There is a growing effort to develop empirically testable representational languages that allow
for the measure of semantic complexity. For instance, so-called {\em languages of thought}
(LOT) have been put to test in various rational probabilistic models that show encouraging
results (see, e.g., \citealt{katz+etal:2008, piantadosi+etal:underreview,
  piantadosi+etal:2012a} and \citealt{piantadosi+jacobs:2016} for recent discussion). At its
core, a LOT defines a set of operations and composition rules from which concepts can be
derived. As a first approximation and for the sake of concreteness, we follow this approach to
make the preference of learners for simpler semantic representations precise and define the
complexity of a representation as a function of its derivation cost in a weighed generative
LOT.

\begin{table}
  \centering
  \begin{align*}
    C & \rightarrow_2 C \wedge C 
    & 
    X & \rightarrow_1 \set{A,B} \\
    C & \rightarrow_2 \neg C 
    & 
    X & \rightarrow_1 X \cap X \\
    C & \rightarrow_1 X \subseteq X
    & 
    X & \rightarrow_1 X \cup X \\
    C & \rightarrow_1 X \neq \emptyset \\
    C & \rightarrow_1 X = \emptyset     
  \end{align*}
  \caption{Toy grammar in a set-theoretic LOT with weighted rules.}
  \label{tab:grammar}
\end{table}

Our toy grammar of concepts is given in Table~\ref{tab:grammar}. This grammar uses basic
set-theoretic operations to form expressions which can be evaluated as true or false in states
$\snone$, $\ssome$ or $\sall$ from above. Applications of generative rules have a cost attached
to them. Here we simply assume that Boolean combinations of concepts are more complex than
atomic concepts and that otherwise each rule application adds the same cost
unit. Table~\ref{tab:concepts} lists all six concepts relevant here, their truth conditions and
the simplest formula that expresses this concept from the grammar in Table~\ref{tab:grammar}.

\begin{table}
  \centering
\begin{center}
  \begin{tabular}{lccclc}
    \toprule
    intuitive name
    & \snone
    & \ssome
    & \sall
    & least complex formula
    & complexity
    \\ \midrule
    ``all''
    & 0
    & 0
    & 1
    & $A \subseteq B$
    & $3$
    \\
    ``some but not all''
    & 0
    & 1
    & 0
    & $A \cap B \neq \emptyset \wedge A \neq \emptyset$
    & $8$
    \\    
    ``some''
    & 0
    & 1
    & 1
    & $A \cap B \neq \emptyset$
    & $4$
    \\
    ``none''
    & 1
    & 0
    & 0
    & $A \cap B = \emptyset$
    & $4$
    \\
    ``none or all''
    & 1
    & 0
    & 1
    & $\neg(A \cap B \neq \emptyset \wedge A \neq \emptyset)$
    & $10$
    \\
    ``not all''
    & 1
    & 1
    & 0
    & $\neg (A \subseteq B)$
    & $5$
    \\
    \bottomrule
  \end{tabular}
\end{center}
\caption{Available concepts and their minimal derivation length}
\label{tab:concepts}
\end{table}

The complexity measure for lexical concepts from Table~\ref{tab:concepts} is used to define a
learning bias that favor simpler concepts over more complex ones. The prior probability of a
type is just the prior probability of its lexicon. The prior of a lexicon is a function of the
complexity of the lexical representations in its image set. Lexica with simpler concepts have a
higher prior. One simple way of defining such priors over lexica (and thereby types) is:
\begin{align*}
  P(L)  & \propto \prod_{c \in Im(L)} P(c)   \ \text{, with} & 
  P(c) & \propto \max_{c'}Compl(c') - Compl(c) + 1\,,
\end{align*}
where $Compl(c)$ is the complexity of the minimal derivation length of $c$ according to the
LOT-grammar (see Table~\ref{tab:concepts}). There are many ways to define priors over lexica
(see, e.g., \citealt{goodman+etal:2008, piantadosi+etal:2012a}) but the key assumption here,
common to all of them, is that simple representational expressions should be favored over more
complex ones. We should stress that these details -- from the generative grammar to its
complexity measure -- are to be regarded as one convenient operationalization of one general
approach to explicating learning biases; this is not a commitment that this general approach is
necessarily superior or that, within it, this particular instrumentalization is the single most
plausible.

\subsection{Simulation results}
\label{sec:simulation-results}

Recall that there are three parameters: soft-max parameter $\lambda$ affects how strongly
pragmatic speakers favor more informative messages; the bottleneck size $k$ influences how
faithfully learners can identify their teacher type; $l$ defines the learners disposition
towards choosing the most likely teacher type from the posterior distribution. We expect that
competitor types (types with lexica of the kind $\Lbound$) have a fitness advantage over target
types (pragmatic agents with lexica of the kind $\Llack$), especially for very low levels of
$\lambda$. Selection based on fitness alone may therefore not lead to prevalence of target
types in the population. On the other hand, lexica of type $\Llack$ are simpler than those of
type $\Lbound$ by the postulated measure from above. This may make them more likely to be
adopted by learners, especially when $k$ is low so that teacher types are relatively
indistinguishable based on their behavior in general and when $l$ is high. Still, lexica of the
kind $\Lall$ are in turn more likely \emph{a priori} than lexica of the kind
$\Llack$. Simulation results will shed light on the question whether target types can emerge
and for which parameter constellations.

To better understand the workings a pressure towards communicative efficiency and a pressure
towards learnability, we look at the behavior of the replicator and mutator step first in
isolation, and then in combination. All simulation runs are initialized with an arbitrary
distribution over types, constituting a population's first generation. All reported results are
the outcome of 50 update steps for each different constellation of parameter values. This
ensures that these outcomes correspond to developmental plateaus in which change is, if not
absent, then at least very slow, In other words, even if the resulting states do not correspond to
an eventual attracting state, they characterize long almost stationary states in which the
system remains for a very long time. As specified in \S\ref{sec:learnability}, the mutation
matrix $Q$ can be obtained by considering all possible state-message sequences of length
$k$. Given that this is intractable for large $k$, the sets of data learners are exposed to are
approximated by sampling $250$ $k$-length sequences from each type's production probabilities.

\subsubsection{Replication only: selection based on expressivity and communicative success} 
\begin{figure}
\centering
\includegraphics[width=\textwidth,height=8cm, keepaspectratio]{./plots/fig1-onlyr}
\caption{Stacked proportion of pragmatic $\Llack$-style types and incumbent types, when other than pragmatic $\Llack$, in $5$ independent populations after $50$ generations under only a pressure for expressivity. The right-most plots zoom in on only the proportion of pragmatic $\Llack$ types.}
\label{fig:only-R}
\end{figure}

As expected, selection based on communicative success is sensitive to $\lambda$ as it
influences signaling behavior. This is showcased in Figure \ref{fig:only-R}, which depicts the
proportion of target types in $5$ independent populations after $50$ generations, as well as
that of the type with the highest proportion in a given population, the incumbent. With low
$\lambda$ many types have very similar behavior, so that evolutionary selection lacks grip and
becomes very slow. The result is a very long transition with near stagnancy in a rather
homogeneous population with many types. Conversely, higher $\lambda$ promotes more rational
linguistic behavior, widening the gap in expressivity between types and promoting more
homogeneous populations. As suggested by Figure \ref{fig:only-R}, the incumbent in most
populations is not one of the six pragmatic $\Llack$-style types. That is, a pressure only for
expressivity does not lend a justification to the systematic prevalence of a lack of
upper-bounds under any $\lambda$-value. For instance, in $1000$ independent populations with
$\lambda = 20$ pragmatic $\Llack$ types only made up $11$ cases of incumbency, corresponding to
a mean proportion of $0.003$ across populations. By contrast, $913$ incumbents were
$\Lbound$-style users with close to an even share between literal ($454$) and pragmatic types
($459$), corresponding to a mean proportion of about $.48$ taken
together. %Across types and populations incumbents make up $.50$ of the mean outcome.

\subsubsection{Iterated learning only}

\begin{figure}
\centering
\includegraphics[width=1\textwidth,height=8cm,keepaspectratio]{./plots/fig2-onlym-pr}
\caption{Stacked proportion of pragmatic $\Llack$-style types and incumbent types, when other than pragmatic $\Llack$, in $5$ independent populations after $50$ generations under only a pressure for learnability ($\lambda = 20, k = 5$). The types' prior probability is shown in the right-most plot.}
\label{fig:only-M}
\end{figure}

The effect of iterated learning without a pressure for expressivity using either posterior
sampling ($l = 1$) or a stronger tendency towards posterior maximization ($l = 15$) is shown in
Figure \ref{fig:only-M} together with the prior over types. The prior shows that while users of
$\Llack$ are not the most favored by the inductive bias (compared, e.g., to $\Lall$) they are
nevertheless more advantaged than others, such as $\Lbound$, in virtue of the relatively simple
semantics they conventionalize. Crucially, $\Llack$ enables its users to convey each state with
a single message when combined with pragmatic reasoning provided sufficiently high
$\lambda$. This makes it less likely to be confused with other types if the learning data is
not too sparse ($k \geq 5$). Put differently, learners have a higher propensity to infer
pragmatic $\Llack$ over types that might be hard or impossible to distinguish it based on overt
linguistic behavior, such as $\Lbound$. Moreover, $\Llack$ is also less likely to be confused
with types with different observable behavior because its pragmatic use approximates a
one-to-one form-meaning mapping. As a consequence, a stronger propensity to maximize the
posterior increases their proportion in the population.

However, in contrast to a pressure only for expressivity with high $\lambda$, learnability
alone does not succeed in selecting for a single prevalent type. Indeed, all six target types
tend to coexist at roughly equal proportion. Each is passed on to next generation with the same
faithfulness and, differently from a pressure for communicative success, they do not stand in
competition with each other. As implicit in Figure \ref{fig:only-M}, in $1000$ independent
populations all incumbents were pragmatic $\Llack$ users provided sufficiently high $\lambda$,
with each reaching approximately the same proportion of users in the population. As with
expressivity only, low values of $\lambda$ make the differences in observable behavior across
types less pronounced and therefore reflect the learners' inductive bias more faithfully,
favoring functionally deficient but a priori preferred types such as those that use $\Lall$. A
pressure for learnability alone may consequently lead to a spread of communicatively suboptimal
types that are easier to learn. In the extreme, when $l = 1$ and $\lambda = 1$ all of $1000$
independent populations had users of $\Lall$ as incumbents.

\subsubsection{Combining pressures of expressivity and learnability}

Pressures for communicative success or learnability are not sufficient on their own to have a
single target type dominate the population. When pressured toward expressivity, the slight
communicative advantage of $\Lbound$ users leads to the its prevalence. When pressured toward
learnability, pragmatic $\Llack$ is promoted over functionally similar but semantically more
complex alternatives such as $\Lbound$. However, on its own learnability does not foment the
propagation of these types across the population. What is more, it leads to heterogeneous
populations that foster an equal share of types of a kind, such as different pragmatic $\Llack$
types.

\begin{figure}
\centering
\includegraphics[width=1\textwidth,height=8cm,keepaspectratio]{./plots/fig3-r+m}
\caption{Stacked proportion of pragmatic $\Llack$-style types and incumbent types, when other than pragmatic $\Llack$, in $5$ independent populations after $50$ generations under both pressures ($k = 5$).}
\label{fig:rmd}
\end{figure}

Figure \ref{fig:rmd} illustrates the combined effects of both pressures for a sample of $\lambda$ and $l$ values. These results show that an inductive learning bias for simpler semantics in tandem with functional pressure can lead to the selection of $\Llack$-like semantics paired with pragmatic language use. The proportion of these types increases with $\lambda$ and $l$. Pressure towards expressivity magnifies the effects of iterated learning and dampens the proliferation of types of a kind that are equal in expressivity {\em and} learnability. A pressure towards learnability favors the transmission of simpler semantics and thereby makes types that exploit pragmatic reasoning to strengthen lexical meaning more likely to be adopted over alternatives that do not. %In short, neither learnability nor functional pressure alone but their combination may lead to the lack of upper-bounds in the lexical meaning of scalar expressions in predominantly monomorphic populations.

As before, low $\lambda$ and $l$ lead to the incumbency of communicatively suboptimal types that are a priori favored, such as $\Lall$. An increase in $\lambda$ leads to the selection and incumbency of pragmatic $\Llack$ but does not lead to monomorphic populations if learners sample from the posterior. Finally, a combination of sufficiently rational signaling behavior with some tendency for posterior maximization leads to increasingly monomorphic populations of types that convey upper-bounds pragmatically. The difference between the mean of the highest pragmatic $\Llack$-like type in $1000$ independent populations and the highest proportion of other types across $\lambda$ and $l$ values is shown in Figure \ref{fig:diff}.

\begin{figure}
\centering
\includegraphics[width=1\textwidth,height=6cm,keepaspectratio]{./plots/fig4-incumbents-difference}
\caption{Difference between mean proportion of highest pragmatic $\Llack$ type and highest other type in $1000$ independent populations after $50$ generations under both pressures ($k = 5$). \tb{I will run more simulations with higher $\lambda$ and $l$. I'm pretty sure the difference still increases.}}
\label{fig:diff}
\end{figure}

As for the effect of the sequence length $k$ not mentioned so far, it influences populations in a predictable way: small values lead to more heterogeneous populations that reflect the learner's prior more faithfully. This is due to the fact that the likelihood that a small sequence was produced by any type is relatively uniform (modulo prior). By contrast, larger values increasingly allow learners to differentiate types with different signaling behaviors.

To recapitulate, other than the involvement of pressure on both expressivity and learnability,
the resulting proportion of pragmatic $\Llack$ speakers primarily hinges on three
factors. First, the degree to which agents try to maximize communicative success. This plays a
role both in expressivity as well as in producing data that allows learners to discriminate
this type from others. Second, the inductive bias, which leads learners to prefer simpler over
more complex semantic representations in acquisition. Lastly, the posterior parameter, as it
magnifies the effects of the learning bias in tandem with replication.

\subsection{Summary}

Under the assumption of a learning bias for simpler lexical representations, our results
suggest that a lack of semantic upper-bounds coupled with pragmatic reasoning can overcome
communicative pressures and come to dominate a population. This prediction hinges on three
assumptions. First, that language is pressured toward both expressivity and
learnability. Second, that language users favor expressions/interpretations expected to have a
higher likelihood of being understood/meant over ones with a lower likelihood. Lastly, that
learners prefer simpler over more complex lexical representations and exhibit a tendency
towards the acquisition of the type that best explains the learning data.

\section{General discussion}\label{sec:discussion}

The approach introduced here combines game theoretic models of functional pressure towards
efficient communication \citep{nowak+krakauer:1999}, effects of transmission perturbations on
(iterated) language learning \citep{griffiths+kalish:2007}, probabilistic speaker and listener
types of varied degrees of pragmatic sophistication \citep{frank+goodman:2012,
  franke+jaeger:2014} as well as reasoning about unobservable lexical representations
\citep{bergen+etal:2012,bergen+etal:2016}. This model allows a structured conceptual
investigation of the co-evolution of conventional meaning and pragmatic language use. Main
innovations of the model are (i) its modular separation of expressivity and learnability,
allowing for their isolated and combined analysis, (ii) the characterization of language
learning as a joint inference over types of pragmatic behavior and lexical meaning, and (iii)
the possibility to trace the co-evolution of conventional semantics and pragmatic use.

The goal to decouple but model both expressivity and learnability has also recently been
addressed by \citet{kirby+etal:2015}. In contrast to our proposal, Kirby et al. model
expressivity as exerting its force only in the production of learning data. The degree of
mutual understanding of interlocutors central to replication and to our notion of expressivity
is absent. That is, while our proposal combines bidirectional horizontal transmission with its
vertical and unidirectional counterpart, Kirby et al. only consider the latter's influence. Our
reasoning behind the inclusion of the former lies in the empirical and theoretical observation
that learnability alone can lead the selection of functionally defective languages, as
showcased by $\Lall$. This outcome has been reported in a number of laboratory experiments
where subjects learned and subsequently reproduced the language produced by a previous
participant over multiple iterations, leading to a proliferation of ambiguous languages that
associate many meanings with a single form (see, e.g., \citealt{silvey+etal:2014} and
experiment 1 in \citealt{kirby+etal:2008}). In contrast, experiments involving an interactive
component have been found to foster languages that enable interlocutors to communicatively
distinguish meanings more accurately (e.g., \citealt{fay+etal:2013}; for a review of laboratory
results under the iterated learning paradigm and further discussion see
\citealt{kirby+etal:2015, tamariz+kirby:2016}). It is not evident how to compare these
empirical findings given that they consider distinct meaning spaces, modes of transmission,
iterations and feedback given to participants. However, on a general level they may suggest
that there is an important difference between a language generating learnable linguistic data
and its actual performance as a means of information transfer. That is, we contend that
successful information transfer in a linguistic community is central to the adoption of a
communication system and that this measure may not be adequately reflected by production alone.

The main result of our case study is that types that correspond to the majority view of scalar
implicatures (scalar readings are non-lexicalized pragmatic enrichments) can come to dominate a
population. This can happen under the assumption that simpler semantic representations are more
likely to be learned (cf. \citealt{chater+vitanyi:2003}). Pragmatic language use can then be
recruited indirectly by a preference for simpler lexical concepts. Under this view, semantics
and pragmatics play a synergic role: pragmatic use allows maintenance of simpler concepts;
pressure towards representational simplicity indirectly promotes pragmatic over literal
language use. As a consequence, iterated transmission and use of language lead to a
regularization that may explain the lack of lexicalization of systematic pragmatic
enrichments.

While the results of this case study are interesting, they also raise a number of critical
issues. First of all, it is not clear to what extent other types should be present in the final
population. It seems reasonable to expect functionally suboptimal lexica such as $\Lall$ to be
ruled out but matters are different for $\Lbound$. The prediction that natural language
communities are homogeneous or that speakers may entertain $\Llack$-like semantics for one
scalar expression and $\Lbound$-like semantics for another is not implausible
(cf. \citealt{NOVECK ETC}). Alternatively, a stronger tendency for posterior
maximization in learning has to be assumed. This empirical issue relates to other two aspects
left undiscussed: disadvantages of pragmatic reasoning and the effect of state frequencies on
the fossilization of pragmatic inferences. We tacitly assumed pragmatic reasoning to come at no
cost. However, there is experimental evidence that suggests that the pragmatic derivation of
upper-bounds costs effort and takes additional processing time
(cf. \citealt{deNeys+schaeken:2007, huang+snedeker:2009}). This raises the question at which
point such usage-based cost undercuts the learnability advantage of simpler semantic
representations. Should cost play a role, then its effect is bound to depend on the frequency
with which a given scalar expression is used. It may be that frequently drawn scalar
implicatures fossilize to avoid cost whereas infrequent ones are derived on-line to avoid
complexity in acquisition. This also opens a possible venue to address the preceding question
about the expected presence of $\Lbound$-like semantics in a population, but further empirical
evidence is needed to assess these matters.

\section{Conclusion}
The cultural evolution of meaning is influenced by intertwined pressures. We set out to investigate this process by putting forward a model that combines a pressure toward successful information transfer with perturbations that may arise in the transmission of linguistic knowledge in acquisition. Its objects of selection and replication are pairs of lexical meanings and patterns of language use. This allows the model to trace the interaction between conventional meaning and pragmatic use. Additionally, it takes the challenge serious of neither semantics nor pragmatics being directly observable. Instead, learners need to infer these unobservables from overt data that results from their combination.  These components and their mutual influence were highlighted in a case study on the lack of lexical upper-bounds in weak scalar expressions that showed that, when pressured for learnability and expressivity, the former force drives for simpler semantic representations inasmuch as pragmatics can compensate for lack of expressivity in use. That is, the relative learning advantage of simpler semantics in combination with functional pressure in use may offer an answer to why natural languages fail to lexicalize systematic pragmatic inferences. And, more broadly, to a division of labors between semantics and pragmatics.

\bibliographystyle{chicago}
\bibliography{./bounds-rmd}

\newpage

\appendix

\section{Snippets}

\subsection{On alternative explanations}
\label{sec:altern-expl}

We see two main explanations for the lack of upper-bounds in the lexical meaning of weak scalar expressions. The first is that their truth-conditional compatibility with stronger expressions endows them with a broader range of applicability by allowing them to occur in contexts in which their upper-bounded reading is absent. This can happen when embedded in downward-entailing contexts, when the speaker is likely uncertain about whether the upper bounded reading is true, or when the distinction between an upper-bounded reading and the simple, only lower-bounded reading, is not relevant. For instance, if for all the speaker knows Chris owns some of Johnny Cash's albums but she does not know whether he owns all, then the use of {\em some} lacking an upper-bound succinctly conveys her uncertainty. This may suggest a functionalist argument for why upper-bounded meanings do not conventionalize: Should contextual cues provide enough information to the hearer to identify whether a bound is intended to be conveyed pragmatically, then this is preferred over expressing it overtly through longer expressions, e.g., by saying {\em some but not all} explicitly. Importantly, although morphosyntactic disambiguation may be dispreferred due to its relative length and complexity \citep{piantadosi+etal:2012b}, it allows speakers to enforce an upper-bound and override contextual cues that might otherwise mislead the hearer. In a nutshell, this explanation posits that scalar implicatures fail to lexicalize because, all else being equal, speakers prefer to communicate as economically as possible and pragmatic reasoning enables them to do so. Compare this with a hypothetical language that lexicalizes two expressions for each weak scalar expression -- one with and one lacking an upper-bound. We see four conditions along this functionalist explanation that may pressure languages for English-like semantics over this alternative. First, contextual cues are very reliable. Second, morphosyntactic disambiguation is seldom necessary. Third, morphosyntactic disambiguation is only marginally dispreferred. Fourth, larger lexica are costly. Overall, neither condition seems convincing as a central explanatory device for such a widespread phenomenon. The first two conditions put a heavy burden on the ability to retrieve contextual cues to a degree that seems unlikely to undercut the benefit of unambiguous communication. It is likely that human language users are very good at retrieving cues from context, but to stipulate that they are so good as to undercut the benefit of safe communication provided by this hypothetical alternative strikes us as too strong of an assumption.  As for the third and fourth condition, these seem mostly like technical solutions without a proper empirical basis.

Instead, the systematicity and typological spread of scalar implicatures together with the observation that monomorphemic expressions that lexically rule out stronger alternatives are unattested across languages (\citealt[252-267]{horn:1984}, \citealt{horn:1972,traugott:2004,vdAuwera:2010}) suggests that other forces may be at play. In what follows we investigate the predictions of our model under the assumption that the lack of lexicalization of scalar inferences may be accounted for by the relative representational simplicity of lexical meanings lacking an upper-bound over those that explicitly codify it. This difference is reflected in a learning bias towards more compressed lexical representations. That is, in a preference of learners for simpler over more complex explanations of the data they witness \citep{feldman:2000, chater+vitanyi:2003, piantadosi+etal:2012a, kirby+etal:2015,piantadosi+etal:underreview}. While we do not want to argue that functional aspects as the ones discussed above do not play a role, we do see a clear benefit in exploring whether matters of transmission biases would not give us additional explanatory leverage.

\end{document}