### 2018-03-13 paper/cogsci-submission2.tex
thorough revision first half; appendix added; clarified effect of l>1

\documentclass[a4paper, 11pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage{natbib}
\bibpunct[:]{(}{)}{,}{a}{}{;}

%--------------------
%\usepackage{gb4e}
%\noautomath

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{mathrsfs}
\usepackage{nicefrac}
%\usepackage{stmaryrd}
%\usepackage{multicol}
\usepackage{graphicx}
\usepackage{color}
\usepackage{booktabs}
\usepackage{pgfplots}
\usepackage{subcaption}
\pgfplotsset{compat=1.3}
\usetikzlibrary{pgfplots.groupplots,decorations.markings}

\newtheoremstyle{Satz}
   {}                      %Space above
   {1em}                   %Space below
   {\normalfont}           %Body font
   {}                      %Indent amount (empty = no indent,
                           %\parindent = para indent)
   {\normalfont}           %Thm head font
   {.}                     %Punctuation after thm head
   {.8em}                  %Space after thm head: " " = normal interword
                           %space; \newline = linebreak
   {\bfseries\thmname{#1}\thmnote{ (#3)}}
                           %Thm head spec (can be left empty, meaning
                           %`normal')

\theoremstyle{Satz}
\newtheorem{example}{Example}

%\newcommand{\mvalueof}[1]{\llbracket#1\rrbracket}
\newcommand{\citeposs}[2][]{\citeauthor{#2}'s (\citeyear[#1]{#2})}
\newcommand{\tuple}[1]{\ensuremath{\left\langle #1 \right\rangle}}

\newcommand{\hl}[1]{\textcolor[rgb]{.8,.33,.0}{#1}}% prints in orange
%\newcommand{\argmax}[1]{\underset{#1}{\operatorname{arg}\,\operatorname{max}}\;}
%\newcommand{\argmin}[1]{\underset{#1}{\operatorname{arg}\,\operatorname{min}}\;}
%\newcommand{\sbna}{\exists\lnot\forall}

\definecolor{Red}{RGB}{178,34,34}
\definecolor{Steelblue}{RGB}{70,130,180}
\newcommand{\mf}[1]{\textcolor{Red}{[MF: #1]}} 
\newcommand{\tb}[1]{\textcolor[rgb]{.8,.33,.0}{[TB: #1]}}% prints in orange
\newcommand{\rvv}[1]{\textcolor{Steelblue}{#1}}% prints in steel blue
\definecolor{indianred}{RGB}{205,92,92}
\definecolor{mygray}{cmyk}{0.35,0.35,0.35,0.35}
\definecolor{goldish}{RGB}{118, 104, 154}

\usepackage{blkarray}
\usepackage{xspace}

\usepackage{tgtermes}
\renewcommand{\baselinestretch}{1.2}

%%% MF's commands
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\card}[1]{\left \lvert \, #1 \, \right\rvert}
\newcommand{\abs}[1]{\lvert #1 \rvert}
\newcommand{\States}{\ensuremath{S}\xspace}		% Set of States
\newcommand{\state}{\ensuremath{s}\xspace}		% single states
\newcommand{\mystate}[1]{\ensuremath{\state_{\text{#1}}}\xspace} %meaningful states
\newcommand{\mylang}[1]{\ensuremath{L_{\text{#1}}}\xspace} %meaningful states
\newcommand{\Messgs}{\ensuremath{M}\xspace}		% Set of Messages
\newcommand{\messg}{\ensuremath{m}\xspace}		% single messages
\newcommand{\mymessg}[1]{\ensuremath{\messg_{\text{#1}}}\xspace} %meaningful messages
\newcommand{\ssome}{\mystate{\ensuremath{\exists\neg\forall}}}
\newcommand{\sall}{\mystate{\ensuremath{\forall}}}
\newcommand{\snone}{\mystate{\ensuremath{\emptyset}}}
\newcommand{\msome}{\mymessg{some}}
\newcommand{\mall}{\mymessg{all}}
\newcommand{\mnone}{\mymessg{none}}
\newcommand{\Lall}{\mylang{all}}
\newcommand{\Lbound}{\mylang{bound}}
\newcommand{\Llack}{\mylang{lack}}
\newcommand{\asome}{\myact{\ensuremath{\exists\neg\forall}}}
\newcommand{\aall}{\myact{\ensuremath{\forall}}}

\definecolor{mygray}{cmyk}{0.35,0.35,0.35,0.35}
\newcommand{\mygray}[1]{{\textcolor{mygray}{#1}}}
%%%

%--------------------
%
\usepackage{setspace}
\doublespacing
%
%-------------------

\title{Co-evolution of lexical meaning \& pragmatic use}

% \author{%\bf NAME1 and NAME2\\
%     ( -- draft \today --- )
% }

\date{}

\begin{document}

  %% for arrow head placement
  \tikzset{->-/.style={decoration={
  markings,
  mark=at position #1 with {\arrow{>}}},postaction={decorate}}}
  %%%

\maketitle

\begin{abstract}
  According to standard linguistic theory, the meaning of an utterance is the product of
  conventional semantic meaning and general pragmatic rules on language use. To investigate how
  cultural evolution of language plays out under this picture of the semantics-pragmatics
  interface, we present a game-theoretic model of the competition between types of language
  users, each endowed with certain lexical representations and a particular pragmatic
  disposition to act on them. Our model traces two evolutionary forces and their interaction:
  (i) pressure towards communicative efficiency and (ii) transmission perturbations during the
  acquisition of linguistic knowledge. We illustrate the model based on a case study on scalar
  implicatures, which suggests that the relationship between underspecified semantics and
  pragmatic inference is one of co-evolution. The evolution of underspecified semantics
  requires pragmatic reasoning, while pragmatic reasoning may not evolve if lexical meanings do
  not require pragmatic enrichment.
\end{abstract}

\section{Introduction}\label{sec:introduction}

What is conveyed usually goes beyond what is said. A request for a blanket can be politely
veiled by uttering ``I'm cold;'' a temporal succession of events can be communicated by the order
in which conjuncts appear as in ``I traveled to Paris and got married;'' an invitation can be
declined by saying ``I have to work.'' An influential explanation of the relation between the
literal meaning of expressions and what they may convey in context is due to
\citet{grice:1975}, who characterizes pragmatic use and interpretation as a process of mutual
reasoning about rational language use. For instance, under the assumption that the speaker is
cooperative and relevant, ``I have to work'' may be interpreted as providing a reason why the
speaker will not be able to accept an invitation, going beyond its literal meaning. Some of
these enrichments are rather \emph{ad hoc}. Others show striking regularities, such as the use
of ability questions for polite requests (``Could you please \dots?''), or certain enrichments
of lexical meanings such as \emph{and} to convey \emph{and then}.

A particularly productive and well studied class of systematic pragmatic enrichments are scalar
implicatures
\citep{horn:1984,Hirschberg1985:A-Theory-of-Sca,LevinsonPragmatics1983,Geurts2010:Quantity-Implic}. Usually,
the utterance of a sentence like ``I own some of Johnny Cash's albums'' will be taken to mean
that the speaker does not own all of them. This is because, if the speaker instead owned them
all, she could have used the word \emph{all} instead of \emph{some} in her utterance, thereby
making a more informative statement. Scalar implicatures, especially the inference from
\emph{some} to \emph{some but not all}, have been studied extensively, both theoretically
\citep[e.g.,][]{Sauerland2004:Scalar-Implicat,ChierchiaFox2008:The-Grammatical,Rooyvan-RooijJagerde-Jager2012:Explaining-Quan}
as well as experimentally
\citep[e.g.,][]{BottNoveck2004:Some-Utterances,huang+snedeker:2009,GrodnerKlein2010:Some-and-Possib,GoodmanStuhlmuller2013:Knowledge-and-I,DegenTanenhaus2012:Processing-Scal}. This
makes them particularly suitable candidates for the study of the evolution of regular pragmatic
inferences.  While there is much dispute in this domain about many details, a position endorsed
by a clear majority is that a scalar item like \emph{some} is underspecified to mean \emph{some
  and maybe all} and that the enrichment to \emph{some but not all} is part of some regular
process with roots in pragmatics. If this majority view is correct, the question arises how
such a division of labor between semantics and pragmatics could have evolved.

Models of language evolution abound. There are simulation-based models studying populations of
communicating agents
\citep{Hurford1989:Biological-Evol,Steels1995:A-Self-Organizi,LenaertsJansen2005:The-Evolutionar,SteelsBelpaeme2005:Coordinating-Pe,BaronchelliPuglisi2008:Cultural-route-,steels:2011,SpikeStadler2016:Minimal-Require}
and there are mathematical models of language evolution, many coming from game theory
\citep{Warneryd1993:Cheap-Talk-Coor,BlumeKim1993:Evolutionary-St,nowak+krakauer:1999,nowak:2006,Huttegger2007:Evolution-and-t,Skyrms2010:Signals}. Much
of this work has focused on explaining basic properties such as compositionality and
combinatoriality
\citep[e.g.,][]{Batali1998:Computational-S,nowak+krakauer:1999,nowak+etal:2000,KirbyHurford2002:The-Emergence-o,kirby:2002,SmithKirby2003:Iterated-Learni,Gong2007:Language-Evolut,kirby+etal:2015,verhoef+etal:2014,Franke2015:Proto-Syntax},
but little attention has been paid to the interaction between conventional meaning and
pragmatic use. What is more, many mathematical models explain evolved meaning as a regularity
in the overt behavior of agents, abstracting from complex interactions between semantic
representations and pragmatic use.

In contrast, we will here look at language users with a richer linguistic make-up. The objects
of replication and selection in our model are pairs consisting of a set of lexical meanings and
a manner of pragmatic behavior, which together define an agent's linguistic choices. Types of
pragmatic behavior are represented using probabilistic models of pragmatic language use
\citep{frank+goodman:2012,FrankeJager2015:Probabilistic-p,GoodmanFrank2016:Pragmatic-Langu}. Replication
and selection are described by the \emph{replicator mutator dynamic}, a general and established
model of evolutionary change in large and homogeneous populations
\citep{Hofbauer1985:The-Selection-M,nowak+etal:2000,NowakKomarova2001:Evolution-of-Un,hofbauer+sigmund:2003,nowak:2006}. This
approach allows us to study the interaction between (i) pressure towards communicative
efficiency and (ii) infidelity in the transmission of linguistic knowledge, caused by factors
such as inductive learning biases and sparse learning data
\citep{KirbyHurford2002:The-Emergence-o,SmithKirby2003:Iterated-Learni,KirbyGriffith2014:Iterated-Learni}. Considering
the effects that the transmission of linguistic knowledge through (iterated) learning have is
important because neither semantic meaning nor pragmatic usage patterns are directly
observable. Instead, language learners have to infer these unobservables from the observable
behavior in which they result. We formalize this process as a form of Bayesian inference. Our
approach thereby contains a well-understood model of iterated Bayesian learning
\citep{griffiths+kalish:2005,griffiths+kalish:2007}, but combines it with functional selection,
here formalized as the most versatile dynamic from evolutionary game theory, the replicator
dynamic \citep{TaylorJonker1978:Evolutionary-St}. Section~\ref{sec:model} introduces this
model.

Section~\ref{sec:si-case-study} applies this model to a case study on scalar implicatures. It
shows that inductive learning biases of Bayesian learners that favor simpler lexical meanings
may play a role in preventing the lexicalization of scalar inferences and thereby lead to the
emergence of Gricean-like pragmatic reasoning types, but only if there is selective pressure
for communicative efficiency and learnability. The results of this study are critically
assessed in the light of the assumptions that feed our model in Section~\ref{sec:discussion}.

% We see the main contribution of this work as conceptual and technical, not as a definite answer
% to the question why scalar implicatures emerge. It rather demonstrates how current
% probabilistic cognitive modeling of language use and evolutionary modeling can be fruitfully
% combined to study the co-evolution of semantics and pragmatics side-by-side. Reversely, the
% approach taken here may be seen as a first step towards giving an evolutionary rationale for
% empirically successful probabilistic models of language use that embrace the majority view of
% the division of labor between semantics and pragmatics. Section~\ref{sec:discussion} elaborates
% on these points.

\section{A model of evolving lexical representations and pragmatic behavior}
\label{sec:model}

\subsection{Communicative success and learnability}

The emergence and change of linguistic structure is influenced by many intertwined
factors. These range from biological and socio-ecological to cultural ones \citep{benz+etal:2005b,steels:2011,tamariz+kirby:2016}. Social and ecological pressures determine communicative needs, while
biology determines the architecture that enables and constrains the means by which they can be
fulfilled. In the following, our focus lies on cultural aspects, wherein processes of
linguistic change are viewed as shaped by language use and its transmission, i.e., as a result
of a process of cultural evolution
\citep{Pagel2009:Human-Language-,ThompsonKirby2016:Culture-Shapes-}.

The idea that language is an adaptation to serve a communicative function is fundamental to
many synchronic and diachronic analyses, at least since \citeposs{zipf:1949} explanation of
word frequency rankings as a result of competing hearer and speaker preferences \citep[e.g.,
in][]{martinet:1962, horn:1984,jaeger+vRooij:2007,jaeger:2007,
  piantadosi:2014,kirby+etal:2015}. If processes of selection, such as conditional imitation or
reinforcement, favor more communicatively efficient types of behavior, languages are driven
towards semantic expressivity \citep[e.g.,][]{nowak+krakauer:1999,Skyrms2010:Signals}. But
pressure towards communicative efficiency is not the only force that shapes
language. Learnability is another, as natural languages need to be learnable to survive their
faithful transmission across generations. Furthermore, the effects of even small learning
biases can have striking effects on an evolving language in a process of iterated learning
\citep{KirbyHurford2002:The-Emergence-o,SmithKirby2003:Iterated-Learni,kirby+etal:2014}.

While natural languages are pressured for both communicative efficiency and learnability, these forces may
pull in opposite directions \citep[\S7]{christiansen+chater:2008}. Their opposition becomes particularly clear when considering the
extreme (\citealt{kemp+regier:2012,kirby+etal:2015}). A language consisting of a single
form-meaning association is easy to learn but likely bad for communication with peers. Conversely, a language
that lexicalizes distinct forms for a large number of different meanings may be better for communication 
but challenging to acquire.

\subsection{The replicator mutator dynamic}
An elegant formal approach to capture the interaction between selection for communicative
efficiency and learnability is the \emph{replicator mutator dynamic}
\citep{Hofbauer1985:The-Selection-M,nowak+etal:2000,NowakKomarova2001:Evolution-of-Un,hofbauer+sigmund:2003,nowak:2006}. The
RMD is a population dynamic which describes the average change in the composition of a
community of agents of different types. The dynamic assumes a virtually infinite population, so
that it is able to abstract away from what individual agents do. It describes the most probable
trajectory of change in the population vector $\vec{x}$, which codifies relative frequencies of
types, so that $x_i$ is the relative frequency of type $i$ and $\sum\nolimits_i x_i = 1$. In
its simplest discrete-time formulation, the RMD defines the frequency $x'_i$ of each type $i$
in the population at the next time step as a function of: (i) the frequency $x_i$ of each type
$i$ before the update step, (ii) the fitness $f_i$ of each type $i$ before the update, and
(iii) the probability $Q_{ji}$ that an agent who observes the behavior of an agent with type
$j$ ends up acquiring type $i$ (with $\sum_k Q_{jk} = 1$ for all $j$):
\begin{align}
  \label{eq:RMD_discrete}
  x'_i = \sum_j \frac{x_jf_j Q_{ji}}{\Phi}\,,
\end{align}
where $\Phi = \sum_k x_k f_k$ is the average fitness in the population.

The RMD consists of two components: fitness-based selection and transmission biases encoded in
the so-called mutation matrix $Q$. This becomes most transparent when we consider an equivalent
formulation in terms of a step-wise application of the discrete-time replicator dynamic
\citep{TaylorJonker1978:Evolutionary-St} on the initial population vector $\vec{x}$ and a
subsequent multiplication with the mutation matrix $Q$:
\begin{align}
  \label{eq:RMD_discrete_recast}
  x'_i & = (\text{M}(\text{RD}(\vec{x})))_i\,,
\end{align}
where
\begin{align*}
      \left ( \text{RD}(\vec{x}) \right )_i 
         = \frac{x_i f_i}{\Phi}
 \ \ \ \ \text{and} \ \ \ \ 
  (\text{M}(\vec{x}))_i = (\vec{x} \cdot Q)_i = \left ( \sum_j
          x_j Q_{ji} \right)_i\,.
\end{align*}
%
If there is no mutation, so that types are always replicated faithfully with $Q_{ji}=1$
whenever $j=i$, the RMD reduces to the replicator dynamic. If fitness is the same for all
types, $f_i = f_j$ for all $i$ and $j$, the RMD reduces to a process of iterated
mutation. Appendix~\ref{sec:minim-example-repl} provides an abstract minimal example.

\subsection{Interpreting the replicator mutator dynamic}

The RMD is an abstract description of population-level changes. It abstracts away from specific
assumptions about how agents interact and what causes selection and mutation. To understand
this better, the following looks at each component in isolation: first replication, then
mutation.

According to the replicator dynamic the relative frequency $x_i$ of type $i$ increases
proportional to its average fitness in the population. This dynamic is popular and versatile
because it can be derived from many processes of biological and cultural transmission and
selection \citep[for overview and several derivations
see][]{Sandholm2010:Population-Game}. Under a biological interpretation, the model describes
population-level changes when the number of offspring under single-parent reproduction is
proportional to the parent's fitness. Under a cultural interpretation, examples of concrete
agent-level processes that lead to population-level changes as described by the replicator
dynamic include conditional imitation
\citep[e.g.,][]{Helbing1996:A-Stochastic-Be,Schlag1998:Why-Imitate-and} or reinforcement
learning \citep[e.g.,][]{BorgersSarin997:Learning-Throug,Beggs2005:On-the-Converge}. In
conditional imitation, for instance, each agent carries a fixed type $i$ which it may change to
a type $j$ when it happens to observe another agent using type $j$. Imitation occurs with a
probability that is a monotonic function of $f_j$, i.e., of how well $j$ fares against the
population: the better the behavior $j$, the more likely it is to be imitated.\footnote{A
  crucial assumption in all of the derivations of the replicator dynamic as a biological or
  cultural process of evolution is that agents themselves are rather limited in their strategic
  capacities. They are, for example, not able to innovate a strategy, i.e., to just choose a
  type, that is optimal given the current population state. They are also not able to
  strategically adapt their behavior to any particular agent they are interacting with on some
  given occasion. Such minimalism about the strategic capacities of agents is a conservative
  methodological choice. It allows us to study ---as the formal thought experiment that any
  such model is--- how fitness-based selection changes a population in the absence of optimal
  agent-level reasoning and it abstracts away from specific assumptions about how fitness-based
  selection might be realized concretely.}

If all types have equal fitness at all times, the replicator mutator dynamic reduces to a
process of iterated mutation. Mutation can be interpreted as biological or cultural as
well. Under a biological interpretation where types are genetically encoded behavioral traits,
$Q_{ji}$ is the probability by which offspring of type $j$ genetically mutates into type
$i$. Under a cultural interpretation where types are transmitted by observation or learning,
such as in conditional imitation, $Q_{ji}$ is the probability that an agent trying to learn,
adopt or imitate the behavior of an agent with type $j$ acquires type $i$. As detailed below,
in this way, the process in (\ref{eq:RMD_discrete}), and its equivalent in
(\ref{eq:RMD_discrete_recast}), can contain a model of iterated learning
\citep{griffiths+kalish:2005,griffiths+kalish:2007}.

In the following application to language evolution, we prefer a cultural interpretation of the
mutator component as the perturbation induced by how easy different linguistic traits are
learnable from observation of language use. Nothing of current relevance hinges on whether the
replicator component is interpreted as a biological process or a cultural process. We believe
that this flexibility of interpretation is an asset. It abstracts away from specific
assumptions and focuses on a general and versatile description of fitness-based
selection. Still, to have something concrete in mind when interpreting the following
application, Appendix~\ref{sec:deriv-repl-mutat} derives the RMD from a specific scheme of
agent-level conditional imitation.

\subsection{Fitness \& learnability of lexical meanings \& pragmatic strategies}
\label{sec:fitn--learn}

Our goal is to apply the RMD to the evolution of lexical representations and pragmatic
behavior. This requires determining three things: (i) what the relevant types are, (ii) how
fitness of a type derives from its communicative success and (iii) how the mutation matrix $Q$
is computed. These issues are addressed, one by one, in the following.

\subsubsection{Types: Lexica and pragmatic strategies}
\label{sec:languages+use}
Types are what evolution operates on. A type is what determines an agent's behavior and thereby
an agent's fitness, so that it makes sense to speak of the fitness of a type itself. To study
the joint evolution of semantic meaning and pragmatic use, types are pairs consisting of a
lexicon and a linguistic strategy of language use.

Agents play signaling games, in which the speaker wants to communicate a world state $s$ with a
message $m$ to a hearer who receives $m$ but does not know $s$
\citep[e.g.][]{lewis:1969,Skyrms2010:Signals}. A lexicon associates each message with a set of
states. A linguistic behavior specifies a probabilistic speaker rule (a probabilistic choice of
message for each state) and a probabilistic hearer rule (a probabilistic choice of state for
each message) given a lexicon. There are many ways of making these general notions
concrete. In the following we aim at a general formulation of speaker and listener
behavior which is as simple as possible but still in line with both a game-theoretic tradition
\citep[e.g.,][]{benz:2006,benz+vRooij:2007,jaeger:2007b,franke+jaeger:2014} and a Bayesian
tradition
\citep[e.g.,][]{frank+goodman:2012,goodman+stuhlmueller:2013,GoodmanFrank2016:Pragmatic-Langu}.

Lexica codify the truth-conditions of expressions. A convenient way to represent lexica is by
$(\card{\States}, \card{\Messgs})$-Boolean matrices, where $\States$ is a set of states
(meanings) and $M$ a set of messages (forms available in the language). For example, suppose
that there are two relevant world states $\States = \set{\ssome, \sall}$. In state $\ssome$
Chris owns some but not all of Johnny Cash's albums while in $\sall$ Chris owns them
all. Suppose that there are two messages $\Messgs = \set{\msome, \mall}$ where $\msome$ is
short for a sentence like \emph{Chris owns some of Johnny Cash's albums} and $\mall$ for the
same sentence with \emph{some} replaced by \emph{all}.  Lexica then assign a Boolean truth
value, either $0$ for false or $1$ for true, to each state-message pair. The following two
lexica are minimal examples for the distinction between a lexicalized upper-bound for
\emph{some}, in $\Lbound$, and the widely assumed logical semantics with only a lower-bound, in
$\Llack$.
\begin{align*}
  \Lbound & = \begin{blockarray}{lcc}
    & \mygray{\msome} & \mygray{\mall} \\
    \begin{block}{l[cc]}
      \mygray{\ssome} & 1 & 0 \\
      \mygray{\sall}  & 0 & 1 \\
    \end{block}
  \end{blockarray} &
  \Llack & = \begin{blockarray}{lcc}
    & \mygray{\msome} & \mygray{\mall} \\
    \begin{block}{l[cc]}
      \mygray{\ssome} & 1 & 0 \\
      \mygray{\sall}  & 1 & 1 \\
    \end{block}
  \end{blockarray}
\end{align*}

Linguistic strategies define dispositions to produce and interpret messages given a lexicon. We
distinguish between two kinds strategies. {\em Literal interlocutors} produce and
interpret messages literally, being guided only by their lexica. {\em Pragmatic interlocutors}
instead engage in mutual reasoning to inform their choices. Recent probabilistic models of
rational language use
\citep{frank+goodman:2012,FrankeJager2015:Probabilistic-p,GoodmanFrank2016:Pragmatic-Langu}
capture different types of pragmatic behavior in a reasoning hierarchy. The hierarchy's bottom,
level $0$, corresponds to literal language use, as in Equations~(\ref{h:level0}) and
(\ref{s:level0}). Pragmatic language users of level $n + 1$ act (approximately) rational with
respect to level-$n$ behavior of their interlocutors, as in Equations~(\ref{h:leveln}) and
(\ref{s:leveln}).
\begin{alignat}{3}
&H_{0}(s \mid m;L) &&\propto pr(s) L_{[s,m]} \label{h:level0}\\
&S_{0}(m \mid s;L) &&\propto \exp(\lambda \; L_{[s,m]}) \label{s:level0}\\
&H_{n+1}(s|m;L) &&\propto pr(s) S_{n}(m|s;L) \label{h:leveln}\\
&S_{n+1}(m|s;L) &&\propto  \exp(\lambda \; H_{n}(s|m;L)) \label{s:leveln}
\end{alignat}

According to (\ref{h:level0}), a literal hearer's interpretation of a message depends on her
lexicon and her prior over states, $pr \in \Delta(S)$, which is here assumed flat for
simplicity. Literal interpreters thereby choose an arbitrary true interpretation for each
message according to their lexicon. Pragmatic hearers, defined in (\ref{h:leveln}), instead use
Bayes' rule to weigh interpretations based on a conjecture about speaker behavior. Speaker
behavior is regulated by a soft-max parameter $\lambda \geq 0$
\citep{luce:1959,sutton+barto:1998}. As $\lambda$ increases, choices approximate strict
maximization of expected utilities. Expected utility of a message $\messg$ in state $\state$
for a level $n+1$ speaker is here defined as $H_{n}(s|m;L)$, the probability that the hearer
will assign to or choose the correct meaning.\footnote{An alternative choice for the speaker's
  utility function is $\log H_{n}(s|m;L)$. Conceptually, the difference is that without the
  logarithm, utilities are derived from action choices, while with the logarithm utilities are
  derived from how closely the listener's beliefs (after hearing an utterance) align with the
  speaker's, as measure by Kullback-Leibler divergence
  \cite{frank+goodman:2012,qing+franke:2015}. In an evolutionary context, where fitness comes
  from actual interactions with the environment, not just beliefs held about it, the
  formulation without the logarithm may be slightly more appealing. Moreover, log-based speaker
  utilities cause the speaker never to use false utterances. Since we aim here for a situation
  in which any type has a reasonable chance to be mistaken for any other based on their
  production behavior (see Section~\ref{sec:learnability}), we want a definition of speaker
  types which also send any message in any state with positive probability.} For literal
speakers, utility only tracks truthfulness. Literal speakers choose any true message with equal
probability but may send false messages as well with a probability dependent on $\lambda$.

In words, an agent's linguistic behavior -- what message to send when or which message to
interpret how -- is defined by (i) her reasoning level and (ii) her lexicon. In what follows,
this is what we identify an agent's type with, in order to analyze under which conditions
combinations of (i) and (ii) emerge through cultural evolution. Intuitively, literal behavior
(level-$0$ reasoning) results from unreflected language use. Such agents produce and comprehend
solely based on the truth-conditions of their lexicon. In a Gricean spirit, pragmatic behavior
(level-$n+1$ reasoning) results from reasoning about rational language use. Such agents produce
and comprehend by reasoning about how they themselves would interpret or use
expressions. Higher order behavior is therefore, even if in tendency rational, quite simple. It
does not assume agents to know their interlocutor's actual type when interacting (e.g., what
lexicon the interlocutor is using); they simply act based on a conjecture about language use
derived from their own (solipsistic) perspective.

The following examples illustrate these behaviors numerically. A literal interpreter with lexicon $\Lbound$ assigns $\ssome$
a probability of $H_0(\ssome \mid \msome; \Lbound) = 1$ after hearing $\msome$, while a literal
interpreter with $\Llack$ has $H_0(\ssome \mid \msome; \Llack) = 0.5$:
\begin{align*}
  H_0(\cdot \mid \cdot, \Lbound) & = \begin{blockarray}{lcc}
    & \mygray{\ssome} & \mygray{\sall} \\
    \begin{block}{l[cc]}
      \mygray{\msome} & 1 & 0 \\
      \mygray{\mall}  & 0 & 1 \\
    \end{block}
  \end{blockarray} &
  H_0(\cdot \mid \cdot, \Llack) & = \begin{blockarray}{lcc}
    & \mygray{\ssome} & \mygray{\sall} \\
    \begin{block}{l[cc]}
      \mygray{\msome} & .5 & .5 \\
      \mygray{\mall}  & 0 & 1 \\
    \end{block}
  \end{blockarray} 
\end{align*}
By contrast, pragmatic hearers of level 1 have the following interpretative behavior for $\lambda = 1$:
\begin{align*}
  H_1(\cdot \mid \cdot, \Lbound) & \approx \begin{blockarray}{lcc}
    & \mygray{\ssome} & \mygray{\sall} \\
    \begin{block}{l[cc]}
      \mygray{\msome} & .73 & .27 \\
      \mygray{\mall}  & .27 & .73 \\
    \end{block}
  \end{blockarray} &
  H_1(\cdot \mid \cdot, \Llack) & \approx \begin{blockarray}{lcc}
    & \mygray{\ssome} & \mygray{\sall} \\
    \begin{block}{l[cc]}
      \mygray{\msome} & .59 & .41 \\
      \mygray{\mall}  & .35 & .65 \\
    \end{block}
  \end{blockarray} 
\end{align*}
This is the outcome of reasoning about their level-$0$ speaker counterparts with $\lambda = 1$:
\begin{align*}
  S_0(\cdot \mid \cdot, \Lbound) & \approx \begin{blockarray}{lcc}
    & \mygray{\msome} & \mygray{\mall} \\
    \begin{block}{l[cc]}
      \mygray{\ssome} & .73 & .27 \\
      \mygray{\sall}  & .27 & .73 \\
    \end{block}
  \end{blockarray} &
  S_0(\cdot \mid \cdot, \Llack) & \approx \begin{blockarray}{lcc}
    & \mygray{\msome} & \mygray{\mall} \\
    \begin{block}{l[cc]}
      \mygray{\ssome} & .73 & .27 \\
      \mygray{\sall}  & .5 & .5 \\
    \end{block}
  \end{blockarray} 
\end{align*}
With low $\lambda$ speakers choose true messages with more slack. Reasoning over this behavior therefore also results in a weaker association of messages with only true states in receivers, but also in a slightly stronger association of $\msome$ with $\ssome$ over $\sall$ for $\Llack$ users because they reason that $S_0(\msome | \ssome;\Llack) > S_0(\msome | \sall;\Llack)$. For $\lambda = 20$, there will be less slack in literal speaker behavior:
\begin{align*}
  S_0(\cdot \mid \cdot, \Lbound) & \approx \begin{blockarray}{lcc}
    & \mygray{\msome} & \mygray{\mall} \\
    \begin{block}{l[cc]}
      \mygray{\ssome} & 1 & 0 \\
      \mygray{\sall}  & 0 & 1 \\
    \end{block}
  \end{blockarray} &
  S_0(\cdot \mid \cdot, \Llack) & \approx \begin{blockarray}{lcc}
    & \mygray{\msome} & \mygray{\mall} \\
    \begin{block}{l[cc]}
      \mygray{\ssome} &  1 & 0 \\
      \mygray{\sall}  & .5 & .5 \\
    \end{block}
  \end{blockarray} 
\end{align*}
And accordingly less slack in level $1$ pragmatic interpretation:
\begin{align*}
  H_1(\cdot \mid \cdot, \Lbound) & \approx \begin{blockarray}{lcc}
    & \mygray{\ssome} & \mygray{\sall} \\
    \begin{block}{l[cc]}
      \mygray{\msome} & 1 & 0 \\
      \mygray{\mall}  & 0 & 1 \\
    \end{block}
  \end{blockarray} &
  H_1(\cdot \mid \cdot, \Llack) & \approx \begin{blockarray}{lcc}
    & \mygray{\ssome} & \mygray{\sall} \\
    \begin{block}{l[cc]}
      \mygray{\msome} & 0.67 & 0.33 \\
      \mygray{\mall}  & 0 & 1 \\
    \end{block}
  \end{blockarray} 
\end{align*}
Lastly, turning to types that have no bearing on the choices of hearers of level 1, with $\lambda = 1$ pragmatic speakers of level 1 have:
\begin{align*}
  S_1(\cdot \mid \cdot, \Lbound) & \approx \begin{blockarray}{lcc}
    & \mygray{\msome} & \mygray{\mall} \\
    \begin{block}{l[cc]}
      \mygray{\ssome} & .73 & .27 \\
      \mygray{\sall}  & .27 & .73 \\
    \end{block}
  \end{blockarray} &
  S_1(\cdot \mid \cdot, \Llack) & \approx \begin{blockarray}{lcc}
    & \mygray{\msome} & \mygray{\mall} \\
    \begin{block}{l[cc]}
      \mygray{\ssome} & .62 & .38 \\
      \mygray{\sall}  & .38 & .62 \\
    \end{block}
  \end{blockarray} 
\end{align*}
For $\lambda = 20$, pragmatic speaker behavior of level 1 is instead as follows:
\begin{align*}
  S_1(\cdot \mid \cdot, \Lbound) & \approx \begin{blockarray}{lcc}
    & \mygray{\msome} & \mygray{\mall} \\
    \begin{block}{l[cc]}
      \mygray{\ssome} & 1 & 0 \\
      \mygray{\sall}  & 0 & 1 \\
    \end{block}
  \end{blockarray} &
  S_1(\cdot \mid \cdot, \Llack) & \approx \begin{blockarray}{lcc}
    & \mygray{\msome} & \mygray{\mall} \\
    \begin{block}{l[cc]}
      \mygray{\ssome} & 1 & 0 \\
      \mygray{\sall}  & 0 & 1 \\
    \end{block}
  \end{blockarray} 
\end{align*}

In contrast to their literal counterparts of level $0$, pragmatic agents of level $1$ who use
$\Llack$ associate $\msome$ preferentially with $\ssome$. This association is not perfect, and
usually less strong than what agents with a lexicalized upper bound in $\Lbound$ can achieve --
with or without pragmatic reasoning. Higher order reasoning beyond level $1$ leads to stronger
associations of $\msome$ and $\ssome$ also for the receiver. Still, the case study presented in
Section~\ref{sec:si-case-study} will consider sender and receiver behavior at levels 0 and 1,
as the latter are the simplest pragmatic reasoning types which show a tendency to
communicatively attuned pragmatic enrichment. Using only level 1 reasoning and possibly small
$\lambda$ is therefore a conservative choice that works \emph{against} the fitness-based
selection of pragmatic language use for a notion of fitness defined as communicative success,
which is introduced next.

When it comes to competition between types of use of lexicon $\Lbound$, pragmatic reasoning at
level $1$ is not advantageous. The reason for this is that literal use of $\Lbound$ already
endows agents with a behavioral strategy that associates a single state with a single message
(in tendency; depending on $\lambda$ for senders). For $\Lbound$-receivers of level $1$,
reasoning over stochasticity introduced at $S_0( \cdot \mid \cdot, \Lbound)$ will generally
decrease the association of one state with one message. This decrease is only slight if
$\lambda$ is high, but nevertheless present. That is to say, level-$1$ reasoning does not
necessarily confer a functional advantage. For some types, such as users of $\Lbound$, literal
signaling is preferable. In sum, the definition of types introduced here constitute conditions
that are purposefully averse to what we would like to show; the evolution of is not immediate
just based on fitness-based selection, to which we turn next.

\subsubsection{Fitness \& fitness-based selection based on communicative success}\label{sec:expressivity}

In the context of language evolution, fitness is usually associated with the ability to
successfully communicate with other language users from the same population
\citep[e.g.,][]{nowak+krakauer:1999,nowak+etal:2000, nowak+etal:2002}. Concretely, the fitness
of type $i$ is its average expected communicative success, or \emph{expected utility} (EU),
given the relative frequencies $x_j$ of types $j$ in the current population:
\begin{align*}
  f_i = \sum_j x_j \text{EU}(t_i,t_j)\,.
\end{align*}
The expected utility $\text{EU}(t_i,t_j)$ for type $i$ when communicating with type $j$ is the
average success of $i$ when talking or listening to $j$. If, as standardly assumed, agents are speakers half of the
time, this yields:
\begin{align*}
  \text{EU}(t_i,t_j) = \nicefrac{1}{2} \, \text{EU}_S(t_i,t_j) + \nicefrac{1}{2} \, \text{EU}_H(t_i,t_j)\,,
\end{align*}
where $\text{EU}_S(t_i,t_j)$ and $\text{EU}_H(t_i,t_j)$ are the expected utilities for $i$ as a
speaker and as a hearer when communicating with $j$, defined as follows, where $n_i$ and $n_j$
are type $i$'s and type $j$'s pragmatic reasoning types and $L_i$ and $L_j$ are their lexica:
\begin{flalign*}
  & \text{EU}_S(t_i,t_j)  = \sum_s P(s)\sum_m S_{n_i}(m \mid s;L_i) \sum_{s'} H_{n_j}(s' \mid m;L_j)
  \delta(s,s') \,, \\
 & \text{EU}_H(t_i,t_j)  = \text{EU}_S(t_j,t_i)\,.
\end{flalign*}
As usual, we assume that agents are cooperative: $\delta(s,s') = 1$ iff $s = s'$ and $0$ otherwise.

In words, expected utility $\text{EU}(t_i,t_j)$ quantifies how successful communication between
agents of types $i$ and $j$ is, with each type's behavior resulting from a combination of a
lexicon and a reasoning level (Section~\ref{sec:languages+use}). Fitness $f_i$ indicates how
well type $i$ fares in a population $\vec{x}$ where the probability of meeting a type $j$ is
$x_j$.

\subsubsection{Learnability}
\label{sec:learnability}

Languages are shaped not only by functionalist forces towards greater communicative
success. Another important factor is the fidelity with which linguistic knowledge is
transmitted. In biological evolution, where types are expressed genetically, transmission
infidelity comes into the picture through infrequent and mostly random mutation and genetic
drift \citep{kimura:1983}. However, an agent's lexicon and a disposition for pragmatic
reasoning are likely not inherited genetically. They need to be learned from observation, such
as in conditional imitation (see Appendix~\ref{sec:deriv-repl-mutat}).  Concretely, when agents
learn from or imitate type $j$, they observe the overt linguistic behavior of type $j$ and need
to infer the covert type that most likely produced the observed behavior.

Iterated learning is a process in which languages are learned repeatedly from the observation
of linguistic behavior of agents who have themselves acquired their behavior from observation
and inference. In the simplest case there is a single teacher and a single learner in each
generation \citep[e.g.,][]{kirby:2001,brighton:2002}. After sufficient training the learner
becomes a teacher and produces behavior that serves as input for a new learner. Due to the
pressure towards learnability it exerts, iterated learning alone generally leads to simpler and
more regular languages (see \citealt{kirby+etal:2014} and \citealt{tamariz+kirby:2016} for
recent surveys).

Following \citet{griffiths+kalish:2005,griffiths+kalish:2007} we model language acquisition as
a process of Bayesian inference in which learners combine the likelihood of a type producing
the witnessed learning input with prior inductive biases. In a Bayesian setting these biases
can be codified in a prior $P \in \Delta(T)$, which reflects the amount of data a learner
requires to faithfully acquire the language of the teacher
\citep[cf.][450]{griffiths+kalish:2007}. Put differently, types favored by the prior are easier
to learn because they require less data than less favored types to be inferred, even if they
are both compatible with the learning data at hand.

The extent of the prior's influence has been shown to heavily depend on the learning strategy
assumed to underlie the inference process. Early simulation results suggested that weak biases
could be magnified by exposing learners to only small data samples
\citep[e.g. in][]{brighton:2002}. \citet{griffiths+kalish:2007} showed that iterated learning
converges to the prior in the limit, i.e., that the resulting distribution over types
corresponds to the learners' prior distribution and is not influenced by the amount of input
given to them, if learners select random hypothesis from their posterior beliefs about which
types may have generated the learning data that they saw. More deterministic strategies such as
the adoption of the type with the highest posterior probability, so-called {\it maximum a
  posterior estimation} (MAP), increase the influence of both the prior and the data
\citep{griffiths+kalish:2007,kirby+etal:2007}. In the following, we use a parameter $l\ge1$ to
modulate between posterior sampling and the MAP strategy. When $l = 1$ learners sample from the
posterior. The learners' propensity to maximize the posterior grows as $l$
increases.\footnote{The way the posterior over types is formed and how learners employ it are
  important design choices. There are other alternatives as well. For instance, instead of
  adopting a single type, learners could keep their uncertainty over types and base their
  production behavior on the posterior over types they form \citep{burkett+griffiths:2010}. A
  straightforward operationalization of this idea would then see production defined as a
  two-step process wherein a speaker first samples from her posterior and then produces
  utterances based on the type she sampled, as done in \citet{kirby+etal:2015}. Beyond adding
  complexity to the model, a disadvantage of this approach is that it could happen that a
  speaker that first uttered a Hungarian utterance speaks Quechua in the next, so to speak. We
  therefore maintain the simpler assumption of a single type being acquired in the present
  context, and postpone further investigation of more complex type-adoption schemes to future
  research.}

Let $D$ be the set of possible data that learners may be exposed to. This set contains all
sequences of state-message pairs of length $k$, e.g.,
$\tuple{\tuple{s_1,m_1},\dots , \tuple{s_k,m_k}}$. As $k$ increases, learners have more data to base their inference on and so tend to
recover the true types that generated a given sequence with higher probability. The mutation
matrix $Q$ of the replicator mutator dynamic in (\ref{eq:RMD_discrete}) can then be defined as
follows: $Q_{ji}$ is the probability that a learner acquires type $i$ when learning from an
agent of type $j$. The learner observes length-$k$ sequences $d$ of state-message pairs, but
the probability $P(d \mid t_j)$ with which sequence $d = \tuple{\tuple{s_1,m_1},\dots , \tuple{s_k,m_k}}$ is observed depends on type $j$'s
linguistic behavior:
\begin{align*}
  P(d = \tuple{\tuple{s_1,m_1},\dots , \tuple{s_k,m_k}} \mid t_j) = \prod_{i = 1}^k S_{n_j}(m_i
  \mid s_i; L_{j})\,,
\end{align*}
where, as before, $n_j$ is $j$'s pragmatic reasoning level and $L_j$ is $j$'s lexicon. For a
given observation $d$, the probability of acquiring type $i$ is $F(t_i \mid d)$, so that:
\begin{flalign*}
  Q_{ji} = \sum_{d \in D} P(d \mid t_j) F(t_i \mid d)\,.
\end{flalign*}

The acquisition probability $F(t_i \mid  d)$ given datum $d$ is obtained by probability matching $l = 1$ or a tendency towards choosing the most likely type $l > 1$ from the posterior distribution $P(\cdot \mid d)$ over types given the data. This is calculated by Bayes' rule:
\begin{flalign*}
  & F(t_i \mid d) \propto P(t_i \mid d)^l \; \text{ and }\\
  & P(t_i \mid d) \propto P(t_i) P(d \mid t_i)\,.
\end{flalign*}

When $l=1$ learners sample a type to adopt from their posterior distribution which they obtain
from observing the teacher produce utterances in particular states. From
\citet{griffiths+kalish:2007}, we know that under $l=1$ iterated Bayesian learning will
converge to a population proportion that reflects that prior over types. If $l>1$ learners tend
to pick types which are \emph{a posteriori} more likely with a higher probability than less
likely types. Even in the absence of a prior bias, iterated Bayesian learning with $l>1$ can
lead to populations in which a particular type $t_i$ has a higher proportion than other
types. For $l>1$, much hinges on particular asymmetries in production likelihoods
$P(d \mid t_j)$, and therefore on the composition of the set of types in general. Intuitively
put, if there are many types $t_j$ which all happen to produce data that is most likely to be
produced by $t_i$, this can lead to a larger proportion of $t_i$ in the long run even when
priors over types are uniform. Appendix~\ref{sec:example-iter-bayes} provides an abstract
example.

\subsection{Model summary}

Communicative success and learnability are central to the cultural evolution of language. These
components can be modelled, respectively, as replication based on a measure of fitness in terms
of communicative efficiency and iterated Bayesian learning. Their interaction is described by
the discrete time replicator mutator dynamic in (\ref{eq:RMD_discrete}), repeated here:
\begin{align*}
  x'_i = \sum_j Q_{ji} \frac{x_jf_j}{\sum_h x_h f_h}\,.
\end{align*}
This equation defines the frequency $x'_i$ of type $i$ at the next time step, based on its
frequency $x_i$ before the update step, its fitness $f_i$, and the probability that a learner
infers $i$ when observing the behavior of a type-$j$ agent. Fitness-based can be thought of as
biological (fitness as expected relative number of offspring) or cultural evolution (fitness as
likelihood of being imitated or repeated, see Appendix~\ref{sec:deriv-repl-mutat} for a
concrete example). A type's communicative success depends on how well it communicates within a
population while its learnability depends on the fidelity by which it is inferred by new
generations of learners. The learners' task is consequently to perform a joint inference over
types of linguistic behavior and lexical meaning.

The model has three parameters: $\lambda$ regulates the degree to which speakers
choose messages that appear optimal from the point of view of the agent's own utility measure
(which may be unrelated to the expected utility when communicating with a given population);
$k$ is the length of observations received by each language learner; $l$ regulates where the learners'
inference behavior lies on a spectrum from probability matching to acquisition of the most likely
teacher type.

\section{Case study: scalar implicatures}
\label{sec:si-case-study}

This section engages in a formal thought experiment to address the question: if a population of
language users could freely combine different lexica with different dispositions to act on
them, what are conditions under which the majority view of scalar implicatures could have
evolved? --- Recall that the majority view is that scalar implicatures are non-lexicalized
pragmatic enrichments. Scalar implicature triggers like \emph{some}, \emph{warm} or \emph{may}
are semantically weak expressions for which logically stronger expressions are salient, e.g.,
\emph{all}, \emph{hot} or \emph{must}. For instance, {\em some} is entailed by {\em
  all}. Still, the use of a less informative expression (\emph{some}) when a more informative
one (\emph{all}) could have been used can license a defeasible inference (\emph{some but not
  all}) that stronger alternative is false (cf. \citealt{horn:1972,gazdar:1979}). According to
the majority view, this is a pragmatic inference, not part of the conventional meaning.

In the following we consider what are perhaps two of the simplest non-trivial setups that speak
to this matter and reflect on their limitations in Section~\ref{sec:discussion}. Section
\ref{subsec:example} first investigates a minimal example intended to isolate effects of
replication and mutation and to gain insights into the sense in which semantic meaning and
pragmatic use can be said to co-evolve. We then turn to a larger and more realistic setup in
Section~\ref{sec:setup:-stat-conc}. Section~\ref{sec:simulation-results} describes simulations
for this case and their results.

\subsection{Co-evolution of semantics \& pragmatics in a restricted type space}
\label{subsec:example}

\begin{figure}
\centering
    % \begin{tikzpicture}

  %   \draw[-,dashed,gray] (0,0) -- (5,0) -- (5,5) -- (0,5) -- (0,0);

  %   \draw (0,0) node[fill,circle,gray,inner sep=1mm,scale=0.6,label={[gray,align=left,label distance=0.1mm]west:{\tiny 100\%} \\[-0.25cm] {\tiny lit. $\Lbound$}}] {};
  %   \draw (5,0) node[fill,circle,gray,inner sep=1mm,scale=0.6,label={[gray,align=left,label distance=0.1mm]east:{\tiny 100\%} \\[-0.25cm] {\tiny prag. $\Lbound$}}] {};
  %   \draw (5,5) node[fill,circle,gray,inner sep=1mm,scale=0.6,label={[gray,align=left,label distance=0.1mm]east:{\tiny 100\%} \\[-0.25cm] {\tiny prag. $\Llack$}}] {};
  %   \draw (0,5) node[fill,circle,gray,inner sep=1mm,scale=0.6,label={[gray,align=left,label distance=0.1mm]west:{\tiny 100\%} \\[-0.25cm] {\tiny lit. $\Llack$}}] {};

  %   \draw[<-, thick, indianred] (0.75,0.25) -- (4.25,0.25) node[midway, above] {\footnotesize functional};
  %   \draw[->, thick, indianred] (0.75,4.75) -- (4.25,4.75) node[midway, below] {\footnotesize functional};
  %   \draw[->, thick, indianred] (0.25,4.25) -- (0.25,0.75) node[midway, below, rotate=90] {\footnotesize functional};
  %   \draw[->, thick, indianred] (4.75,4.25) -- (4.75,0.75) node[midway, above, rotate=90] {\footnotesize functional};

  % \path[->, thick, densely dashdotted, Steelblue] (0.75,0.75) edge [bend left] node[sloped,below] {\footnotesize learning\textsubscript{flat}} (2.25,2.45);
  % \path[->, thick, densely dashdotted, Steelblue] (4.25,0.75) edge [bend right] node[sloped,below] {\footnotesize learning\textsubscript{flat}} (2.75,2.45);
  % \path[->, thick, densely dashdotted, Steelblue] (0.75,4.25) edge [bend right] node[sloped,above] {\footnotesize learning\textsubscript{flat}} (2.25,2.75);
  % \path[->, thick, densely dashdotted, Steelblue] (4.25,4.25) edge [bend left] node[sloped,above] {\footnotesize learning\textsubscript{flat}} (2.75,2.75);

  %   \draw[->, thick, goldish] (0.75,-0.25) -- (2.25,-0.25) node[midway, below] {\footnotesize learning\textsubscript{biased}};
  %   \draw[<-, thick, goldish] (2.75,-0.25) -- (4.25,-0.25) node[midway, below] {\footnotesize learning\textsubscript{biased}};
  %   \draw[->, thick, goldish] (-0.25,0.75) -- (-0.25,4.25) node[midway, above,rotate=90] {\footnotesize learning\textsubscript{biased}};
  %   \draw[->, thick, goldish] (5.25,0.75) -- (5.25,4.25) node[midway, below,rotate=90] {\footnotesize learning\textsubscript{biased}};
  %   \draw[->, thick, goldish] (0.75,5.25) -- (4.25,5.25) node[midway, above] {\footnotesize learning\textsubscript{biased}};

  %   \end{tikzpicture}
\caption{Sketch of dynamics on a two-dimensional type space with four types. The $y$-axis
  represents the proportion of types with $\Llack$; the $x$-axis that of pragmatic ones. Arrows
  sketch population trajectories under functional pressure (replicator steps, red arrows) and
  learnability pressure (mutator steps) with either a flat prior (blue dash-dotted arrows) or a
  prior that favors $\Llack$ (solid purple arrows).\mf{I did not change the upper-most arrow
    for biased learnability, because I cannot compile this anyway; I did change some typos in
    arrow labelling though.}}
\label{fig:sketch}
\end{figure}

This section illustrates how replication and mutation act on the evolution of semantic meaning
and pragmatic use in the simplest non-trivial type space, which consists of only the four types
from Section~\ref{sec:languages+use}: a type has either lexicon $\Llack$ or $\Lbound$ and is
either a literal (level-$0$) or a pragmatic (level-$1$) language user. We may think of such a
population as a point in a two-dimensional space. The grey dotted square in
Fig.~\ref{fig:sketch} outlines this space. The space's first dimension determines the
proportion of types that lexicalize ambiguous $\msome$ (the $y$-axis in
Fig.~\ref{fig:sketch}). The second dimension determines the proportion of pragmatic language
users (the $x$-axis in Fig.~\ref{fig:sketch}). Populations at the corners of this space are
monomorphic; all agents in them are of a single type. Points away from the corners represent
mixed populations of various types.\footnote{This two-dimensional representation underspecifies
  the exact proportions of types found in a mixed population. The space's center, for instance,
  could be composed in equal parts of pragmatic $\Llack$ and of literal $\Lbound$. But it could
  equally well be a population that has all four types in equal proportions. This level of
  granularity will nevertheless be sufficient for our purposes, as we are only interested in
  sketching rough trajectories after replicator or mutator steps.} We can predict general
trends in evolutionary trajectories in this space based on what we know from the literature and
the setup from Section~\ref{sec:fitn--learn}, even before any further simulation results.

When it comes to communicative success, types with $\Lbound$ have a functional advantage over
$\Llack$; literal $\Lbound$ has a slight advantage over pragmatic $\Lbound$; all have a sizable
advantage over literal $\Llack$. The red arrows in Fig.~\ref{fig:sketch} sketch trajectories we
can expect if there is only pressure for communicative success. How much each replicator step
moves a population through this space ultimately depends on the rationality parameter
$\lambda$. Functional differences decrease as $\lambda$ increases: if $\lambda$ is sufficiently
high, all types except for literal $\Llack$ speakers will show a strong tendency to associate
$\ssome$ with $\msome$ and $\sall$ with $\mall$.

Under probability matching $l=1$, iterated Bayesian learning converges to the prior over types
\citep{griffiths+kalish:2007}. If the prior on types is uniform, populations gravitate toward
the space's center under repeated mutator steps. This is sketched by the blue dash-dotted
arrows in Fig.~\ref{fig:sketch}. If instead the prior favors $\Llack$ over $\Lbound$, we expect
repeated mutator steps to drive the population toward the upper-half of the space. This is
suggested by the solid purple arrows in Fig.~\ref{fig:sketch}. The system's predictions are
more difficult to predict when $l>1$, because now individual mutator steps may also be
influenced not only by the priors, but also by the differential likelihood with which a
particular type $i$ can be confused for any other type $j$ (see Section~\ref{sec:learnability}
and Appendix~\ref{sec:example-iter-bayes}).

Despite its simplicity, the diagram in Figure~\ref{fig:sketch} makes clear the sense in which
semantics and pragmatics might co-evolve. Pragmatic language use will not evolve under a strong
and fully expressive semantics (bottom part of Figure~\ref{fig:sketch}). A weak semantic
convention will not evolve without disposition towards pragmatic enrichments (left part of
Figure~\ref{fig:sketch}). But together a weak semantics and a pragmatic disposition for
enrichment can co-evolve. These abstract theoretical considerations also suggest that we need
both fitness based selection of communicative efficiency and pressure towards learnability to
see underspecified semantics and pragmatic language use emerge together.

\begin{figure}[t]
\centering
\includegraphics[width=1.08\textwidth,height=10cm, keepaspectratio]{./plots/fig-contourb105k5.png}
\caption{Dynamics on a two-dimensional type space with four types. The $y$-axis represents the
  proportion of types with $\Llack$; the $x$-axis that of pragmatic ones. Arrows indicate
  directionality of trajectories after (a) replicator steps and (b) mutator steps with a prior
  that favors types with $\Llack$ by a factor of $1.05$ over those with $\Lbound$. Colored
  contours show the proportion of pragmatic $\Llack$ users after a step. \tb{I don't think the
    contours are adding much, so they could be removed. Then again, I think they are rather
    non-intrusive, look nice, and illustrate how the step-size, depending on where we are in
    the space, varies.} \mf{I think that the contours are sweet. But I worried about the size
  of the arrows. I made the picture a bit bigger}}

\label{fig:quiver}
\end{figure}

Fig.~\ref{fig:quiver} shows actual evolutionary trajectories in this type space. The replicator
steps shown in the top row of Fig.~\ref{fig:quiver} make the functional disadvantage of
$\Llack$ against $\Lbound$ apparent, particularly for its literal users and/or low
$\lambda$. They also highlight that the functional difference between literal and pragmatic
$\Lbound$ is small. Under functional pressure only, populations move toward the lower part of
the square but the steps become small as the lower edge, representing populations composed only
of $\Llack$ users, are reached. As for the mutator diagrams (bottom row of
Fig.~\ref{fig:quiver}), there are three main things to note. First, as expected from the
literature, a small bias for a type will lead to larger proportions to adopt it under $l=1$
(see lower left space in Figure~\ref{fig:quiver}). Second, if $l>0$, the dynamic is influenced
heavily also by asymmetries in production likelihoods. Third, these asymmetries in the
production likelihoods are in turn influenced by $\lambda$, so that $\lambda$ also has an
effect on iterated Bayesian learning (compare the middle and right-most diagram in the bottom
row of Fig.~\ref{fig:quiver}).

Fig.~\ref{fig:sketch} and \ref{fig:quiver} investigate replicator and mutator steps in
isolation. It remains to investigate how exactly these forces shape a population over time when
they apply together. This will also hinge on which types are present because the space of all
types determines whether production likelihoods asymmetrically favor certain types over others
(when $l>1$; see Appendix~\ref{sec:example-iter-bayes}). For this reason, we turn to a larger
setup next with a larger and more realistic set of types. Additionally, we want to explore
the effects of a learning bias in favor of simple lexical representations. One way of
motivating and formalizing such a bias is introduced in Section~\ref{sec:an-induct-learn}.

\subsection{Setup}
\label{sec:setup:-stat-conc}

\subsubsection{States, messages, lexical representations, and lexica} 
\label{sec:stat-mess-conc}

Consider a state space with three states $\States = \set{\snone, \ssome, \sall}$ and think of
it as a partition of possible worlds into cells where none, some or all of the $A$s are $B$s,
for some arbitrary fixed predicates $A$ and $B$. Eight lexical representations can be
distinguished based on their truth or falsity in three world states, six of which are not
contradictory or tautological (see Table~\ref{tab:concepts} below).

A lexicon $L$ is a mapping $\Messgs \rightarrow R$ from messages to representations. With three
messages there are $6^3 = 216$ possible lexica. Some assign the same representations to more
than one message and others lexicalize the same representations but associate them with
different messages. Out of these possible lexica, three kinds are of particular
relevance. First, lexica that assign the same lexical representations to more than one
message. Such lexica lack in expressivity but may be favored by particular learning biases
nonetheless (see below). Second, lexica that conventionalize upper-bounds to realize a
one-to-one mapping of messages to states. Finally, lexica that do not lexicalize an upper
bound but allow it to be conveyed pragmatically due to the presence of a stronger lexical
item. There are six lexica of the second kind and six of the third.  The following three lexica
exemplify each kind:
\begin{align*}
  \begin{blockarray}{lccc}
    & & \underline{\Lall} & \\
    & \mygray{\mnone} & \mygray{\msome} & \mygray{\mall} \\
    \begin{block}{l[ccc]}
     \mygray{\snone}  & 0 & 0 & 0\\
     \mygray{\ssome}  & 0 & 0 & 0\\
    \mygray{\sall}   & 1 & 1 & 1\\
    \end{block}
% \\ & & P(\Lall)\rlap{ $\propto 9$} & \\
  \end{blockarray} & &
 \begin{blockarray}{lccc}
    & & \underline{\Lbound} & \\
    & \mygray{\mnone} & \mygray{\msome} & \mygray{\mall} \\
    \begin{block}{l[ccc]}
       & 1 & 0 & 0\\
       & 0 & 1 & 0\\
       & 0 & 0 & 1\\
    \end{block}
%  \\& & P(\Lbound)\rlap{ $\propto 96$} & \\
  \end{blockarray} & &
  \begin{blockarray}{lccc}
    & & \underline{\Llack} & \\
    & \mygray{\mnone} & \mygray{\msome} & \mygray{\mall} \\
    \begin{block}{l[ccc]}
       & 1 & 0 & 0\\
       & 0 & 1 & 0\\
       & 0 & 1 & 1\\
    \end{block}
%   \\& & P(\Llack)\rlap{ $\propto 48$} & \\
  \end{blockarray}
\end{align*}
Lexicon $\Lall$ is clearly bad for communication: all message and interpretation choices will
be equally likely; no information about the observed world state will be conveyed
by its users. In contrast, users of $\Lbound$ can communicate world states perfectly, no matter
whether they are literal or pragmatic users. Users of $\Llack$ can also communicate information
about the actual world state but need pragmatic language use to approximate a one-to-one
mapping between message use and states (see Section~\ref{sec:languages+use}).

Recall that types are a combination of a lexicon and a manner of language use. We analyze the
model's predictions in populations of types with one of the two behaviors introduced earlier:
literal or pragmatic. The former correspond to level-$0$ reasoners and the latter to level $1$. Accordingly, there is a total of $432$ types. Six types are pragmatic language
users with $\Llack$-like lexica. We refer to these as \emph{target types} because they
represent lexica and language use that conform to the majority view of scalar
implicatures. Twelve types are either literal or pragmatic users of lexica of the
$\Lbound$ kind. We refer to these as \emph{competitor types}, because they are expected to be
the target types' main contenders in evolutionary competition.

Note that while
different types may lexicalize the same representations, they may nevertheless map different
states to different overt messages. More informally, they speak different languages that
lexicalize the same concepts. Consequently, more often than not different types of the same
kind fail to understand each other.

\subsubsection{An inductive learning bias for semantic simplicity}
\label{sec:an-induct-learn}

There is a growing effort to develop empirically testable representational languages that allow
for the measure of semantic complexity. For instance, so-called {\em languages of thought}
(LOTs) have been put to test in various rational probabilistic models that show encouraging
results (see, e.g., \citealt{katz+etal:2008}, \citeauthor{piantadosi+etal:underreview} under review; 2012, and \citealt{piantadosi+jacobs:2016} for recent discussion). At its
core, a LOT defines a set of operations and composition rules from which lexical representations can be
derived. As a first approximation and for the sake of concreteness, we follow this approach to
motivate and formalize a preference of learners for simpler semantic representations
(\citealt{feldman:2000, chater+vitanyi:2003, piantadosi+etal:2012a,kirby+etal:2015}, \citeauthor{piantadosi+etal:underreview} under review). 
In a weighted generative LOT a representation's complexity is a function of its derivation cost.

\begin{table}
  \centering
  \begin{align*}
    R & \rightarrow_2 R \wedge R 
    & R & \rightarrow_2 \neg R  \\ 
    R & \rightarrow_1 X \subseteq X &
    R & \rightarrow_1 X \neq \emptyset &
    R & \rightarrow_1 X = \emptyset  \\
    X & \rightarrow_1 \set{A,B} & 
    X & \rightarrow_1 X \cap X & 
    X & \rightarrow_1 X \cup X    
  \end{align*}
  \caption{Toy grammar in a set-theoretic LOT with weighted rules.}
  \label{tab:grammar}
\end{table}

Our toy grammar of lexical representations is given in Table~\ref{tab:grammar}. This grammar uses basic
set-theoretic operations to form expressions which can be evaluated as true or false in states
$\snone$, $\ssome$, or $\sall$ from above. Applications of generative rules have a cost attached
to them. Here we simply assume that the formation of Boolean combinations of representations incurs 2
cost units, while all other rule applications incur only 1 cost unit. Table~\ref{tab:concepts}
lists all six lexical representations relevant here, their truth conditions and the simplest formula that
expresses this representation in the grammar in Table~\ref{tab:grammar}.

\begin{table}
  \centering
\begin{center}
  \begin{tabular}{lccclc}
    \toprule
    intuitive name
    & \snone
    & \ssome
    & \sall
    & least complex formula
    & complexity
    \\ \midrule
    ``all''
    & 0
    & 0
    & 1
    & $A \subseteq B$
    & $3$
    \\
    ``some but not all''
    & 0
    & 1
    & 0
    & $A \cap B \neq \emptyset \wedge A \neq \emptyset$
    & $8$
    \\    
    ``some''
    & 0
    & 1
    & 1
    & $A \cap B \neq \emptyset$
    & $4$
    \\
    ``none''
    & 1
    & 0
    & 0
    & $A \cap B = \emptyset$
    & $4$
    \\
    ``none or all''
    & 1
    & 0
    & 1
    & $\neg(A \cap B \neq \emptyset \wedge A \neq \emptyset)$
    & $10$
    \\
    ``not all''
    & 1
    & 1
    & 0
    & $\neg (A \subseteq B)$
    & $5$
    \\
    \bottomrule
  \end{tabular}
\end{center}
\caption{Available lexical representations and their minimal derivation cost.}
\label{tab:concepts}
\end{table}

A complexity measure for lexical representations from Table~\ref{tab:concepts} is used to define a
learning bias that favors simpler representations over more complex ones. The prior probability of a
type is just the prior probability of its lexicon. The prior of a lexicon is a function of the
complexity of the lexical representations in its image set. Lexica with simpler representations accordingly have a
higher prior. One simple way of defining such priors over lexica (and thereby types) is:
\begin{align*}
  P(L)  & \propto \prod_{r \in Im(L)} P(r)   \ \text{, with} & 
  P(r) & \propto \max_{r'}Compl(r') - Compl(r) + 1\,,
\end{align*}
where $Compl(r)$ is the complexity of the minimal derivation cost of representation $r$ according to the
LOT-grammar (see Table~\ref{tab:concepts}). Applied to our space of lexica, this construal assigns the highest probability 
to a lexicon of type $\Lall$, which only uses the simplest lexical representation ``all'' for
all messages. Lexica of type $\Llack$ are less likely, but more likely than $\Lbound$.

The menu of inductive biases argued to shape language acquisition is steadily being
refined. Apart from simplicity, prominent examples include mutual exclusivity
\citep{merriman+bowman:1989,clark:2009}, regularization \citep{hudson+etal:2005}, and
generalization \citep{smith:2011}. Even when these biases are considered in isolation, there
are many ways in which they can be translated into priors over types. The key assumption here,
common to all simplicity-biased priors, is that simple representational expressions should be
favored over more complex ones (see, e.g., \citealt{goodman+etal:2008, piantadosi+etal:2012a,
  kirby+etal:2015}). In view of these numerous possibilities, we should stress that these
details -- from the generative grammar to its complexity measure -- are to be regarded as one
convenient operationalization of one general approach to explicating learning biases; this is
not a commitment that this general approach is necessarily superior or that, within it, this
particular instrumentalization is the single most plausible. We return to this issue in
Section~\ref{sec:discussion}.

\subsection{Simulation results}
\label{sec:simulation-results}
Recall that there are three parameters held constant across types: soft-max parameter $\lambda$
affects how strongly speakers favor messages that appear best from their subjective point of
view; the bottleneck size $k$ influences how faithfully learners can identify their teacher
type; $l$ defines the learners' disposition towards choosing the most likely teacher type from
the posterior distribution. Based on considerations and results from
Section~\ref{subsec:example}, we expect that competitor types (types with lexica of the kind
$\Lbound$) have a fitness advantage over target types (pragmatic types with lexica of the kind
$\Llack$), especially for very low levels of $\lambda$ (see Fig~\ref{fig:sketch}). Selection
based on fitness alone may therefore not lead to prevalence of target types in the
population. On the other hand, lexica of type $\Llack$ are simpler than those of type $\Lbound$
by the postulated measure from above. This may make them more likely to be adopted by learners,
especially when $l$ is high. Still, lexica of the kind $\Lall$ are in turn even more likely
\emph{a priori} than lexica of the kind $\Llack$. Simulation results will shed light on the
question whether target types can emerge, and for which parameter constellations.

To better understand the workings of a pressure towards communicative efficiency and a pressure
towards learnability, we look at the behavior of the replicator and mutator step first in
isolation, and then in combination. All simulation runs are initialized with an arbitrary
distribution over types, constituting a population's first generation. All reported results are
the outcome of 50 update steps. These outcomes correspond to developmental plateaus in which
change is, if not absent, then at least very slow. In other words, even if the resulting states
do not correspond to an eventual attracting state, they characterize population states in which the system remains for a long time. As specified in Section~\ref{sec:learnability}, the
mutation matrix $Q$ can be obtained by considering all possible state-message sequences of
length $k$. Given that this is intractable for large $k$, the sets of data which learners are
exposed to are approximated by sampling $250$ $k$-length sequences from each type's production
probabilities.

\subsubsection{Replication only: selection based on communicative success}

\begin{figure}[t]
\centering
%\includegraphics[width=\textwidth,height=8cm, keepaspectratio]{./plots/alt-fig1-onlyr}
\includegraphics[width=\textwidth,height=8cm, keepaspectratio]{./plots/barh-onlyr}
\caption{Proportion of target types, the six competitor types with the highest proportion, the
  most frequent type of the $\Lall$-kind, and the population's majority type, in representative
  populations after $50$ generations under only a pressure for communicative success
  (replicator steps).}
\label{fig:only-R}
\end{figure}

Selection based on communicative success is sensitive to $\lambda$ since $\lambda$ influences
signaling behavior, which in turn determines communicative success. Fig. \ref{fig:only-R} shows
the proportion of target types, the highest competitor types and the highest type with $\Lall$
in a representative population after $50$ replicator steps. The plot also indicates the
proportion of the \emph{majority type}: the type with the highest proportion in the
population. With low $\lambda$ many types have very similar behavior, so that evolutionary
selection lacks grip and becomes very slow. The result is a very long transition with near
stagnancy in a rather homogeneous population with many types. Conversely, higher $\lambda$
promotes less-stochastic linguistic behavior, widening the gap in expressivity between some
types and promoting more homogeneous populations. As suggested by Fig. \ref{fig:only-R}, the
majority in most populations is not one of the six pragmatic $\Llack$-style types. That is, a
pressure only for communicative success does not lead to a prevalence of target types under any
$\lambda$-value. For instance, with $\lambda = 20$, $1000$ independent populations only had $11$
cases in which the target type was the majority type, corresponding to a mean proportion of
$.003$ across populations. By contrast, in $913$ cases the majority types had $\Lbound$ with
close to an even share between literal ($454$) and pragmatic types ($459$), corresponding to a
mean proportion of about $.48$ taken together. In sum, fitness-based selection of single types
requires sufficiently high $\lambda$ but does not often single out types with an underspecified
semantic representation. \mf{rewrite this paragraph or section}

\subsubsection{Iterated learning only: transmission fidelity and learnability}

\begin{figure}[t]
\centering
%\includegraphics[width=1\textwidth,height=8cm,keepaspectratio]{./plots/alt-fig2-onlym-pr}
\includegraphics[width=1\textwidth,height=8cm,keepaspectratio]{./plots/barh-onlym}

\caption{Proportion of target types, the six competitor types with the highest proportion, the highest type of the $\Lall$-kind, and the population's majority type in representative populations after $50$ generations under only a
  pressure for learnability ($\lambda = 20, k = 5$). The learning prior is shown in the
  right-most plot with top-most groupings corresponding to types, literal and pragmatic, with
  lexica of kinds $\Llack$, $\Lbound$, and $\Lall$.}
\label{fig:only-M}
\end{figure}

The effect of iterated learning without a pressure for communicative success using either
posterior sampling ($l = 1$) or a stronger tendency towards posterior maximization ($l = 15$)
is shown in Fig.~\ref{fig:only-M} together with the prior over types. The prior shows that
while users of $\Llack$ are not the most favored by the inductive bias (compared, e.g., to
$\Lall$) they are nevertheless more advantaged than others, such as $\Lbound$, in virtue of the
relatively simple semantics they conventionalize. Crucially, $\Llack$ enables its users to
convey each state with a single message when combined with pragmatic reasoning and sufficiently
high $\lambda$. This makes it less likely to be confused with other types if the learning data
is not too sparse ($k \geq 5$). What is more, learners have a propensity to infer pragmatic
$\Llack$ when the teacher's type produces very similar data, such as when using $\Lbound$ (see
Section~\ref{subsec:example}). Moreover, $\Llack$ is less likely to be confused with types with
different observable behavior because its pragmatic use approximates a one-to-one form-meaning
mapping. As a consequence, a stronger propensity to maximize the posterior increases their
proportion in the population.

However, in contrast to a pressure only for communicative success with high $\lambda$ (see
Fig.~\ref{fig:only-R}), learnability alone does not succeed in selecting for a single
prevalent type; all six target types tend to coexist at roughly equal proportion. Each
is passed on to the next generation with the same faithfulness and, differently from a pressure
for communicative success, they do not stand in competition with each other. In $1000$
independent populations with $\lambda = 20$ all majority types were target types,
with each reaching approximately the same proportion of users in the population.

As with a pressure only for communicative success, low values of $\lambda$ make the differences in observable behavior across types less pronounced. This makes differences in the likelihood of particular types having generated a learning input less pronounced. Therefore, low $\lambda$ leads populations to reflect the learners' inductive bias more faithfully. This favors functionally deficient but a priori preferred types such as those that use  $\Lall$. A pressure for learnability alone may consequently lead to a spread of communicatively suboptimal types that lexicalize simpler semantics. In the extreme, when $l = 1$ and $\lambda = 1$ all
of $1000$ independent populations had users of $\Lall$ as majority types. For higher $\lambda$, it becomes clear that a high prior is not the only thing that counts when it comes to learnability (see Section~\ref{subsec:example}). As soon as there is information for learners to discern whether one type is more likely to have generated the data (depending on $\lambda$ and $k$), it becomes paramount for types to produce data that makes them easily identifiable if they are to be inferred more often.

\subsubsection{Combining pressures of communicative success and learnability}

Pressure for communicative success and learnability are not sufficient on their own to have a
single target type dominate the population. When pressured for communicative success, the  communicative advantage of $\Lbound$ users leads to their prevalence. When pressured for
learnability, pragmatic $\Llack$ is promoted over functionally similar but semantically more
complex alternatives such as $\Lbound$. However, learnability alone does not foment the
propagation of a single target type across the population.

\begin{figure}[t]
\centering
\includegraphics[width=1\textwidth,height=8cm,keepaspectratio]{./plots/alt-fig3-r+m}
\caption{Proportion of target types, the six competitor types with the highest proportion, the highest type of the $\Lall$-kind, and the population's majority type in representative populations after $50$ generations under both pressures ($k = 5$).}
\label{fig:rmd}
\end{figure}

Combining pressures for communicative efficiency and learnability can lead to the selection of
a single target type (see Fig.~\ref{fig:rmd}). The proportion of a single majority target type
increases with $\lambda$ and $l$.

As before, low $\lambda$ and $l$ lead to the prevalence of communicatively suboptimal types
that are a priori favored, such as $\Lall$. An increase in $\lambda$ leads to the selection of
target types but does not necessarily lead to monomorphic populations if learners sample from
the posterior (see the uppermost row in Fig.~\ref{fig:rmd}). Finally, a combination of high
$\lambda$ and $l$ leads to increasing proportions of a single majority target type. This joint
influence is summarized in Fig. \ref{fig:diff}, which shows the mean difference between the
most frequent target type and the proportion of the most frequent non-target type of a
different kind in $1000$ independent populations across $\lambda$ and $l$ values. Higher values
of $\lambda$ and $l$ increase the prevalence of a single target type, whereas lower values lead
to less pronounced differences, with a valley resulting from low $\lambda$ and high $l$
(cf. Fig. \ref{fig:rmd} with $\lambda = 1$ and $l=15$).

\begin{figure}[t]
\centering
\includegraphics[width=1\textwidth,height=10cm,keepaspectratio]{./plots/fig4-3d-incumbents-difference} %alternatively without '-3d'
\caption{Mean difference between proportion of highest target type and highest other type in
  $1000$ independent populations after $50$ generations under both pressures ($k = 5$).}
\label{fig:diff}
\end{figure}

Effects of manipulating the sequence length $k$ have not been addressed so far, but are rather
predictable: small values lead to more heterogeneous populations that reflect the learner's
prior more faithfully. This is due to the fact that the likelihood that a small sequence was
produced by any type is relatively uniform (modulo prior). By contrast, larger values
increasingly allow learners to differentiate types with different signaling behaviors.

In sum, when pressured for communicative success only, target types are outperformed by
competitor types. When pressured for learnability only, populations are polymorphic due to lack
of competition. When we combine both pressures, the (slight) functional disadvantage of targets
is counterbalanced by an advantage in learnability. This leads to large proportions of targets
and, due to the competition among types applied by each replicator steps, to more monomorphic
populations. To conclude, target types can come to dominate the population if three assumptions
are met: (i) language is pressured toward both communicative success {\em and} learnability;
(ii) pragmatic language use is an option; (iii) learners prefer simpler over more complex
lexical representations and exhibit a tendency towards the acquisition of the type that best
explains the learning data.

\section{General discussion}\label{sec:discussion}
The approach introduced here combines game-theoretic models of functional pressure towards
successful communication \citep{nowak+krakauer:1999}, effects of transmission perturbations on
(iterated) language learning \citep{griffiths+kalish:2007}, probabilistic speaker and listener
types of varied degrees of pragmatic sophistication \citep{frank+goodman:2012,
  franke+jaeger:2014}, as well as reasoning about unobservable lexical representations
\citep{bergen+etal:2012,bergen+etal:2016}. This allows for a conceptual investigation of the
(co-)evolution of conventional meaning and pragmatic language use. Main contributions of the
model are (i) its modular separation of communicative success and learnability on evolutionary
trajectories, (ii) the characterization of language learning as a joint inference over
linguistic behavior and lexical meaning, and (iii) the possibility to trace the co-evolution of
conventional semantics and pragmatic use.

With respect to (i), \citet{kirby+etal:2015} propose a comparable model of the interaction
between a lexicon's expressivity and its learnability. A main difference is that here we
considered communicative pressure for mutual understanding. This pressure can indirectly select
for expressive types, those that can convey states unequivocally, whereas Kirby et al. only
consider the bearing that the latter ability has on the production of learnable data. We see
three main reasons for considering communicative success rather than just expressivity, and for
looking at communication and learning rather than just learning. First, learning alone can
promote populations with non-negligible proportions of functionally defective types. This is
true both of simulations and of laboratory experiments with human subjects (see
\citealt{kirby+etal:2008,silvey+etal:2014}, and see \citealt{fay+etal:2013} for review of
laboratory results). Second and more importantly, types may be equally expressive but their
performance as a means of information transfer depends not only on themselves but on the
population they find themselves in (compare the competition of target types in Fig.
\ref{fig:rmd} and their lack of competition in Fig. \ref{fig:only-M}). That is, we contend that
adopting an expressive type that generates learnable data does not, in itself, capture a type's
arguably central communicative function of transferring information to peers. Taking
communication into consideration allows the model to be responsive to the task for which
language is learned. Lastly, chains of iterated learning alone do not put types in direct
competition. Accordingly, learning leads to polymorphous populations in which multiple types of
a kind coexist \citep{nowak:2006}.

The main result of our case study is that types that correspond to the majority view of scalar
implicatures -- scalar readings are non-lexicalized pragmatic enrichments -- can come to
dominate a population. This can happen if pragmatic language use is recruited indirectly by a
preference for simpler lexical representations (relative to more complex ones that lead to
comparable overt linguistic behavior without pragmatic language use). Under this view,
semantics and pragmatics play a synergic role and can co-evolve: pragmatic use allows
maintenance of simpler representations; pressure towards representational simplicity indirectly
promotes pragmatic over literal language use.

While the results of this case study are interesting, they also raise a number of critical
issues. First of all, while many favorable parameter settings exist which lead to a prevalence
of target types, other types are usually represented in non-negligible proportions as well (see
Fig.~\ref{fig:rmd} and \ref{fig:diff}). This may just be a technical quirk of the mutator step;
but there is a related issue of empirical importance. Several experimental studies on scalar
implicatures suggest that participants can be classified as either semantic or pragmatic users
of, in particular, \emph{some}
\citep[e.g.,][]{BottNoveck2004:Some-Utterances,NieuwlandDitman2010:On-the-incremen,DegenTanenhaus2012:Processing-Scal}. The
former consistently accept \emph{some} where \emph{all} would be true as well, the latter do
not. Interestingly, in our simulations when a target type is the majority type an inflated
proportion of the population uses compatible lexica with a lexicalized upper bound. In other
words, we find a tendency toward a similar co-existence of semantic and pragmatic
types. Whether this analogy has any further explanatory value is an interesting path for future
exploration.

Another important issue that is not addressed in the model are potential costs associated with
pragmatic reasoning. Here, we simply assumed that literal and pragmatic reasoning strategies
exist from the start and are equally costly to apply. In contrast, empirical results suggest
that the computation of a scalar implicature may involve additional cognitive effort
\citep[e.g.,][]{BrehenyKatsos2006:Are-Generalised,deNeys+schaeken:2007,huang+snedeker:2009,Jr.Bailey2013:Possibly-all-of}. Extensions
of the model presented here to include processing costs for pragmatic language use would be
interesting future work. It seems plausible that effects of reasoning cost may trade off with
the frequency with which a given scalar expression is used. It may be that frequently drawn
scalar implicatures lexicalize to avoid cost, whereas infrequent ones are derived on-line to
avoid more complex lexical representations during acquisition. Such a prediction would lend itself to
empirical testing in line with a recent interest in differences between various scalar
implicature triggers \citep{Tielvan-TielMiltenburgvan-Miltenburg2014:Scalar-Diversit}.

% Our case study could be criticized as follows: all it shows is that scalar implicatures do not
% lexicalize because upper bounds are dispreferred lexical representations. This criticism would be too
% superficial and highly unjust. Dispreferred lexical representations can thrive under evolutionary
% selection. Lexicalized upper-bounds can dominate a population because they may boost
% communicative success. But they do not have to. Moreover, as stressed throughout this investigation, even without selective pressure
% for communicative success, it is not necessarily the case that the types that are most
% likely \emph{a priori} will dominate. The dynamics of iterated learning are not that
% trivial. Iterated learning does not necessarily promote the \emph{a priori} more likely type,
% but tends to promote a type $t$ based on a gradient of how many other types might likely mutate
% into $t$, so to speak. Taken together, without an explicit model of the interaction between
% pressure for communicative success and learnability, it is far from trivial to judge whether or when
% preferred or dispreferred representations can be adaptive. This is why a major contribution of this
% paper is the arrangement of many different ingredients into a joined model of the (co-)evolution
% of lexical meaning and pragmatic use.

We tried to motivate and formalize a general assumption about lexical representations'
complexity with a concrete, albeit provisional proposal. The specification of a learning bias
in terms of a ``grammar of representations'' can and should be seen critically, however. Much
depends on the primitives of such a grammar. For instance, the lexical representation ``none or
all'' is the most complex in Table~\ref{tab:concepts}. But consider adding a new primitive
relation between sets $A \smile B$ which is true if and only if
$\neg(A \cap B \neq \emptyset \wedge A \neq \emptyset)$. The lexical representation ``none or
all'' would then be one of the simplest. Clearly, further research, empirical and conceptual,
into the role of representational complexity, processing costs and learning biases is
needed. The model here makes a clear and important contribution nonetheless: it demonstrates
how simplicity of representations can interact with use and evolutionary selection and shows
that for simple representations to emerge it may require pragmatic strategies to compensate
their potential expressive deficiencies. Hence a model of co-evolving semantics and pragmatics
is needed. Future work should also include the possibility that representational simplicity may
itself be a notion that is subject to evolutionary pressure
\citep[cf.][]{ThompsonKirby2016:Culture-Shapes-}, as well for the evolution of elements that
define the agents' cognitive make-up; $\lambda$ and $l$.

This case study should not be interpreted as a proposal for a definite explanation of how
scalar implicatures evolved. Other factors should be considered eventually even if they will
lead to much more complex modeling. One such factor is the observation that non-lexicalized
upper bounds allow a broader range of applicability. For example, when the speaker is not
certain as to whether \emph{all} is true. This may suggest an alternative and purely
functionalist argument for why upper-bounded meanings do not conventionalize: should contextual
cues provide enough information to the hearer to identify whether a bound is intended to be
conveyed pragmatically, then this is preferred over expressing it overtly through longer
expressions. For example, by saying {\em some but not all} explicitly. Importantly, although
morphosyntactic disambiguation may be dispreferred due to its relative length and complexity
\citep{piantadosi+etal:2012b}, it allows speakers to enforce an upper-bound and override
contextual cues that might otherwise mislead the hearer. In a nutshell, this explanation posits
that scalar implicatures fail to lexicalize because, all else being equal, speakers prefer to
communicate as economically as possible and pragmatic reasoning enables them to do so. What
this alternative argument does not explain is why functional pressure does not lead to the
emergence of different, equally costly lexical items to express different knowledge states of
the speaker (\citealt[252-267]{horn:1984},
\citealt{horn:1972,traugott:2004,vdAuwera:2010}). That is, this argument does not explain why
English and other languages do not have a monomorphemic dual for, e.g., {\em some} that
lexicalizes an upper-bound. If this hypothetical expression existed, it could be deployed to
signal that the speaker knows that {\em some but not all} holds, and unbounded {\em some} could
exclusively signal epistemic uncertainty. Looking at pressure from learnability might come in
again.

Beyond scalar implicatures, the model can generate predictions about likely lexicalization
trajectories of pragmatic inferences, or a lack thereof. In this realm an interesting issue is
whether proposed principles, such as the semantic conventionalization of once highly
context-dependent inferences, if they become regular enough
\citep{levinson:2000,traugott:2004}, can be given a formal rationale and inform postulated
directionalities of change. The present investigation made a first start and gave a framework
for exploring these issues systematically.

\section{Conclusion}
The cultural evolution of meaning is influenced by intertwined pressures. We set out to
investigate this process by putting forward a model that combines pressure toward successful
information transfer with perturbations that may arise in the transmission of linguistic
knowledge in acquisition. Its objects of selection and replication are pairs of lexical
meanings and patterns of language use. This allows the model to trace the evolutionary interaction between
conventional meaning and pragmatic use. Additionally, it takes the challenge seriously of
neither semantics nor pragmatics being directly observable. Instead, learners need to infer
these unobservables from overt data that results from their combination.  These components and
their mutual influence were highlighted in a case study on the lack of lexical upper-bounds in
weak scalar expression. This study showed that, when pressured for learnability and communicative
success, the former force can drive for simpler semantic representations inasmuch as pragmatics
can compensate for lack of expressivity in use. That is, the relative learning advantage of
simpler semantics in combination with functional pressure in use may offer an answer to why
natural languages fail to lexicalize systematic pragmatic inferences. And, more broadly, lead to a
division of labor between semantics and pragmatics.

\appendix

\section{A minimal example of the replicator mutator dynamic}
\label{sec:minim-example-repl}

\begin{figure}[t]
  \centering
  %% notice that the figure is hacked, because somehow the curve does not come out right under
  %% PGFplots; I don't know why!?!?

    \begin{subfigure}[b]{0.45\textwidth}
      \centering
      \begin{tikzpicture}
      \begin{axis}[xlabel=$x$, ylabel = $x'$,
                   grid=major, width=6.85cm, height = 5.5cm,
                   legend style = {legend pos = north west},
                   ymin=0,ymax=1, xmin=0,xmax=1]

      \addplot[smooth, color=blue] {x^2 / (3*x^2 - 4*x + 2) + 0.019 * x};
      \addplot[smooth, color=green] {.8*x + .1};

      \addplot[smooth, color=red] {(.9*x^2 - .2*x +.2)/(3*x^2 - 4*x + 2)};

      \addplot[smooth, color=gray] {x};

      \legend{RD, M, RMD}

      \end{axis}
    \end{tikzpicture}

        \caption{Update functions: the population state $x$ is mapped onto $x'$ in one update step.}
        \label{fig:Updates_RMS}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.5\textwidth}    
      \centering
    \begin{tikzpicture}[x=120]
      \node[draw=black, circle, fill = black, minimum size = 0.25cm]
      (0) at (0,0) {};

      \node[draw=black, circle, fill = black, minimum size = 0.25cm]
      (1) at (1,0) {};

      \draw[-, thick] (0) -- (1);

      \node[draw=black, circle, fill = white, minimum size = 0.25cm, thick]
      (mid) at (2/3,0) {};

      \node[]  (0.label) at (0,-0.5) {0};

      \node[]  (1.label) at (1,-0.5) {1};

      \node[]  (label) at (0.5,-0.5) {$x$};

      \node[]  (label) at (0.5,0.55) {RD};

      \draw[->-=0.8,very thick] (mid) -- (0);

      \draw[->-=0.8,very thick] (mid) -- (1);

    \end{tikzpicture}

    \vspace*{0.35cm}

    \begin{tikzpicture}[x=120]

      \node (0) at (-0.05,0) {};

      \node (1) at (1.05,0) {};

      \draw[-, thick] (0) -- (1);

      \node[draw=black, circle, fill = black, minimum size = 0.25cm, thick]
      (mid) at (0.5,0) {};

      \node[]  (0.label) at (0,-0.3) {0};

      \node[]  (1.label) at (1,-0.3) {1};

      \node[]  (label) at (0.5,-0.5) {$x$};

      \node[]  (label) at (0.5,0.55) {M};

      \draw[->-=0.5,very thick] (0) -- (mid);

      \draw[->-=0.5,very thick] (1) -- (mid);

    \end{tikzpicture}

    \vspace*{0.35cm}

    \begin{tikzpicture}[x=120]

      \node (0) at (-0.05,0) {};

      \node (1) at (1.05,0) {};

      \node[draw=black, circle, fill = black, minimum size = 0.25cm]
      (S0) at (0.121,0) {};

      \node[draw=black, circle, fill = black, minimum size = 0.25cm]
      (S1) at (0.903,0) {};

      \draw[-, thick] (0) -- (1);

      \node[draw=black, circle, fill = white, minimum size = 0.25cm, thick]
      (mid) at (0.609,0) {};

      \node[]  (0.label) at (0,-0.3) {0};

      \node[]  (1.label) at (1,-0.3) {1};

      \node[]  (label) at (0.5,-0.5) {$x$};

      \node[]  (label) at (0.5,0.55) {RMD};

      \draw[->-=0.5,very thick] (mid) -- (S0);

      \draw[->-=0.5,very thick] (mid) -- (S1);

      \draw[->-=0.5,very thick] (0) -- (S0);

      \draw[->-=0.5,very thick] (1) -- (S1);

    \end{tikzpicture}

    \caption{Phase portraits for RD, M and RMD: unstable rest points are hollow, attractors are
      solid.}
        \label{fig:Phase_RD}
    \end{subfigure}

  \caption{Example.}
  \label{fig:Example_RMD}
\end{figure}

Consider a simple and abstract coordination game. Agents are of two types: positive or
negative. If agents of different type interact with each other, they obtain a payoff of $0$. If
negative meets negative, each receives a payoff of $1$. If positive meets positive, they get a
payoff of $2$.  A population state is completely characterized by the proportion $x$ of
negatives. The fitness of negatives in population state $x$ is $f_n(x) = x$, that of positives
is $f_p(x) = 2-2x$.  The average fitness is
$\Phi(x) = x f_n(x) + (1-x) f_p(x) = 3x^2 - 4x + 2$. Without mutation, the replicator dynamic
will update $x$ to $\text{RD}(x) = \nicefrac{f(x)x}{\Phi(x)} = \nicefrac{x^2}{\Phi(x)}$. The
update function $\text{RD}(x)$ of the replicator step is plotted in Fig.~\ref{fig:Updates_RMS}
as the blue line. Rest points, for which $\text{RD}(x)=x$, are at $x=0$, $x=1$ and
$x= \nicefrac{2}{3}$.  The former are attractors; nearby points converge to them. Points near
$x=\nicefrac{2}{3}$ also move towards 0 or 1. This is schematically pictured in the topmost
phase portrait in Fig.~\ref{fig:Phase_RD}. Adding mutation changes the dynamic and its rest
points. Let us assume that $Q_{ji} = .9$ when $j=i$. This is the proportion of types that are
replicated faithfully. Conversely, a proportion of $.1$ will change their type from positive to
negative, and vice-versa. The update effect of mutation on its own are described by
$\text{M}(x) = .9 x + .1 (1-x) = .8x + .1$, plotted as the linear green line in
Fig.~\ref{fig:Updates_RMS}. As shown in Fig.~\ref{fig:Phase_RD}, in this example mutation alone
has only one stable rest point. It is located at $x = .5$. If we first take the replicator step
and then the mutation step in sequence, we obtain the replicator mutator dynamic
$\text{RMD}(x) = \text{M}(\text{RD}(x)) = \nicefrac{.9x^2 - .2x +.2}{3x^2 - 4x + 2}$, which is
plotted in red in Fig.~\ref{fig:Updates_RMS}. The rest points are at $x=.121$, $x=.903$ and
$x=.609$. The former two are attractors (see Fig.~\ref{fig:Phase_RD}).

\section{Deriving the replicator mutator dynamic from conditional imitation}
\label{sec:deriv-repl-mutat}

Consider a huge, virtually infinite population of agents. At any point in time each agent has a
fixed type $i$ which determines its fitness $f_i$ at that moment in time when playing with an
arbitrary member of the population. The relative frequency of type $i$ is $x_i$ with
$\sum_i x_i = 1$. The average population fitness is $\Phi = \sum_k x_k f_k$. The mutation
probability from $j$ to $i$ is $Q_{ji}$, with $\sum_k Q_{jk} = 1$. The proportion of type $i$
at the next time step $x'_i$ is given by the discrete-time replicator mutator dynamic:
\begin{align}
  \label{eq:RMD_discrete_repeated}
  x'_i = \sum_j \frac{x_jf_j Q_{ji}}{\Phi}\,,
\end{align}
We will derive this dynamic from an agent-level update scheme of conditional imitation. There
are many variants on this derivation. Our choice is to be as simple as
possible. \citet{Sandholm2010:Population-Game} provides much more detail and discusses many
further interesting derivations of similar evolutionary dynamics.

A simple conditional imitation dynamic runs as follows. Every now and then an agent is selected
at random from the population and given a chance to possibly change its behavior. Call this
agent (perhaps slightly misleadingly) \emph{learner} $L$. The learner chooses a random agent
from the population to possibly imitate. Call this agent (somewhat more misleadingly)
\emph{teacher} $T$. Whether $L$ actually chooses to imitate the behavior of $T$ depends on how
good $T$'s behavior is at the moment. For simplicity, we assume that $L$ adopts $T$'s type with
probability $f_j$, where $j$ is the type of $T$.\footnote{We assume here that $f_i \in [0;1]$
  for all $i$. This is actually the case for the definition of fitness as communicative success
  from Section~\ref{sec:fitn--learn}. See \citet{Sandholm2010:Population-Game} for more complex
  alternative conditional update rules that still derive the same population dynamic.}  If $L$
chooses to imitate $T$'s behavior, it may not be able to directly observe $T$'s actual
type. Rather it may have to infer $T$'s latent type from its overt behavior. The probability
that $L$ actually acquires type $k$ when it chose to imitate a teacher of type $j$ is
$Q_{jk}$. In sum, we assume that learners have (at least roughly) an estimate of how good
another agent's behavior is, decide to imitate that behavior conditionally on how good it is
and the adopt whatever mechanism (type) they infer may cause the observed behavior.

This model of occasional conditional imitation gives rise to expected changes in relative
frequencies in the population. Given a huge population, the expected probability
$P(k \rightarrow i)$ that some agent of type $k$ switches by conditional imitation to type $i$
is the product of probabilities for the following four events:
\begin{enumerate}
\item a learner $L$ of type $k$ is sampled, which happens with probability $x_k$
\item a teacher $T$ of type $j$ is sampled, which happens with probability $x_j$
\item $L$ chooses to imitate $T$, which happens with probability $f_j$
\item $L$ infers $T$'s type to be $i$, which happens with probability $Q_{ji}$
\end{enumerate}
The average expected change $\dot{x}_i$ in proportion of type $i$ is then:
\begin{align*}
  \dot{x}_i & = \sum_k P(k \rightarrow i) - \sum_k P(i \rightarrow k) \\ 
  & = \sum_k x_k \sum_j x_j f_j Q_{ji} -   \sum_k x_i \sum_j x_j f_j Q_{jk} \\
  & = \sum_j x_j f_j Q_{ji} -   x_i \sum_j x_j f_j \\
  & =   (x'_i - x_i) \Phi
\end{align*}
In this way, the discrete-time replicator mutator equation in (\ref{eq:RMD_discrete_repeated})
captures the expected difference in relative frequency of type $i$ from one discrete time step
to the next (relative to current baseline fitness $\Phi$, which is constant for all types, and
so only governs the absolute speed of change).

\section{Example of iterated Bayesian learning}
\label{sec:example-iter-bayes}

We consider three types and three possible data observations they can produce. Types are
equally likely \emph{a priori}, but show different and crucially asymmetric production
behavior. The likelihood with which type $t_i$ produces datum $d_j$ is $LH_{ij}$:
\begin{align*}
  LH & = \begin{blockarray}{lccc}
    & d_1 & d_2 & d_3 \\
    \begin{block}{l[ccc]}
      t_1 & 1-2\epsilon & \epsilon & \epsilon \\
      t_2 & \delta & 1-(\epsilon + \delta) & \epsilon \\
      t_2 & \delta & \epsilon & 1-(\epsilon + \delta)  \\
    \end{block}
  \end{blockarray}
\end{align*}
If $\delta > \epsilon$, production likelihoods are asymmetrically skewed to possibly favor
$t_1$, in the sense that both types $t_2$ and $t_3$  produce behavior that is strong
evidence for $t_1$ more often than behavior that is strong evidence for any other
type other then themselves. Type $t_1$ makes no such ``asymmetric mistakes''.

Using the definitions from Section~\ref{sec:learnability}, we calculate mutation matrices $Q$ for
$l=1$ and $l=10$ when $\epsilon = 0.1$ and $\delta = 0.2$:
\begin{align*}
  Q_{l = 1} & \approx \begin{blockarray}{lccc}
    & t_1 & t_2 & t_3 \\
    \begin{block}{l[ccc]}
      t_1 & 0.56 & 0.22 & 0.22 \\
      t_2 & 0.22 & 0.59 & 0.19 \\
      t_2 & 0.22 & 0.19 & 0.59  \\
    \end{block}
  \end{blockarray} & 
  Q_{l = 10} & \approx \begin{blockarray}{lccc}
    & t_1 & t_2 & t_3 \\
    \begin{block}{l[ccc]}
      t_1 & 0.8 & 0.1 & 0.1 \\
      t_2 & 0.2 & 0.7 & 0.1 \\
      t_2 & 0.2 & 0.1 & 0.7  \\
    \end{block}
  \end{blockarray}
\end{align*}
Despite flat priors over types, the mutation matrix for $l=10$ is skewed towards $t_1$ in the
sense that the probability of any type mutating to $t_1$, $\sum_i Q_{i1} = 1.2$, is higher than
the probability of $\sum_i Q_{ij} = 0.8$ for $j \in \set{2,3}$. For $l=1$, these probabilities
are all equal for all three types. Consequently, under iterated Bayesian learning with $l=1$
the population will converge to a uniform distribution over all three types in the limit, but
for $l=10$ it converges to a skewed population vector $\vec{x} \approx \tuple{0.5, 0.25, 0.25}$
where $t_1$ is a clear majority type. This demonstrates how iterated Bayesian learning with
$l>1$ can be sensitive to the set of types represented in the model, in particular when
production likelihoods are asymmetrically in favor of some types. With probability matching
learners ($l = 1$), these considerations do not play a role, as the system will gravitate to
the prior distribution over types nonetheless.

\bibliographystyle{chicago}
\bibliography{./bounds-rmd}

\end{document}